---
title: "Figure 4 - gene networks"
author: "Kimberly Olney"
date: "2023-07-13"
output: html_document
---
# Setup
```{r setup}
knitr::opts_knit$set(root.dir = ".")
```
# User defined variables
```{r set_variables}
source(here::here("scripts/R", "file_paths_and_colours.R"))
library("readxl")
library(WGCNA)
library(anRichmentMethods)
library(tidyverse)
library(CorLevelPlot)
library(gridExtra)
library("limmaDE2")

condition <- c("TYPE")
tool <- "star"

addSmallLegend <- function(myPlot, pointSize = 4, textSize = 7, spaceLegend = .5) {
    myPlot +
        guides(shape = guide_legend(override.aes = list(size = pointSize)),
               color = guide_legend(override.aes = list(size = pointSize))) +
        theme(legend.title = element_text(size = textSize), 
              legend.text  = element_text(size = textSize),
              legend.key.size = unit(spaceLegend, "lines"))
}
```
# Read in data
```{r}
voomCounts <- readRDS(paste0("../../../rObjects/", condition, ".voomCountsMatrix.rds"))
load("../../../rObjects/bwnet.Rdata")
load("../../../rObjects/WGCNA_allSampleInput.RData")
module_eigengenes <- bwnet$MEs
head(module_eigengenes) # inspect 
ngenes_per_module <- table(bwnet$colors) # number of genes in each module 

dge.filtered.norm <- readRDS(paste0("../../../rObjects/dge.filtered.norm.rds"))
info <- as.data.frame(dge.filtered.norm$samples)
genes <- as.data.frame(dge.filtered.norm$genes)

# keep only the columns that hold information we need. 
dataTraits <- info[, c(6,11:19,21,22,24,29,37,108,109,114,117:124)] # Keep
dim(dataTraits)
# disease type isn't numeric, make numeric as this is required down stream
numeric_traits <- sapply(dataTraits, as.numeric)
numeric_traits <- as.data.frame(numeric_traits) # save as data frame
rownames(numeric_traits) <- rownames(dataTraits) # add NPID as row names
```
# Plot dendrogram
This function plots a hierarchical clustering dendrogram and color annotation(s) of objects in the dendrogram underneath.
guideHang - fraction of the dendrogram height to leave between the top end of the guide line and the dendrogram merge height. 
```{r plot_dendrogram}
plotDendroAndColors(bwnet$dendrograms[[1]], cbind(bwnet$unmergedColors, bwnet$colors), 
                    c("unmerged", "merged"), 
                    dendroLabels = FALSE, 
                    addGuide = TRUE, 
                    hang = 0.03,
                    guideHang = 0.05)

path <- paste0("../../../results/manuscript_figures/Figure_4_dendrogram")
saveToPDF(paste0(path, ".pdf"), width = 6, height = 4)

```
# Association between modules and traits
Which modules are associated with our traits

binarizeCategoricalColumns - Given a data frame with (some) categorical columns, this function creates a set of indicator variables for the various possible sets of levels.

includePairwise - logical should pairwise binary indicators be included? For each pair of levels, the indicator is val1 for the lower level (earlier in levelOrder), val2 for the higher level and NA otherwise.
includeLevelVsAll - Logical: should binary indicators for each level be included? The indicator is val2 where x equals the level and val1 otherwise.
minCount - 	Levels of x for which there are fewer than minCount elements will be ignored.
```{r binary_traits}
controlvsdisease <- binarizeCategoricalColumns(dataTraits$TYPE, 
                           includePairwise = TRUE, 
                           includeLevelVsAll = TRUE, 
                           minCount = 1)
# remove data. from column names
names(controlvsdisease) <- gsub(x = names(controlvsdisease), 
                                pattern = "data.", replacement = "")  
traits <- cbind(numeric_traits, controlvsdisease) # column combine with dataTraits
# disease type isn't numeric, make numeric as this is required down stream
numeric_traits <- sapply(traits, as.numeric) # make sure all are numeric 
numeric_traits <- as.data.frame(numeric_traits) # save as data frame
rownames(numeric_traits) <- rownames(dataTraits) # add NPID as row names
```
# Correlation
Add multiple test correction - BH 
```{r correlation}
nSamples <- nrow(counts) # number of samples
nGenes <- ncol(counts) # number of genes

# correlation between modules and traits 
module_trait_corr <- cor(module_eigengenes, numeric_traits, use = 'p')
# get the pval
# Calculates Student asymptotic p-value for given correlations.
module_trait_corr_pval <- corPvalueStudent(module_trait_corr, nSamples)
module_trait_corr_pval
module_trait_corr_pval_df <- as.data.frame(module_trait_corr_pval)
unadjust <- module_trait_corr_pval_df$AD.vs.CONTROL
adjust <- p.adjust(module_trait_corr_pval_df$AD.vs.CONTROL, method = p.adjust.methods, n = length(module_trait_corr_pval_df$AD.vs.CONTROL))

library(magrittr)
module_trait_corr_pval_fdr <- module_trait_corr_pval %>% 
     as.vector %>% 
     p.adjust(method='fdr') %>% # FDR is Benjamini & Hochberg
     matrix(ncol=35)

colnames(module_trait_corr_pval_fdr) <- colnames(module_trait_corr_pval)
rownames(module_trait_corr_pval_fdr) <- rownames(module_trait_corr_pval)
```
# Heatmap
```{r heatmap}
heatmap.data <- merge(module_eigengenes, numeric_traits, by = "row.names")
heatmap.data <- heatmap.data %>%
  column_to_rownames(var = 'Row.names')

# some data has missing values with na, such as lewy body count. 
# Replace na with zero 
heatmap.data["Cing.LB"][is.na(heatmap.data["Cing.LB"])] <- 0
heatmap.data["Cing.Synuclein"][is.na(heatmap.data["Cing.Synuclein"])] <- 0
plot_colors <- names(heatmap.data)[1:15]
plot_colors <- gsub("ME", "", plot_colors)

#par(mar= c(5, 5, 3, 3)) # bottom, left, top and right 
cor_disease_plot <- CorLevelPlot(heatmap.data, 
             x = c(names(heatmap.data)[43:44], names(heatmap.data)[47]), # which traits
             y = names(heatmap.data)[1:15], # gene modules 
             col = c("blue1", "skyblue", "white", "pink", "red"),
             rotLabX = 45, 
             colLabY = c("black"), #plot_colors
            # main = "Module-trait relationships", 
            # cexMain = 1, 
             fontLabX = 1,
             fontLabY = 1) # color scale
cor_disease_plot
```
# Pretty heatmap correlation
```{r}
numeric_traits_selective <- numeric_traits[, c(27:29,32)] # Keep
module_trait_corr_selective <- module_trait_corr[, c(27:29,32)] # Keep
module_trait_corr_pval_selective <- module_trait_corr_pval_fdr[, c(27:29,32)] # Keep
#Print correlation heatmap between modules and traits
textMatrix= paste(signif(module_trait_corr_selective, 2), "\n(",
                        signif(module_trait_corr_pval_selective, 1), ")", sep= "")

data  <- melt(module_trait_corr_selective)
#data$Var2 <- factor(data$Var2, levels = c("PA.vs.CONTROL",  "AD.vs.CONTROL", "LBD.vs.CONTROL", "LBD.vs.AD"))
data$Var2 <-
      gsub(
       ".vs.",
        " vs ",
        data$Var2
      )
data$Var2 <-
      gsub(
       "CONTROL",
        "Control",
        data$Var2
      )
data$Var2 <- factor(data$Var2, levels = c("PA vs Control",  "AD vs Control", "LBD vs Control", "LBD vs AD"))

FDRq <- melt(module_trait_corr_pval_selective)
data$stars <- ifelse(FDRq$value < .0001, "***", ifelse(FDRq$value < .01, "**", ifelse(FDRq$value < .05, "*", " ")))
data$sig <- paste0(round(data$value, digits = 2), data$stars)
  
ME_heatmap <- ggplot(data = data) +
  geom_tile(aes(x = Var2, y =Var1, fill = value)) +
  geom_text(aes(x = Var2, y =Var1, label = sig), size = 2.8) +
  #scale_color_manual(values = c('black' = 'black', 'white' = 'white'), guide = "none") +
  scale_fill_gradientn(colours = c("blue","white","red"), 
                         values = rescale(c(-.3,.3)),
                         guide = "colorbar", limits=c(-.3, .3),
                         breaks = c(-.3, 0, .3), name = "Pearson's r") +
  scale_x_discrete(expand = c(0,0)) +
 theme_bw() +
    theme(strip.text = element_text(size = 8), 
          axis.text.y = element_text(size = 8, margin = margin(t = 0, r = 10, b = 0, l = 0)),
          axis.title.y = element_blank(), 
          axis.text.x = element_text(size = 8, color = "black", angle = 45, vjust = 1, hjust=1), 
          axis.title.x = element_blank(), 
          legend.text = element_text(size = 8),
          legend.title = element_text(size = 8),
          legend.margin=margin(0,0.5,0,0),
          legend.box.margin=margin(-10,-2,-10,-7.5), 
          plot.margin = margin(0.1, 0.3, 0.2, 0.3, "cm"), #t,r,b,l
          plot.title = element_text(size = 8, margin = margin(.1,.1,.4,0))) +
  ggtitle("Module-trait relationships") +
    theme(
    legend.position = "right", 
    legend.text = element_text(size =6), 
    legend.title = element_text (size = 8), 
    legend.margin=margin(0,0,1,0),
    legend.key.size = unit(0.5, "cm"))
ME_heatmap <- addSmallLegend(ME_heatmap)
ME_heatmap
```
# Plot gene count per module 
```{r gene_count_per_module}
color_counts <- as.data.frame(bwnet$colors) # color modules 
color_counts$gene_id <- row.names(color_counts) # gene_ids as row names
colnames(color_counts)[1] ="hcolors"
color_counts$hcolors <- factor(color_counts$hcolors, levels = rev(plot_colors)) # arrange to match previous plot 
rev_plot_colors <- rev(plot_colors)
gene_count_per_module <- ggplot(color_counts, aes(x=hcolors)) + 
    geom_bar(fill = rev_plot_colors) +
  theme_classic() +
      theme(strip.text = element_text(size = 8), 
          axis.text.y = element_text(size = 8),
          axis.title.y = element_text(size = 8),
          axis.text.x = element_text(size = 8), 
          axis.title.x = element_text(size = 8), 
          plot.margin = margin(0.1, 0.3, 0.2, 0.4, "cm"), #t,r,b,l
          plot.title = element_text(size = 8, margin = margin(.1,.1,.8,0), hjust = -.1)) +
    ggtitle("Gene count in modules") +
    xlab("Module") +
    ylab("Gene count") +
  theme(axis.text.x=element_text(color = "black", angle = 90, vjust = 0.5, hjust=1)) +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-.25, size = 2.8) 
gene_count_per_module
```
# igraph
### DEGs
```{r eval = FALSE}
# First read in DEGs from limma/voom 
# read in DEGs 
LBDvsControl <- read.delim("../../../results/star/DEGs/TYPE_LBDvsControl_gene_DEGs_FDRq0.05_logFC_0.25.txt")
ADvsControl <- read.delim("../../../results/star/DEGs/TYPE_ADvsControl_gene_DEGs_FDRq0.05_logFC_0.25.txt")
# PAvsControl <- read.delim("../../../results/star/DEGs/TYPE_PAvsControl_gene_DEGs_FDRq0.05_logFC_0.25.txt")
LBDvsAD <- read.delim("../../../results/star/DEGs/TYPE_LBDvsAD_gene_DEGs_FDRq0.05_logFC_0.25.txt")

### AD vs Control
df_ADvsControl <- merge(color_counts, ADvsControl, by = "gene_id", all = TRUE)
df_ADvsControl$FDRq <- ifelse(df_ADvsControl$adj.P.Val < 0.05, "< 0.05", "No difference")
df_ADvsControl$FDRq <- df_ADvsControl$FDRq %>% replace_na('Not DE')
colnames(df_ADvsControl)[2] ="hcolors"
#### LBD vs Control
df_LBDvsControl <- merge(color_counts, LBDvsControl, by = "gene_id", all = TRUE)
df_LBDvsControl$FDRq <- ifelse(df_LBDvsControl$adj.P.Val < 0.05, "< 0.05", "No difference")
df_LBDvsControl$FDRq <- df_LBDvsControl$FDRq %>% replace_na('Not DE')
colnames(df_LBDvsControl)[2] ="hcolors"
#### LBD vs AD
df_LBDvsAD <- merge(color_counts, LBDvsAD, by = "gene_id", all = TRUE)
df_LBDvsAD$FDRq <- ifelse(df_LBDvsAD$adj.P.Val < 0.05, "< 0.05", "No difference")
df_LBDvsAD$FDRq <- df_LBDvsAD$FDRq %>% replace_na('Not DE')
colnames(df_LBDvsAD)[2] ="hcolors"
```
### magenta graph all 
```{r eval = FALSE}
magenta_graph <- wgcna2igraph(net = bwnet, datExpr = counts,
                    modules2plot = c("magenta"),
                    colors2plot = c("magenta"),
                    kME.threshold = 0.75, adjacency.threshold = 0.1,
                    adj.power = 12, verbose = T,
                    node.size = 0, frame.color = NA, node.color = NA,
                    edge.alpha = .5, edge.width =1)
# Add gene names 
magenta_genes <- genes[c(genes$gene_id %in% V(magenta_graph)$name),]
V(magenta_graph)$gene_name <- magenta_genes$gene_name
plot(magenta_graph, edge.arrow.size=.2, edge.color="magenta",
     vertex.color="magenta", vertex.frame.color="#ffffff",
     vertex.label=V(magenta_graph)$gene_name, vertex.label.color="black") 
path <- paste0("../../../results/manuscript_figures/magenta")
saveToPDF(paste0(path, ".pdf"), width = 4, height = 4)
```
```{r test_loop}
module_colors <- unique(bwnet$colors)
module_colors <- module_colors[module_colors != "grey"] # too large to be meaningful 

module_colors <- c("cyan")
for (i in module_colors) {
set.seed(28)
module_graph <- wgcna2igraph(net = bwnet, datExpr = counts,
                    modules2plot = i,
                    colors2plot = i,
                    kME.threshold = 0.75, adjacency.threshold = 0.1,
                    adj.power = 12, verbose = T,
                    node.size = 0, frame.color = NA, node.color = NA,
                    edge.alpha = .5, edge.width =1)
#-- all network with all gene names 
# Add gene names 
module_genes <- genes[c(genes$gene_id %in% V(module_graph)$name),]
V(module_graph)$gene_name <- module_genes$gene_name
plot(module_graph, edge.arrow.size=.2, edge.color=i,
     vertex.color=i, vertex.frame.color="#ffffff",
     vertex.label=V(module_graph)$gene_name, vertex.label.color="black") 
path <- paste0("../../../results/star/WGCNA/network_graphs/", i)
saveToPDF(paste0(path, ".pdf"), width = 8, height = 8)


# LBD vs Control
graph_DEGs <- subset(df_LBDvsControl, hcolors == i & FDRq == "< 0.05")
V(module_graph)$selection <- V(module_graph)$gene_name %in% graph_DEGs$gene_name
DEG_graph <- subgraph.edges(graph=module_graph, eids=which(V(module_graph)$selection == "TRUE"), delete.vertices = TRUE)
plot(DEG_graph, edge.arrow.size=.2, edge.color=i,
     vertex.color=i, vertex.frame.color="#ffffff",
     vertex.label=V(DEG_graph)$gene_name, vertex.label.color="black", 
     edge.label=NULL, 
     vertex.label.font = ifelse(V(DEG_graph)$selection == "TRUE", 2, 1)) 

plot(DEG_graph, vertex.label = ifelse(V(DEG_graph)$selection == "TRUE", V(DEG_graph)$gene_name, "."), vertex.label.color="black")

path <- paste0("../../../results/star/WGCNA/network_graphs/", i, "_LBDvsControl_DEGs")
saveToPDF(paste0(path, ".pdf"), width = 5, height = 5)


plot(DEG_graph, vertex.label = ifelse(V(DEG_graph)$selection == "TRUE", V(DEG_graph)$gene_name, NA), vertex.label.color="black", vertex.label.font = 2)

plot(module_graph, vertex.label = ifelse(V(module_graph)$selection == "TRUE", V(module_graph)$gene_name, NA), vertex.label.color="black")
path <- paste0("../../../results/star/WGCNA/network_graphs/", i, "_all_network_LBDvsControl_DEGs_gene_labels")
saveToPDF(paste0(path, ".pdf"), width = 6, height = 6)
remove(graph_DEGs)

# AD vs Control
graph_DEGs <- subset(df_ADvsControl, hcolors == i & FDRq == "< 0.05")
V(module_graph)$selection <- V(module_graph)$gene_name %in% graph_DEGs$gene_name
DEG_graph <- subgraph.edges(graph=module_graph, eids=which(V(module_graph)$selection == "TRUE"), delete.vertices = TRUE)
plot(DEG_graph, edge.arrow.size=.2, edge.color=i,
     vertex.color=i, vertex.frame.color="#ffffff",
     vertex.label=V(DEG_graph)$gene_name, vertex.label.color="black", 
     edge.label=NULL) 
path <- paste0("../../../results/star/WGCNA/network_graphs/", i, "_ADvsControl_DEGs")
saveToPDF(paste0(path, ".pdf"), width = 5, height = 5)
plot(module_graph, vertex.label = ifelse(V(module_graph)$selection == "TRUE", V(module_graph)$gene_name, "."), vertex.label.color="black")
path <- paste0("../../../results/star/WGCNA/network_graphs/", i, "_all_network_ADvsControl_DEGs_gene_labels")
saveToPDF(paste0(path, ".pdf"), width = 6, height = 6)
remove(graph_DEGs)

# LBD vs AD
graph_DEGs <- subset(df_LBDvsAD, hcolors == i & FDRq == "< 0.05")
V(module_graph)$selection <- V(module_graph)$gene_name %in% graph_DEGs$gene_name
DEG_graph <- subgraph.edges(graph=module_graph, eids=which(V(module_graph)$selection == "TRUE"), delete.vertices = TRUE)
plot(DEG_graph, edge.arrow.size=.2, edge.color=i,
     vertex.color=i, vertex.frame.color="#ffffff",
     vertex.label=V(DEG_graph)$gene_name, vertex.label.color="black", 
     edge.label=NULL) 
path <- paste0("../../../results/star/WGCNA/network_graphs/", i, "_LBDvsAD_DEGs")
saveToPDF(paste0(path, ".pdf"), width = 5, height = 5)

plot(module_graph, vertex.label = ifelse(V(module_graph)$selection == "TRUE", V(module_graph)$gene_name, NA), vertex.label.color="black")
path <- paste0("../../../results/star/WGCNA/network_graphs/", i, "_all_network_LBDvsAD_DEGs_gene_labels")
saveToPDF(paste0(path, ".pdf"), width = 6, height = 6)

remove(graph_DEGs, module_graph)
}
```
### magenta subset to only DEGs LBD vs AD 
```{r eval = FALSE}
set.seed(28)
magenta_DEGs <- subset(df_LBDvsControl, hcolors == "magenta" & FDRq == "< 0.05")

V(magenta_graph)$selection <- V(magenta_graph)$gene_name %in% magenta_DEGs$gene_name

magenta_DEG_graph <- subgraph.edges(graph=magenta_graph, eids=which(V(magenta_graph)$selection == "TRUE"), delete.vertices = TRUE)
plot(magenta_DEG_graph)
plot(magenta_DEG_graph, edge.arrow.size=.2, edge.color="magenta",
     vertex.color="magenta", vertex.frame.color="#ffffff",
     vertex.label=V(magenta_DEG_graph)$gene_name, vertex.label.color="black", 
     edge.label=NULL) 
path <- paste0("../../../results/manuscript_figures/magenta_LBDvsControl_DEGs")
saveToPDF(paste0(path, ".pdf"), width = 5, height = 5)

plot(magenta_graph, vertex.label = ifelse(V(magenta_graph)$selection == "TRUE", V(magenta_DEG_graph)$gene_name, NA))

```
### Salmon graph all 
```{r eval = FALSE}
salmon_graph <- wgcna2igraph(net = bwnet, datExpr = counts,
                    modules2plot = c("salmon"),
                    colors2plot = c("salmon"),
                    kME.threshold = 0.75, adjacency.threshold = 0.1,
                    adj.power = 12, verbose = T,
                    node.size = 0, frame.color = NA, node.color = NA,
                    edge.alpha = .5, edge.width =1)
# Add gene names 
salmon_genes <- genes[c(genes$gene_id %in% V(salmon_graph)$name),]
V(salmon_graph)$gene_name <- salmon_genes$gene_name
plot(salmon_graph, edge.arrow.size=.2, edge.color="salmon",
     vertex.color="salmon", vertex.frame.color="#ffffff",
     vertex.label=V(salmon_graph)$gene_name, vertex.label.color="black") 
path <- paste0("../../../results/manuscript_figures/salmon_LBDvsAD_DEGs")
saveToPDF(paste0(path, ".pdf"), width = 4, height = 4)

```
### Salmon subset to only DEGs LBD vs AD 
```{r eval = FALSE}
salmon_DEGs <- subset(df_LBDvsAD, hcolors == "salmon" & FDRq == "< 0.05")

V(salmon_graph)$selection <- V(salmon_graph)$gene_name %in% salmon_DEGs$gene_name

salmon_DEG_graph <- subgraph.edges(graph=salmon_graph, eids=which(V(salmon_graph)$selection == "TRUE"), delete.vertices = TRUE)
plot(salmon_DEG_graph)
plot(salmon_DEG_graph, edge.arrow.size=.2, edge.color="salmon",
     vertex.color="salmon", vertex.frame.color="#ffffff",
     vertex.label=V(salmon_DEG_graph)$gene_name, vertex.label.color="black", 
     edge.label=NULL) 
path <- paste0("../../../results/manuscript_figures/salmon_LBDvsAD_DEGs")
saveToPDF(paste0(path, ".pdf"), width = 4, height = 4)
```
#### green
```{r eval = FALSE}
green_graph <- wgcna2igraph(net = bwnet, datExpr = counts,
                    modules2plot = c("green"),
                    colors2plot = c("green"),
                    kME.threshold = 0.75, adjacency.threshold = 0.1,
                    adj.power = 12, verbose = T,
                    node.size = 0, frame.color = NA, node.color = NA,
                    edge.alpha = .5, edge.width =1)
# Add gene names 
green_genes <- genes[c(genes$gene_id %in% V(green_graph)$name),]
V(green_graph)$gene_name <- green_genes$gene_name
plot(green_graph, edge.arrow.size=.2, edge.color="green",
     vertex.color="green", vertex.frame.color="#ffffff",
     vertex.label=V(green_graph)$gene_name, vertex.label.color="black") 
path <- paste0("../../../results/manuscript_figures/green")
saveToPDF(paste0(path, ".pdf"), width = 12, height = 12)

green_DEGs <- subset(df_LBDvsAD, hcolors == "green" & FDRq == "< 0.05")
V(green_graph)$selection <- V(green_graph)$gene_name %in% green_DEGs$gene_name
green_DEG_graph <- subgraph.edges(graph=green_graph, eids=which(V(green_graph)$selection == "TRUE"), delete.vertices = TRUE)
plot(green_DEG_graph)
plot(green_DEG_graph, edge.arrow.size=.2, edge.color="green",
     vertex.color="green", vertex.frame.color="#ffffff",
     vertex.label=V(gbf)$gene_name, vertex.label.color="black", 
     edge.label=NULL) 
path <- paste0("../../../results/manuscript_figures/green_LBDvsAD_DEGs")
saveToPDF(paste0(path, ".pdf"), width = 4, height = 4)
```
# Metascape 
#### salmon
```{r}
salmon_enrich_results <-
  read_excel(
    paste0(
      "../../../results/",
      tool,
      "/WGCNA/metascape/salmon/metascape_result.xlsx"
    ), sheet = 2
  )


GO_ID_salmon <- c(
    "1_Summary",
    "2_Summary",
    "3_Summary",
    "4_Summary",
    "5_Summary",
    "7_Summary",
    "8_Summary",
    "9_Summary"
)

salmon_enrich_results_subset <-
  salmon_enrich_results[salmon_enrich_results$GroupID %in% GO_ID_salmon, ]
salmon_enrich_results_subset$Cluster <- c("salmon")
salmon_enrich_results_subset$Description <-
  factor(salmon_enrich_results_subset$Description,
         levels = salmon_enrich_results_subset$Description)
salmon_enrich_results_subset$Description <-
  fct_rev(salmon_enrich_results_subset$Description)

salmon_gene_count <-
  strsplit(as.character(salmon_enrich_results_subset$InTerm_InList),
           "/",
           fixed = T)
salmon_gene_count_df <-
  data.frame(matrix(
    unlist(salmon_gene_count),
    nrow = length(salmon_gene_count),
    byrow = TRUE
  ), stringsAsFactors = FALSE)
salmon_enrich_results_subset$InTerm <- as.numeric(salmon_gene_count_df$X1)

#salmon_enrich_results_subset$Description <-
#      gsub(
#       "regulation of cysteine-type endopeptidase activity involved in apoptotic process",
#        "reg cysteine endopeptidase in apoptotic process",
#        salmon_enrich_results_subset$Description
#      )
#
salmon_enrich_results_subset$Description <- factor(salmon_enrich_results_subset$Description, levels = salmon_enrich_results_subset$Description)

salmon_enrich_results_subset$Description <- factor(salmon_enrich_results_subset$Description, levels = rev(salmon_enrich_results_subset$Description))

salmon_enrich_plot <-
  ggplot(data = salmon_enrich_results_subset, aes(x = InTerm, y = Description)) +
  ggplot2::facet_grid(~ Cluster) +
  geom_bar(stat = "identity", aes(fill = LogP), width = .65) +
  theme_bw() +
  scale_y_discrete(expand = c(0, 0)) +
  labs(x = "Gene count", y = "") +
  guides(fill = guide_legend(
    size = 1,
    title = expression(log[10] ~ "(" ~ italic("p") ~ "-value)")
  )) +
  scale_fill_gradientn(
    colours = c("#800026FF","#FD8D3CFF", "salmon"), # midnightblue
    guide = "legend",
    limits = c(-100,-2)
  ) + theme(strip.text = element_text(size = 8), 
          axis.text.y = element_text(size = 8),
          axis.text.x = element_text(size = 8), 
          axis.title.x = element_text(size = 8), 
          legend.text = element_text(size = 8),
          legend.title = element_text(size = 8),
          legend.margin=margin(0,0.5,0,0),
          legend.box.margin=margin(-10,-2,-10,-7.5), 
          plot.margin = margin(0.1, 0.3, 0.3, 0.1, "cm"), 
          plot.title = element_text(size = 8, hjust = -2.25, vjust=0, margin = margin(0,0,0,0)))
 #  theme(aspect.ratio = 1.5) # to reduce the size of the boxes 
addSmallLegend <- function(salmon_enrich_plot, pointSize = 3, textSize = 6, spaceLegend = .5) {
    salmon_enrich_plot +
        guides(shape = guide_legend(override.aes = list(size = pointSize)),
               color = guide_legend(override.aes = list(size = pointSize))) +
        theme(legend.title = element_text(size = textSize), 
              legend.position = "right",
              legend.text  = element_text(size = textSize),
              legend.key.size = unit(spaceLegend, "lines"))
}
# Apply on original plot
salmon_enrich_plot <- addSmallLegend(salmon_enrich_plot)
salmon_enrich_plot
```

#### green
```{r}
green_enrich_results <-
  read_excel(
    paste0(
      "../../../results/",
      tool,
      "/WGCNA/metascape/green/metascape_result.xlsx"
    ), sheet = 2
  )


GO_ID_green <- c(
    "1_Summary",
    "2_Summary",
    "3_Summary",
    "4_Summary",
    "5_Summary",
    "6_Summary",
    "7_Summary",
    "8_Summary",
    "9_Summary",
    "11_Summary"

)

green_enrich_results_subset <-
  green_enrich_results[green_enrich_results$GroupID %in% GO_ID_green, ]
green_enrich_results_subset$Cluster <- c("green")
green_enrich_results_subset$Description <-
  factor(green_enrich_results_subset$Description,
         levels = green_enrich_results_subset$Description)
green_enrich_results_subset$Description <-
  fct_rev(green_enrich_results_subset$Description)

green_gene_count <-
  strsplit(as.character(green_enrich_results_subset$InTerm_InList),
           "/",
           fixed = T)
green_gene_count_df <-
  data.frame(matrix(
    unlist(green_gene_count),
    nrow = length(green_gene_count),
    byrow = TRUE
  ), stringsAsFactors = FALSE)
green_enrich_results_subset$InTerm <- as.numeric(green_gene_count_df$X1)

green_enrich_results_subset$Description <-
      gsub(
       "Diseases of signal transduction by growth factor receptors and second messenger",
        "Diseases of signal transduction",
        green_enrich_results_subset$Description
      )

green_enrich_results_subset$Description <-
      gsub(
       "Epithelial to mesenchymal transition in colorectal cancer",
        "Epithelial to mesenchymal transition",
        green_enrich_results_subset$Description
      )
green_enrich_results_subset$Description <- factor(green_enrich_results_subset$Description, levels = green_enrich_results_subset$Description)

green_enrich_results_subset$Description <- factor(green_enrich_results_subset$Description, levels = rev(green_enrich_results_subset$Description))

green_enrich_plot <-
  ggplot(data = green_enrich_results_subset, aes(x = InTerm, y = Description)) +
  ggplot2::facet_grid(~ Cluster) +
  geom_bar(stat = "identity", aes(fill = LogP), width = .65) +
  theme_bw() +
  scale_y_discrete(expand = c(0, 0)) +
  labs(x = "Gene count", y = "") +
  guides(fill = guide_legend(
    size = 1,
    title = expression(log[10] ~ "(" ~ italic("p") ~ "-value)")
  )) +
  scale_fill_gradientn(
    colours = c("green", "gray"), # midnightblue
    guide = "legend",
    limits = c(-9,-1)
  ) + theme(strip.text = element_text(size = 8), 
          axis.text.y = element_text(size = 8),
          axis.text.x = element_text(size = 8), 
          axis.title.x = element_text(size = 8), 
          legend.text = element_text(size = 8),
          legend.title = element_text(size = 8),
          legend.margin=margin(0,0.5,0,0),
          legend.box.margin=margin(-10,-2,-10,-7.5), 
          plot.margin = margin(0.1, 0.3, 0.3, 0.1, "cm"), 
          plot.title = element_text(size = 8, hjust = -2.25, vjust=0, margin = margin(0,0,0,0)))
 #  theme(aspect.ratio = 1.5) # to reduce the size of the boxes 
addSmallLegend <- function(green_enrich_plot, pointSize = 3, textSize = 6, spaceLegend = .5) {
    green_enrich_plot +
        guides(shape = guide_legend(override.aes = list(size = pointSize)),
               color = guide_legend(override.aes = list(size = pointSize))) +
        theme(legend.title = element_text(size = textSize), 
              legend.position = "right",
              legend.text  = element_text(size = textSize),
              legend.key.size = unit(spaceLegend, "lines"))
}
# Apply on original plot
green_enrich_plot <- addSmallLegend(green_enrich_plot)
green_enrich_plot
```
#### magenta
```{r}
magenta_enrich_results <-
  read_excel(
    paste0(
      "../../../results/",
      tool,
      "/WGCNA/metascape/magenta/metascape_result.xlsx"
    ), sheet = 2
  )


GO_ID_magenta <- c(
    "1_Summary",
    "2_Summary",
    "3_Summary",
    "4_Summary",
    "5_Summary",
    "6_Summary",
    "7_Summary",
    "8_Summary",
    "9_Summary",
    "10_Summary",
    "11_Summary",
    "12_Summary",
    "13_Summary",
    "14_Summary",
    "15_Summary",
    "16_Summary",
    "17_Summary",
    "18_Summary",
    "19_Summary",    
    "20_Summary"

)

magenta_enrich_results_subset <-
  magenta_enrich_results[magenta_enrich_results$GroupID %in% GO_ID_magenta, ]
magenta_enrich_results_subset$Cluster <- c("magenta")
magenta_enrich_results_subset$Description <-
  factor(magenta_enrich_results_subset$Description,
         levels = magenta_enrich_results_subset$Description)
magenta_enrich_results_subset$Description <-
  fct_rev(magenta_enrich_results_subset$Description)

magenta_gene_count <-
  strsplit(as.character(magenta_enrich_results_subset$InTerm_InList),
           "/",
           fixed = T)
magenta_gene_count_df <-
  data.frame(matrix(
    unlist(magenta_gene_count),
    nrow = length(magenta_gene_count),
    byrow = TRUE
  ), stringsAsFactors = FALSE)
magenta_enrich_results_subset$InTerm <- as.numeric(magenta_gene_count_df$X1)

magenta_enrich_results_subset$Description <-
      gsub(
       "regulation of myeloid leukocyte mediated immunity",
        "reg of myeloid leukocyte mediated immunity",
        magenta_enrich_results_subset$Description
      )

magenta_enrich_results_subset$Description <- factor(magenta_enrich_results_subset$Description, levels = magenta_enrich_results_subset$Description)

magenta_enrich_results_subset$Description <- factor(magenta_enrich_results_subset$Description, levels = rev(magenta_enrich_results_subset$Description))

magenta_enrich_plot <-
  ggplot(data = magenta_enrich_results_subset, aes(x = InTerm, y = Description)) +
  ggplot2::facet_grid(~ Cluster) +
  geom_bar(stat = "identity", aes(fill = LogP), width = .65) +
  theme_bw() +
  scale_y_discrete(expand = c(0, 0)) +
  labs(x = "Gene count", y = "") +
  guides(fill = guide_legend(
    size = 1,
    title = expression(log[10] ~ "(" ~ italic("p") ~ "-value)")
  )) +
  scale_fill_gradientn(
    colours = c("magenta", "gray"), # midnightblue
    guide = "legend",
    limits = c(-32,-2)
  ) + theme(strip.text = element_text(size = 8), 
          axis.text.y = element_text(size = 8),
          axis.text.x = element_text(size = 8), 
          axis.title.x = element_text(size = 8), 
          legend.text = element_text(size = 8),
          legend.title = element_text(size = 8),
          legend.margin=margin(0,0.5,0,0),
          legend.box.margin=margin(-10,-2,-10,-7.5), 
          plot.margin = margin(0.1, 0.3, 0.3, 0.1, "cm"), 
          plot.title = element_text(size = 8, hjust = -2.25, vjust=0, margin = margin(0,0,0,0)))
 #  theme(aspect.ratio = 1.5) # to reduce the size of the boxes 
addSmallLegend <- function(magenta_enrich_plot, pointSize = 3, textSize = 6, spaceLegend = .5) {
    magenta_enrich_plot +
        guides(shape = guide_legend(override.aes = list(size = pointSize)),
               color = guide_legend(override.aes = list(size = pointSize))) +
        theme(legend.title = element_text(size = textSize), 
              legend.position = "right",
              legend.text  = element_text(size = textSize),
              legend.key.size = unit(spaceLegend, "lines"))
}
# Apply on original plot
magenta_enrich_plot <- addSmallLegend(magenta_enrich_plot)
magenta_enrich_plot
```
#### brown
```{r}
brown_enrich_results <-
  read_excel(
    paste0(
      "../../../results/",
      tool,
      "/WGCNA/metascape/brown/metascape_result.xlsx"
    ), sheet = 2
  )


GO_ID_brown <- c(
    "1_Summary",
    "2_Summary",
    "3_Summary",
    "4_Summary",
    "5_Summary",
    "6_Summary",
    "7_Summary",
    "8_Summary",
    "9_Summary",
    "10_Summary",
    "11_Summary",
    "12_Summary",
    "13_Summary",
    "14_Summary",
    "15_Summary",
    "16_Summary",
    "17_Summary",
    "18_Summary",
    "19_Summary",    
    "20_Summary"

)

brown_enrich_results_subset <-
  brown_enrich_results[brown_enrich_results$GroupID %in% GO_ID_brown, ]
brown_enrich_results_subset$Cluster <- c("brown")
brown_enrich_results_subset$Description <-
  factor(brown_enrich_results_subset$Description,
         levels = brown_enrich_results_subset$Description)
brown_enrich_results_subset$Description <-
  fct_rev(brown_enrich_results_subset$Description)

brown_gene_count <-
  strsplit(as.character(brown_enrich_results_subset$InTerm_InList),
           "/",
           fixed = T)
brown_gene_count_df <-
  data.frame(matrix(
    unlist(brown_gene_count),
    nrow = length(brown_gene_count),
    byrow = TRUE
  ), stringsAsFactors = FALSE)
brown_enrich_results_subset$InTerm <- as.numeric(brown_gene_count_df$X1)

brown_enrich_results_subset$Description <-
      gsub(
       "regulation of plasma membrane bounded cell projection organization",
        "reg of plasma membrane bounded cell projection",
        brown_enrich_results_subset$Description
      )

brown_enrich_results_subset$Description <- factor(brown_enrich_results_subset$Description, levels = brown_enrich_results_subset$Description)

brown_enrich_results_subset$Description <- factor(brown_enrich_results_subset$Description, levels = rev(brown_enrich_results_subset$Description))

brown_enrich_plot <-
  ggplot(data = brown_enrich_results_subset, aes(x = InTerm, y = Description)) +
  ggplot2::facet_grid(~ Cluster) +
  geom_bar(stat = "identity", aes(fill = LogP), width = .65) +
  theme_bw() +
  scale_y_discrete(expand = c(0, 0)) +
  labs(x = "Gene count", y = "") +
  guides(fill = guide_legend(
    size = 1,
    title = expression(log[10] ~ "(" ~ italic("p") ~ "-value)")
  )) +
  scale_fill_gradientn(
    colours = c("brown", "gray"), # midnightblue
    guide = "legend",
    limits = c(-30,-5)
  ) + theme(strip.text = element_text(size = 8), 
          axis.text.y = element_text(size = 8),
          axis.text.x = element_text(size = 8), 
          axis.title.x = element_text(size = 8), 
          legend.text = element_text(size = 8),
          legend.title = element_text(size = 8),
          legend.margin=margin(0,0.5,0,0),
          legend.box.margin=margin(-10,-2,-10,-7.5), 
          plot.margin = margin(0.1, 0.3, 0.3, 0.1, "cm"), 
          plot.title = element_text(size = 8, hjust = -2.25, vjust=0, margin = margin(0,0,0,0)))
 #  theme(aspect.ratio = 1.5) # to reduce the size of the boxes 
addSmallLegend <- function(brown_enrich_plot, pointSize = 3, textSize = 6, spaceLegend = .5) {
    brown_enrich_plot +
        guides(shape = guide_legend(override.aes = list(size = pointSize)),
               color = guide_legend(override.aes = list(size = pointSize))) +
        theme(legend.title = element_text(size = textSize), 
              legend.position = "right",
              legend.text  = element_text(size = textSize),
              legend.key.size = unit(spaceLegend, "lines"))
}
# Apply on original plot
brown_enrich_plot <- addSmallLegend(brown_enrich_plot)
brown_enrich_plot
```



# Gene-level combine all plots 
```{r}
row1 <- ggarrange(NULL,
    gene_count_per_module,
    ncol = 2,
    labels = c("a)", "b)"),
    font.label = list(size = 8)
  )

row2.a <- ggarrange(
  ME_heatmap,
  ncol = 1,
  labels = c("c)"),
  font.label = list(size = 8)
  )

row2.b <- ggarrange(
  salmon_enrich_plot,
  green_enrich_plot,
  nrow = 2,
  labels = c("d)", "e)"),
  heights = c(.9, 1),
  font.label = list(size = 8)
  )


row2 <- ggarrange(
  row2.a, 
  row2.b, 
  ncol = 2)
  
  
row3 <-
  ggarrange(brown_enrich_plot,
    magenta_enrich_plot,
    ncol = 2,
    labels = c("f)", "g)"),
    font.label = list(size = 8)
  )

combind <-
  ggarrange(
    row1,
    row2,
    row3,
    nrow = 3,
    heights = c(.8,1,.8)
  )
combind

path <- paste0("../../../results/manuscript_figures/Figure_4_WGCNA_TYPE")
saveToPDF(paste0(path, ".pdf"), width = 8.5, height = 11)
```
```{r}
```
