---
title: "Figure 1. LBD vs Control"
author: "Kimberly Olney"
date: "07/10/2023"
output: html_document
---
# Setup
```{r setup}
knitr::opts_knit$set(root.dir = ".")
```
# User defined variables
```{r set_variables}
source(here::here("scripts/R", "file_paths_and_colours.R"))
library("readxl")
#source(here::here("scripts/R", "gtf_path.R"))
condition <- c("TYPE")
tool <- "star"
dge.filtered.norm <- readRDS(paste0("../../../rObjects/dge.filtered.norm.rds"))
```
# Binary list shared gene function
```{r}
fromList <- function (input) {
  # Same as original fromList()...
  elements <- unique(unlist(input))
  data <- unlist(lapply(input, function(x) {
      x <- as.vector(match(elements, x))
      }))
  data[is.na(data)] <- as.integer(0)
  data[data != 0] <- as.integer(1)
  data <- data.frame(matrix(data, ncol = length(input), byrow = F))
  data <- data[which(rowSums(data) != 0), ]
  names(data) <- names(input)
  # ... Except now it conserves your original value names!
  row.names(data) <- elements
  return(data)
  }
```
# Volcano 
```{r volcano, warning=FALSE}
i <- "LBDvsControl"
group1_vs_group2 <- read.delim(paste0("../../../results/", tool, "/DEGs/", condition, "_", i, "_gene_DEGs_FDRq1.00.txt"))
  color_values <- vector()
  max <- nrow(group1_vs_group2)
  for (row in 1:max) {
    if (group1_vs_group2$adj.P.Val[row] < 0.05) {
      if (group1_vs_group2$logFC [row] > 0.25) {
        color_values <- c(color_values, 1)
      }
      else if (group1_vs_group2$logFC[row] < -.25) {
        color_values <- c(color_values, 2)
      }
      else {
        color_values <- c(color_values, 3)
      }
    }
    else{
      color_values <- c(color_values, 3)
    }
  }
  group1_vs_group2$color_adjpval_0.05 <- factor(color_values)
  data <- group1_vs_group2
  # plot only if there are DEGs with p_val_adj < 0.05
  num <- subset(data, (adj.P.Val < 0.05 & logFC < -.25)  | (adj.P.Val < 0.05 & logFC > .25 ))
  num <- nrow(num)
    up <- data[data$color_adjpval_0.05 == 1,]
    up10 <- up[1:10,]
    upFold <- subset(up, logFC > 1)
    upFold <- upFold[!(upFold$gene_name %in% up10$gene_name),]
    down <- data[data$color_adjpval_0.05 == 2,]
    down10 <- down[1:10,]
    downFold <- subset(down, logFC < -1)
    downFold <- downFold[!(downFold$gene_name %in% down10$gene_name),]
    if (!1 %in% unique(data$color_adjpval_0.05)) {
      my_colors <- c("blue", "gray")
    } else if (!2 %in% unique(data$color_adjpval_0.05)) {
      my_colors <- c("red", "gray")
    } else if (!1 %in% unique(data$color_adjpval_0.05) &&
               !2 %in% unique(data$color_adjpval_0.05)) {
      my_colors <- c("gray")
    } else {
      my_colors <- c("red", "blue", "gray")
    }
    hadjpval <- (-log10(max(data$P.Value[data$adj.P.Val < 0.05],
                            na.rm = TRUE)))
    negFC <- c(-.25)
    posFC <- c(.25) 
    i <- gsub("vs", " vs ", i)
volcano <- ggplot(data = data,
             aes(
               x = logFC,
               y = -log10(P.Value),
               color = color_adjpval_0.05
             )) +
      geom_point(alpha = 0.8, size = 1) +
      theme_bw() +
      theme(legend.position = "none", 
        plot.margin = margin(0.1, 0.2, 0, 0.2, "cm"), 
             axis.title.x = element_text(size = 6),
             axis.text.x = element_text(size = 6),
             axis.title.y = element_text(size = 6),
             axis.text.y = element_text(size = 6),
             plot.title = element_text(size = 7, margin = margin(0,0,0,0))) +
      scale_color_manual(values = my_colors) +
      labs(
        title = "",
        x = expression(log[2](FC)),
        y = expression(-log[10] ~ "(" ~ italic("p") ~ "-value)")
      )  +
      #theme(plot.title = element_text(size = 7, margin = margin(0,0,0,0))) +
      geom_hline(yintercept = hadjpval,
                 #  horizontal line
                 colour = "#000000",
                 linetype = "dashed") +
      geom_vline(xintercept = negFC,
                 #  horizontal line
                 colour = "#000000",
                 linetype = "dashed") +
      geom_vline(xintercept = posFC,
                 #  horizontal line
                 colour = "#000000",
                 linetype = "dashed") +
      ggtitle(paste0(i)) +
      geom_text_repel(
        data = up10,
        aes(
          x = logFC,
          y = -log10(P.Value),
          label = gene_name
        ),
        color = "maroon",
        fontface = "italic",
        size = 2,
        max.overlaps = getOption("ggrepel.max.overlaps", default = 10)
      ) +
      geom_text_repel(
        data = upFold,
        aes(
          x = logFC,
          y = -log10(P.Value),
          label = gene_name
        ),
        color = "maroon",
        fontface = "italic",
        size = 2,
        max.overlaps = getOption("ggrepel.max.overlaps", default = 10)
      ) +
      geom_text_repel(
        data = down10,
        aes(
          x = logFC,
          y = -log10(P.Value),
          label = gene_name
        ),
        color = "navyblue",
        fontface = "italic",
        size = 2,
        max.overlaps = getOption("ggrepel.max.overlaps", default = 10)
      ) +
      geom_text_repel(
        data = downFold,
        aes(
          x = logFC,
          y = -log10(P.Value),
          label = gene_name
        ),
        color = "navyblue",
        fontface = "italic",
        size = 2,
        max.overlaps = getOption("ggrepel.max.overlaps", default = 10)
      )
volcano
i <- gsub(" vs ", "vs", i)
```

```{r cleanup}
# clean up
remove(up, up10, upFold, downFold, down10, data)
```

# Metascape Enrichment 
### Format the dataframes
```{r}
# read in enrichment analysis results
up_enrich_results <-
  read_excel(
    paste0(
      "../../../results/",
      tool,
      "/metascape/",
      i,
      "_up_regulated/metascape_result.xlsx"
    ), sheet = 2
  )

down_enrich_results <-
  read_excel(
    paste0(
      "../../../results/",
      tool,
      "/metascape/",
      i,
      "_down_regulated/metascape_result.xlsx"
    ), sheet = 2
  )
```

```{r}
# select the GO term IDs we want to show in the plot
GO_ID_up <-
  c(
    "1_Summary",
    "2_Summary",
    "3_Summary",
    "4_Summary",
    "5_Summary",
    "6_Summary",
    "7_Summary",
    "8_Summary",
    "9_Summary", 
    "10_Summary",
    "11_Summary",
    "12_Summary",
    "13_Summary",
    "14_Summary",
    "15_Summary",
    "16_Summary",
    "17_Summary",
    "18_Summary",
    "19_Summary", 
    "20_Summary"
  )
GO_ID_down <- c(
    "1_Summary",
    "2_Summary",
    "3_Summary",
    "4_Summary",
    "5_Summary",
    "6_Summary",
    "7_Summary",
    "8_Summary",
    "9_Summary", 
    "10_Summary",
    "11_Summary",
    "12_Summary",
    "13_Summary",
    "14_Summary",
    "15_Summary",
    "16_Summary",
    "17_Summary",
    "18_Summary",
    "19_Summary", 
    "20_Summary"
)

up_enrich_results_subset <-
  up_enrich_results[up_enrich_results$GroupID %in% GO_ID_up, ]
up_enrich_results_subset$Cluster <- c("up-regulated")
up_enrich_results_subset$Description <-
  factor(up_enrich_results_subset$Description,
         levels = up_enrich_results_subset$Description)
up_enrich_results_subset$Description <-
  fct_rev(up_enrich_results_subset$Description)

down_enrich_results_subset <-
  down_enrich_results[down_enrich_results$GroupID %in% GO_ID_down, ]
down_enrich_results_subset$Cluster <- c("down-regulated")
down_enrich_results_subset$Description <-
  factor(down_enrich_results_subset$Description,
         levels = down_enrich_results_subset$Description)
down_enrich_results_subset$Description <-
  fct_rev(down_enrich_results_subset$Description)

# get the number of genes in each summary
up_gene_count <-
  strsplit(as.character(up_enrich_results_subset$InTerm_InList),
           "/",
           fixed = T)
up_gene_count_df <-
  data.frame(matrix(
    unlist(up_gene_count),
    nrow = length(up_gene_count),
    byrow = TRUE
  ), stringsAsFactors = FALSE)
up_enrich_results_subset$InTerm <- as.numeric(up_gene_count_df$X1)

down_gene_count <-
  strsplit(as.character(down_enrich_results_subset$InTerm_InList),
           "/",
           fixed = T)
down_gene_count_df <-
  data.frame(matrix(
    unlist(down_gene_count),
    nrow = length(down_gene_count),
    byrow = TRUE
  ), stringsAsFactors = FALSE)
down_enrich_results_subset$InTerm <- as.numeric(down_gene_count_df$X1)

# combine together
up_and_down_enrich_results_subset <- rbind(up_enrich_results_subset, down_enrich_results_subset)

up_and_down_enrich_results_subset$Description <-
      gsub(
       "regulation of tumor necrosis factor superfamily cytokine production",
        "reg tumor necrosis factor cytokine production",
        up_and_down_enrich_results_subset$Description
      )
up_and_down_enrich_results_subset$Description <-
      gsub(
       "Trans-sulfuration, one-carbon metabolism and related pathways",
        "Trans-sulfuration, one-carbon metabolism & pathways",
        up_and_down_enrich_results_subset$Description
      )
up_and_down_enrich_results_subset$Description <-
  factor(up_and_down_enrich_results_subset$Description, levels = unique(up_and_down_enrich_results_subset$Description)) 

up_and_down_enrich_results_subset$Description <- fct_rev(up_and_down_enrich_results_subset$Description)
up_and_down_enrich_results_subset$Cluster <- factor(up_and_down_enrich_results_subset$Cluster, levels = c("up-regulated", "down-regulated"))
```

remove files 
```{r}
remove(
  down_enrich_results,
  down_gene_count,
  down_gene_count_df,
  up_enrich_results,
  up_gene_count,
  up_gene_count_df
)
```

### Enrichment plot showing log10 P-value
#### down and up seperately
```{r}
down_enrich_plot <-
  ggplot(data = down_enrich_results_subset, aes(x = InTerm, y = Description)) +
  ggplot2::facet_grid(~ Cluster) +
  geom_bar(stat = "identity", aes(fill = LogP), width = .65) +
  theme(axis.title.x = element_text(size = 6),
        axis.text.x = element_text(size = 6)) +
  theme(axis.title.y = element_text(size = 6),
        axis.text.y = element_text(size = 6)) +
  theme_bw() +
  scale_y_discrete(expand = c(0, 0)) +
  labs(x = "Gene count", y = "") +
  #  xlim(0,60) +
  guides(fill = guide_legend(
    size = 1,
    title = expression(log[10] ~ "(" ~ italic("p") ~ "-value)")
  )) +
  scale_fill_gradientn(
    colours = c("darkmagenta", "mediumorchid3", "grey"),
    guide = "legend",
    limits = c(-45,-1),
  ) +
  theme(legend.text = element_text(size = 6),
        legend.title = element_text(size = 6))
down_enrich_plot

up_enrich_plot <-
  ggplot(data = up_enrich_results_subset, aes(x = InTerm, y = Description)) +
  ggplot2::facet_grid(~ Cluster) +
  geom_bar(stat = "identity", aes(fill = LogP), width = .65) +
  theme(axis.title.x = element_text(size = 6),
        axis.text.x = element_text(size = 6)) +
  theme(axis.title.y = element_text(size = 6),
        axis.text.y = element_text(size = 6)) +
  theme(legend.text = element_text(size = 6)) +
  theme_bw() +
  scale_y_discrete(expand = c(0, 0)) +
  labs(x = "Gene count", y = "") +
  guides(fill = guide_legend(title = expression(log[10] ~ "(" ~ italic("p") ~ "-value)"))) +
  scale_fill_gradientn(
    colours = c("darkmagenta", "mediumorchid3", "grey"),
    guide = "legend",
    limits = c(-45,-1)
  ) +
  theme(legend.text = element_text(size = 6),
        legend.title = element_text(size = 6))
up_enrich_plot
```

#### down and up together
```{r}
up_and_down_enrich_plot <-
  ggplot(data = up_and_down_enrich_results_subset, aes(x = InTerm, y = Description)) +
  ggplot2::facet_wrap(~ Cluster, scales = "free_y", ncol = 1) +
  geom_bar(stat = "identity", aes(fill = LogP), width = .7, position = position_dodge(width = .2)) +
  theme_bw() +
  ggtitle(paste0("Enrichment summaries")) +
  labs(x = "Gene count", y = NULL) +
  guides(fill = guide_legend(title = expression(log[10] ~ "(" ~ italic("p") ~ "-value)"))) +
  scale_fill_gradientn(
    colours = c("darkmagenta", "mediumorchid3", "grey"),
    guide = "legend",
    limits = c(-45,-1)
  ) +
    theme(strip.text = element_text(size = 6), 
          axis.text.y = element_text(size = 7),
          axis.text.x = element_text(size = 6), 
          axis.title.x = element_text(size = 6), 
          legend.text = element_text(size = 6),
          legend.title = element_text(size = 6),
          legend.margin=margin(0,0.5,0,0),
          legend.box.margin=margin(-10,-2,-10,-7.5), 
          plot.margin = margin(0.1, 0.2, 0, 0.2, "cm"), 
          plot.title = element_text(size = 7, hjust = -2.25, vjust=0, margin = margin(0,0,0,0)))
 #  theme(aspect.ratio = 1.5) # to reduce the size of the boxes 
addSmallLegend <- function(myPlot, pointSize = 3, textSize = 6, spaceLegend = .5) {
    myPlot +
        guides(shape = guide_legend(override.aes = list(size = pointSize)),
               color = guide_legend(override.aes = list(size = pointSize))) +
        theme(legend.title = element_text(size = textSize), 
              legend.text  = element_text(size = textSize),
              legend.key.size = unit(spaceLegend, "lines"))
}

# Apply on original plot
up_and_down_enrich_plot <- addSmallLegend(up_and_down_enrich_plot)
```

```{r}
remove(up_enrich_plot,
       down_enrich_plot)
```
## Combine volcano and enrichment plot 
```{r}
row1 <- ggarrange(
  volcano,
  up_and_down_enrich_plot,
  ncol = 2,
  labels = c("a)", "b)"),
  widths = c(2.5, 4.25), 
  font.label = list(size = 8)
  )
row1

path <- paste0("../../../results/manuscript_figures/Figure_1_row1_test")
saveToPDF(paste0(path, ".pdf"), width = 8, height = 4.25)
```
# logCPM data
```{r}
lcpm <- edgeR::cpm(dge.filtered.norm$counts, log = TRUE) # obtain log CPM counts
```
### Fold change information 
```{r}
up_regulated <- subset(
  group1_vs_group2, adj.P.Val <= 0.05 & logFC > 0.25)
down_regulated <- subset(
  group1_vs_group2, adj.P.Val <= 0.05 & logFC <= -0.25)

# sort by log2FC and adjusted p-val
up_regulated_sort <-
  up_regulated[order(-up_regulated$logFC, up_regulated$adj.P.Val), ]
down_regulated_sort <-
  down_regulated[order(down_regulated$logFC, down_regulated$adj.P.Val), ]

# get the lcpm information for those genes of interest
up_regulated_lcpm <- subset(lcpm, row.names(lcpm) %in%
                              up_regulated_sort$gene_id)
down_regulated_lcpm <- subset(lcpm, row.names(lcpm) %in%
                                down_regulated_sort$gene_id)

# reform the data
up_regulated_sort_df <- up_regulated_lcpm %>%
  as.data.frame() %>%
  rownames_to_column("gene_id") %>%
  pivot_longer(-c(gene_id), names_to = "samples", values_to = "counts") %>%
  mutate(samples = fct_relevel(samples, colnames(up_regulated_lcpm)))

down_regulated_sort_df <- down_regulated_lcpm %>%
  as.data.frame() %>%
  rownames_to_column("gene_id") %>%
  pivot_longer(-c(gene_id), names_to = "samples", values_to = "counts") %>%
  mutate(samples = fct_relevel(samples, colnames(down_regulated_lcpm)))

# lock in gene order which is sorted by adjusted p-value
up_regulated_sort_df$gene_id <-
  factor(up_regulated_sort_df$gene_id,
         levels = unique(up_regulated_sort_df$gene_id))
up_regulated_sort_df$gene_id <-
  fct_rev(up_regulated_sort_df$gene_id)
down_regulated_sort_df$gene_id <-
  factor(down_regulated_sort_df$gene_id,
         levels = unique(down_regulated_sort_df$gene_id))
down_regulated_sort_df$gene_id <-
  fct_rev(down_regulated_sort_df$gene_id)

DEG_df <-
  rbind(up_regulated_sort_df, down_regulated_sort_df)
colnames(DEG_df)[2] ="NPID"
NPID_TYPE <- metadata[,c("NPID","TYPE")]
DEG_df <- merge(DEG_df, NPID_TYPE, by = "NPID")
DEG_sort <- rbind(up_regulated_sort, down_regulated_sort)
names(DEG_sort)[names(DEG_sort) == "gene_name"] <- "gene"
```
remove files
```{r}
remove(
  lcpm,
  down_regulated,
  down_regulated_lcpm,
  down_regulated_sort,
  down_regulated_sort_df,
  up_regulated,
  up_regulated_lcpm,
  up_regulated_sort,
  up_regulated_sort_df
)
```

### Genes in enrichment plot all
## up 
```{r}
up_enrich_df <-
  up_enrich_results_subset[, c("GroupID", "Term", "Description", "Symbols")]
up_enrich_df_melt <- reshape2::melt(up_enrich_df)
up_gene_char <- str_split(up_enrich_df_melt$Symbols, ",")

up_gene_pathway <-
  data.frame(gene = unlist(up_gene_char),
             value = rep(up_enrich_df$Description, lengths(up_gene_char)))
up_pathways_levels <- levels(fct_rev(up_gene_pathway$value))

up_enrich_df <-
  up_enrich_results_subset[, c("GroupID", "Term", "Description", "Symbols")]
up_enrich_df_melt <- reshape2::melt(up_enrich_df)
up_gene_char <- str_split(up_enrich_df_melt$Symbols, ",")

up_gene_pathway <-
  data.frame(gene = unlist(up_gene_char),
             value = rep(up_enrich_df$Description, lengths(up_gene_char)))
up_pathways_levels <- levels(fct_rev(up_gene_pathway$value))

up_input <- list(
  "positive regulation of immune response" = up_gene_char[[1]],
  "tube morphogenesis" = up_gene_char[[7]],
  "Hemostasis" = up_gene_char[[10]],
  "positive regulation of cell migration" = up_gene_char[[11]],
  "TYROBP causal network in microglia" = up_gene_char[[19]]
)
up_data <- fromList(up_input)
up_data$gene <- row.names(up_data)
up_data_melt <- reshape2::melt(up_data)

# get the fold change value for those genes
up_df <- merge(up_data_melt, DEG_sort,
                       by = "gene")
# order by fold change 
up_df <- up_df[order(-up_df$logFC, -up_df$adj.P.Val),]

# subset by GO termes
GO_1 <-
  subset(up_df,
         up_df$variable == "positive regulation of immune response"
         & up_df$value == 1)
GO_1_top <- head(GO_1, 15)

GO_2 <-
  subset(up_df,
         up_df$variable == "tube morphogenesis"
         & up_df$value == 1)
GO_2_top <- head(GO_2, 15)

GO_3 <-
  subset(up_df,
         up_df$variable == "Hemostasis"
         & up_df$value == 1)
GO_3_top <- head(GO_3, 15)

GO_4 <-
  subset(up_df,
         up_df$variable == "positive regulation of cell migration"
         & up_df$value == 1)
GO_4_top <- head(GO_4, 15)

GO_5 <-
  subset(up_df,
         up_df$variable == "TYROBP causal network in microglia"
         & up_df$value == 1)
GO_5_top <- head(GO_5, 15)
```

#### plot lcpm GO heatmaps
##### up 1
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_1_DEG <- subset(DEG_df, DEG_df$gene_id %in% GO_1_top$gene_id)
GO_1_DEG <- merge(GO_1_DEG, GO_1_top, by = "gene_id", all.x = TRUE)
GO_1_DEG$gene <- factor(GO_1_DEG$gene,
                          levels = unique(GO_1_top$gene))
GO_1_DEG$gene <- fct_rev(GO_1_DEG$gene)
GO_1_DEG <- subset(GO_1_DEG, TYPE == "CONTROL" | TYPE ==  "LBD")
# heatmap 
up_GO_1_lcpm_heat <- ggplot(data = GO_1_DEG, aes(x = reorder(NPID, -counts), y = gene)) +
  geom_tile(aes(fill = counts)) +
  ggplot2::facet_grid(~ TYPE, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "#FFFFCCFF",
    mid = "#FD8D3CFF",
    high = "#800026FF",
    midpoint = 6,
    guide = "colourbar",
    breaks = c(-3,0, 3, 6, 12),
) +
  scale_x_discrete(expand = c(0,0)) +
  scale_y_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0, 0.2, 0, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank(),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank())+
    ggtitle("positive regulation 
of immune response") 
up_GO_1_lcpm_heat
```
##### up 2
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_2_DEG <- subset(DEG_df, DEG_df$gene_id %in% GO_2_top$gene_id)
GO_2_DEG <- merge(GO_2_DEG, GO_2_top, by = "gene_id", all.x = TRUE)
GO_2_DEG$gene <- factor(GO_2_DEG$gene,
                          levels = unique(GO_2_top$gene))
GO_2_DEG$gene <- fct_rev(GO_2_DEG$gene)
GO_2_DEG <- subset(GO_2_DEG, TYPE == "CONTROL" | TYPE ==  "LBD")

# heatmap 
up_GO_2_lcpm_heat <- ggplot(data = GO_2_DEG, aes(x = reorder(NPID, -counts), y = gene)) +
  geom_tile(aes(fill = counts)) +
  ggplot2::facet_grid(~ TYPE, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "#FFFFCCFF",
    mid = "#FD8D3CFF",
    high = "#800026FF",
    midpoint = 6,
    guide = "colourbar",
    breaks = c(-3,0, 3, 6, 12),
) +
  scale_x_discrete(expand = c(0,0)) +
  scale_y_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0, 0.2, 0, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank(),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank())+
    ggtitle("tube morphogenesis") 
up_GO_2_lcpm_heat

```
##### up 3
```{r}
GO_3_DEG <- subset(DEG_df, DEG_df$gene_id %in% GO_3_top$gene_id)
GO_3_DEG <- merge(GO_3_DEG, GO_3_top, by = "gene_id", all.x = TRUE)
GO_3_DEG$gene <- factor(GO_3_DEG$gene,
                          levels = unique(GO_3_top$gene))
GO_3_DEG$gene <- fct_rev(GO_3_DEG$gene)
GO_3_DEG <- subset(GO_3_DEG, TYPE == "CONTROL" | TYPE ==  "LBD")

# heatmap 
up_GO_3_lcpm_heat <- ggplot(data = GO_3_DEG, aes(x = reorder(NPID, -counts), y = gene)) +
  geom_tile(aes(fill = counts)) +
  ggplot2::facet_grid(~ TYPE, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "#FFFFCCFF",
    mid = "#FD8D3CFF",
    high = "#800026FF",
    midpoint = 6,
    space = "rgb",
    guide = "colourbar",
    breaks = c(-3,0, 3, 6, 12),
) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0, 0.2, 0, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("Hemostasis") 
up_GO_3_lcpm_heat
```
##### up 4
```{r}
GO_4_DEG <- subset(DEG_df, DEG_df$gene_id %in% GO_4_top$gene_id)
GO_4_DEG <- merge(GO_4_DEG, GO_4_top, by = "gene_id", all.x = TRUE)
GO_4_DEG$gene <- factor(GO_4_DEG$gene,
                          levels = unique(GO_4_top$gene))
GO_4_DEG$gene <- fct_rev(GO_4_DEG$gene)
GO_4_DEG <- subset(GO_4_DEG, TYPE == "CONTROL" | TYPE ==  "LBD")

# heatmap 
up_GO_4_lcpm_heat <- ggplot(data = GO_4_DEG, aes(x = reorder(NPID, -counts), y = gene)) +
  geom_tile(aes(fill = counts)) +
  ggplot2::facet_grid(~ TYPE, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "#FFFFCCFF",
    mid = "#FD8D3CFF",
    high = "#800026FF",
    midpoint = 6,
    space = "rgb",
    guide = "colourbar",
    breaks = c(-3,0, 3, 6, 12),
) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0, 0.2, 0, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("positive regulation 
of cell migration") 
up_GO_4_lcpm_heat
```
##### up 5
```{r}
GO_5_DEG <- subset(DEG_df, DEG_df$gene_id %in% GO_5_top$gene_id)
GO_5_DEG <- merge(GO_5_DEG, GO_5_top, by = "gene_id", all.x = TRUE)
GO_5_DEG$gene <- factor(GO_5_DEG$gene,
                          levels = unique(GO_5_top$gene))
GO_5_DEG$gene <- fct_rev(GO_5_DEG$gene)
GO_5_DEG <- subset(GO_5_DEG, TYPE == "CONTROL" | TYPE ==  "LBD")

# heatmap 
up_GO_5_lcpm_heat <- ggplot(data = GO_5_DEG, aes(x = reorder(NPID, -counts), y = gene)) +
  geom_tile(aes(fill = counts)) +
  ggplot2::facet_grid(~ TYPE, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "#FFFFCCFF",
    mid = "#FD8D3CFF",
    high = "#800026FF",
    midpoint = 6,
    space = "rgb",
    guide = "colourbar",
    breaks = c(-3,0, 3, 6, 12),
) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0, 0.2, 0, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("TYROBP causal network 
in microglia") 
up_GO_5_lcpm_heat
```
clean up
```{r, eval = FALSE}
remove(
  up_data,
  up_data_melt,
  up_df,
  up_enrich_df,
  up_enrich_df_melt,
  up_enrich_results_subset,
  up_gene_char,
  up_gene_pathway,
  up_input,
  GO_1,
  GO_1_DEG,
  GO_1_top,
  GO_2,
  GO_2_DEG,
  GO_2_top,
  GO_3,
  GO_3_DEG,
  GO_3_top,
  GO_4,
  GO_4_DEG,
  GO_4_top,
  GO_5, 
  GO_5_DEG,
  GO_5_top
)
```
## down
```{r}
down_enrich_results_subset$Description
down_enrich_df <-
  down_enrich_results_subset[, c("GroupID", "Term", "Description", "Symbols")]
down_enrich_df_melt <- reshape2::melt(down_enrich_df)
down_gene_char <- str_split(down_enrich_df_melt$Symbols, ",")

down_gene_pathway <-
  data.frame(gene = unlist(down_gene_char),
             value = rep(down_enrich_df$Description, lengths(down_gene_char)))
down_pathways_levels <- levels(fct_rev(down_gene_pathway$value))

down_input <- list(
  "organic acid catabolic process" = down_gene_char[[1]],
  "response to toxic substance" = down_gene_char[[2]],
  "alcohol metabolic process" = down_gene_char[[10]], 
  "microtubule-based movement" = down_gene_char[[13]],
  "carboxylic acid transport" = down_gene_char[[16]]
)
down_data <- fromList(down_input)
down_data$gene <- row.names(down_data)
down_data_melt <- reshape2::melt(down_data)


# get the fold change value for those genes
down_df <- merge(down_data_melt, DEG_sort,
                       by = "gene")
# order by fold change 
down_df <- down_df[order(down_df$logFC, -down_df$adj.P.Val),]

# subset by GO termes
GO_1 <-
  subset(down_df,
         down_df$variable == "organic acid catabolic process"
         & down_df$value == 1)
GO_1_top <- head(GO_1, 15)

GO_2 <-
  subset(down_df,
         down_df$variable == "response to toxic substance"
         & down_df$value == 1)
GO_2_top <- head(GO_2, 15)

GO_3 <-
  subset(down_df,
         down_df$variable == "alcohol metabolic process"
         & down_df$value == 1)
GO_3_top <- head(GO_3, 15)

GO_4 <-
  subset(down_df,
         down_df$variable == "microtubule-based movement"
         & down_df$value == 1)
GO_4_top <- head(GO_4, 15)

GO_5 <-
  subset(down_df,
         down_df$variable == "carboxylic acid transport"
         & down_df$value == 1)
GO_5_top <- head(GO_5, 15)
```
#### plot lcpm GO heatmaps
##### down 1
```{r}
GO_1_DEG <- subset(DEG_df, DEG_df$gene_id %in% GO_1_top$gene_id)
GO_1_DEG <- merge(GO_1_DEG, GO_1_top, by = "gene_id", all.x = TRUE)
GO_1_DEG$gene <- factor(GO_1_DEG$gene,
                          levels = unique(GO_1_top$gene))
GO_1_DEG$gene <- fct_rev(GO_1_DEG$gene)
GO_1_DEG <- subset(GO_1_DEG, TYPE == "CONTROL" | TYPE ==  "LBD")

# heatmap 
down_GO_1_lcpm_heat <- ggplot(data = GO_1_DEG, aes(x = reorder(NPID, -counts), y = gene)) +
  geom_tile(aes(fill = counts)) +
  ggplot2::facet_grid(~ TYPE, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "#FFFFCCFF",
    mid = "#FD8D3CFF",
    high = "#800026FF",
    midpoint = 6,
    space = "rgb",
    guide = "colourbar",
    breaks = c(-3,0, 3, 6, 12),
) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0, 0.2, 0, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("organic acid 
catabolic process") 
down_GO_1_lcpm_heat

```
##### down 2
```{r}
GO_2_DEG <- subset(DEG_df, DEG_df$gene_id %in% GO_2_top$gene_id)
GO_2_DEG <- merge(GO_2_DEG, GO_2_top, by = "gene_id", all.x = TRUE)
GO_2_DEG$gene <- factor(GO_2_DEG$gene,
                          levels = unique(GO_2_top$gene))
GO_2_DEG$gene <- fct_rev(GO_2_DEG$gene)
GO_2_DEG <- subset(GO_2_DEG, TYPE == "CONTROL" | TYPE ==  "LBD")

# heatmap 
down_GO_2_lcpm_heat <- ggplot(data = GO_2_DEG, aes(x = reorder(NPID, -counts), y = gene)) +
  geom_tile(aes(fill = counts)) +
  ggplot2::facet_grid(~ TYPE, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "#FFFFCCFF",
    mid = "#FD8D3CFF",
    high = "#800026FF",
    midpoint = 6,
    space = "rgb",
    guide = "colourbar",
    breaks = c(-3,0, 3, 6, 12),
) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0, 0.2, 0, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("response to 
toxic substance") 
down_GO_2_lcpm_heat
```
##### down 3
```{r}
GO_3_DEG <- subset(DEG_df, DEG_df$gene_id %in% GO_3_top$gene_id)
GO_3_DEG <- merge(GO_3_DEG, GO_3_top, by = "gene_id", all.x = TRUE)
GO_3_DEG$gene <- factor(GO_3_DEG$gene,
                          levels = unique(GO_3_top$gene))
GO_3_DEG$gene <- fct_rev(GO_3_DEG$gene)
GO_3_DEG <- subset(GO_3_DEG, TYPE == "CONTROL" | TYPE ==  "LBD")

# heatmap 
down_GO_3_lcpm_heat <- ggplot(data = GO_3_DEG, aes(x = reorder(NPID, -counts), y = gene)) +
  geom_tile(aes(fill = counts)) +
  ggplot2::facet_grid(~ TYPE, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "#FFFFCCFF",
    mid = "#FD8D3CFF",
    high = "#800026FF",
    midpoint = 6,
    space = "rgb",
    guide = "colourbar",
    breaks = c(-3,0, 3, 6, 12),
) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0, 0.2, 0, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("alcohol metabolic 
process") 
down_GO_3_lcpm_heat

```
##### down 4
```{r}
GO_4_DEG <- subset(DEG_df, DEG_df$gene_id %in% GO_4_top$gene_id)
GO_4_DEG <- merge(GO_4_DEG, GO_4_top, by = "gene_id", all.x = TRUE)
GO_4_DEG$gene <- factor(GO_4_DEG$gene,
                          levels = unique(GO_4_top$gene))
GO_4_DEG$gene <- fct_rev(GO_4_DEG$gene)
GO_4_DEG <- subset(GO_4_DEG, TYPE == "CONTROL" | TYPE ==  "LBD")

# heatmap 
down_GO_4_lcpm_heat <- ggplot(data = GO_4_DEG, aes(x = reorder(NPID, -counts), y = gene)) +
  geom_tile(aes(fill = counts)) +
  ggplot2::facet_grid(~ TYPE, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "#FFFFCCFF",
    mid = "#FD8D3CFF",
    high = "#800026FF",
    midpoint = 6,
    space = "rgb",
    guide = "colourbar",
    breaks = c(-3,0, 3, 6, 12),
) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0, 0.2, 0, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("microtubule-based 
movement") 
down_GO_4_lcpm_heat
```
##### down 5
```{r}
GO_5_DEG <- subset(DEG_df, DEG_df$gene_id %in% GO_5_top$gene_id)
GO_5_DEG <- merge(GO_5_DEG, GO_5_top, by = "gene_id", all.x = TRUE)
GO_5_DEG$gene <- factor(GO_5_DEG$gene,
                          levels = unique(GO_5_top$gene))
GO_5_DEG$gene <- fct_rev(GO_5_DEG$gene)
GO_5_DEG <- subset(GO_5_DEG, TYPE == "CONTROL" | TYPE ==  "LBD")

# heatmap 
down_GO_5_lcpm_heat <- ggplot(data = GO_5_DEG, aes(x = reorder(NPID, -counts), y = gene)) +
  geom_tile(aes(fill = counts)) +
  ggplot2::facet_grid(~ TYPE, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "#FFFFCCFF",
    mid = "#FD8D3CFF",
    high = "#800026FF",
    midpoint = 6,
    space = "rgb",
    guide = "colourbar",
    breaks = c(-3,0, 3, 6, 12),
) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0, 0.2, 0, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("carboxylic acid 
transport") 
down_GO_5_lcpm_heat

```
### legend only
```{r}
# Extract the legend. Returns a gtable
# lock in gene order which is sorted by adjusted p-value
GO_1_DEG <- subset(DEG_df, DEG_df$gene_id %in% GO_1_top$gene_id)
GO_1_DEG <- merge(GO_1_DEG, GO_1_top, by = "gene_id", all.x = TRUE)
GO_1_DEG$gene <- factor(GO_1_DEG$gene,
                          levels = unique(GO_1_top$gene))
GO_1_DEG$gene <- fct_rev(GO_1_DEG$gene)
GO_1_DEG <- subset(GO_1_DEG, TYPE == "CONTROL" | TYPE ==  "LBD")

# heatmap 
legend_data <- ggplot(data = GO_1_DEG, aes(x = reorder(NPID, -counts), y = gene)) +
  geom_tile(aes(fill = counts)) +
  ggplot2::facet_grid(~ TYPE, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "#FFFFCCFF",
    mid = "#FD8D3CFF",
    high = "#800026FF",
    midpoint = 6,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 4, 8, 12, 16),
    name = expression(log[2](CPM))
  ) +
  theme(
    legend.position = "top", 
    legend.text = element_text(size =6), 
    legend.title = element_text (size = 6), 
    legend.margin=margin(0,0,1,0),
  #  legend.direction = "horizontal", 
    legend.key.size = unit(0.5, "cm"))

legend_data <- addSmallLegend(legend_data)

leg <- get_legend(legend_data)
# Convert to a ggplot and print
legend_heat_bar <- as_ggplot(leg)
legend_heat_bar
```
clean up
```{r, eval =FALSE}
remove(
  down_data,
  down_data_melt,
  down_df,
  down_enrich_df,
  down_enrich_df_melt,
  down_enrich_results_subset,
  down_gene_char,
  down_gene_pathway,
  down_input,
  GO_1,
  GO_1_DEG,
  GO_1_top,
  GO_2,
  GO_2_DEG,
  GO_2_top,
  GO_3,
  GO_3_DEG,
  GO_3_top,
  GO_4,
  GO_4_DEG,
  GO_4_top,
  GO_5, 
  GO_5_DEG,
  GO_5_top
)
```

# Gene-level combine all plots 
```{r}
row2 <-
  ggarrange(
    NULL,
    legend_heat_bar,
    NULL,
    NULL,
    ncol = 4,
    labels = c("up-regulated", "", "", ""),
    font.label = list(size = 7, color = "red")
  )

row3 <- ggarrange(
  up_GO_1_lcpm_heat,
  up_GO_2_lcpm_heat,
  up_GO_3_lcpm_heat,
  up_GO_4_lcpm_heat,
  up_GO_5_lcpm_heat,
  ncol = 5,
  labels = c("c)", "d)", "e)", "f)", "g)"),
  font.label = list(size = 8)
  )

row4 <-
  ggarrange(
    NULL,
    NULL,
    NULL,
    NULL,
    ncol = 4,
    labels = c("down-regulated", "", "", ""),
    font.label = list(size = 7, color = "blue")
  )

row5 <- ggarrange(
  down_GO_1_lcpm_heat,
  down_GO_2_lcpm_heat,
  down_GO_3_lcpm_heat,
  down_GO_4_lcpm_heat,
  down_GO_5_lcpm_heat,
  ncol = 5,
  labels = c("h)", "i)", "j)", "k)", "l)"),
  font.label = list(size = 8)
  )

combind <-
  ggarrange(
    row1,
    row2,
    row3,
    row4,
    row5,
    nrow = 5,
    heights = c(1.45, 0.1, 1, 0.1, 1)
  )
combind

path <- paste0("../../../results/manuscript_figures/Figure_1_LBD_vs_Control")
#saveToPDF(paste0(path, ".pdf"), width = 8, height = 11.5)

png(file="../../../results/manuscript_figures/Figure_1_LBD_vs_Control.png",
width=8, height=11, units="in", res = 1500)
combind
dev.off()

jpeg(file="../../../results/manuscript_figures/Figure_1_LBD_vs_Control.jpeg",
width=8, height=11, units="in", res = 1500)
combind
dev.off()
```
```{r}
```
