---
title: "Linseed deconvolution"
author: "Kimberly Olney"
date: "4/20/2023"
output: html_document
---
Linseed (LINear Subspace identification for gene Expresion Deconvolution) is a package that provides tools and interface to explore gene expression datasets in linear space.
https://github.com/ctlab/linseed
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Install packages
```{r install, eval=FALSE}
library(vctrs, lib.loc = "/usr/local/biotools/rpackages/R-4.1.2-2021-11-11")
devtools::install_github("ctlab/linseed")
```
# Load libraries
```{r load}
set.seed(28)
library(linseed)
```
# Load GEO example dataset
To start working with gene expression data, we need to create a new LinseedObject, in this tutorial we will use GSE19830 (mixture of Liver, Brain and Lung), we will take only mixed samples (10-42) and will take only 10000 most expressed genes.
```{r dge}
dge.filtered.norm <- readRDS(paste0("../../rObjects/dge.filtered.norm.rds"))
lo_dge <- LinseedObject$new(dge.filtered.norm$counts, topGenes=10000)
```
# Coolinearity networks
To build a coolinearity network we first have to evaluate all pairwise collinearity coefficients, all pairwise spearman correlation and then run significance test which will calculate p value for each each by shuffling network weights randomly.
```{r}
lo_dge$calculatePairwiseLinearity()
lo_dge$calculateSpearmanCorrelation()
lo_dge$calculateSignificanceLevel(100)
lo_dge$significancePlot(0.01)
```
# Filter genes
```{r}
lo_dge$filterDatasetByPval(0.01)
lo_dge$svdPlot()
```
# After Filtering projection
To visualiaze what left after filtering we can call projection plot from our object. But we have to project the data to the simplex first.
```{r}
lo_dge$setCellTypeNumber(5)
lo_dge$project("full") # projecting full dataset
lo_dge$projectionPlot(color="filtered")
```
# Deconvolution
To deconvolve the dataset, you first have to project (full or filtered dataset) to the simplex, and then find corners of it.
```{r}
lo_dge$project("filtered")
lo_dge$smartSearchCorners(dataset="filtered", error="norm")
```
# Plot
```{r}
lo_dge$deconvolveByEndpoints()
ct_proportions <- as.data.frame(lo_dge$proportions)
plotProportions(ct_proportions)
```
# TSNE
```{r}
# lets select 100 genes closest to the simplex corners 
lo_dge$selectGenes(100)
lo_dge$tsnePlot()
```
# Cell markers
```{r}
cell_type_markers <- lo_dge$markers
expression <- lo_dge$exp
```
# PCA
```{r}
gene_pvals <- lo_dge$genes
df <- as.data.frame(gene_pvals)
# Running PCA
pca_exp <- prcomp(t(highly_variable_exp), scale = F, center = T)
# scale a logical value indicating whether the variables should be scaled
# to have unit variance before the analysis takes place.
# a logical value indicating whether the variables should be shifted to be zero centered.
head(pca_exp$x)[, 1:3]
summary(pca_exp)
# Dataframe with the first 10 PCs
dim1_10 <- data.frame(pca_exp$x[, 1:10])
# Adding metadata
dim1_10$NPID <- rownames(dim1_10)
pcaWithMetadata <- merge(dim1_10, info_with_scale, by = "NPID", all = TRUE)
pcaWithMetadata$group <- pcaWithMetadata$TYPE

# Plotting
ggplot(data = pcaWithMetadata, aes(x = PC1, y = PC2, shape = group, color = group)) +
  geom_point(size = 2.5) +
  theme_bw() +
  scale_color_manual(values = colorbindColors) 

```
# Merge with sample information
```{r}
dge.filtered.norm <- readRDS(paste0("../../rObjects/dge.filtered.norm.rds"))
# get the gene id that corresponds to the gene name 
genes <- dge.filtered.norm$genes # gene information 
goi <- subset(genes, gene_id == "ENSG00000152583.13") # gene of interest 
ENSG00000162734.13
```