---
title: "PSEA deconvolution"
author: "Kimberly Olney"
date: "4/20/2023"
output: html_document
---
PSEA: Expression deconvolution of neural RNA mixtures
Tutorial: https://www.bioconductor.org/packages/devel/bioc/vignettes/PSEA/inst/doc/PSEA_RNAmixtures.pdf 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Install packages
```{r install, eval=FALSE}
BiocManager::install("PSEA")
devtools::install_github("ctlab/linseed")
```
# Load libraries
```{r load}
library(PSEA)
library(affy)
library(GEOquery)
```
# Load GEO example dataset
```{r GEO}
data(expression_GSE19380)
# We start by removing the control probesets.
expression <- expression_GSE19380[1:31042,]
```
# Define marker probes
```{r}
neuron_probesets<-list(c("1370058_at","1370059_at"),"1387073_at","1367845_at")
astro_probesets<-list("1372190_at","1386903_at",c("1375120_at","1375183_at", "1385923_at"))
oligo_probesets<-list("1398257_at","1368861_a_at",c("1368263_a_at","1370434_a_at", "1370500_a_at"))
```
We normalize the signals using mixed samples only as we re-strict the use of pure samples to the validation of deconvoluted expression. Mixed samples correspond to column 17 to 24 of the expression matrix, as indicated in the corresponding sample information. 

Generates a population-specific reference signal.
marker(expr,id,sampleSubset=NULL,targetMean=1)
expr = The matrix contains the expression data
id = Names of the transcripts to use to generate the reference signal.
sampleSubset = integer vector. Represents a subset of samples (specified as column indices in expr) to use to normalize the reference signal. Default is all samples. 
targetMean = Numeric. Mean of the reference signal (defaults to 1)
```{r}
mixedsamples<-c(17:24)
neuron_reference<-marker(expression,neuron_probesets,sampleSubset=mixedsamples,targetMean = 100)
astro_reference<-marker(expression,astro_probesets,sampleSubset=mixedsamples,targetMean = 100)
oligo_reference<-marker(expression,oligo_probesets,sampleSubset=mixedsamples,targetMean = 100)
```
# Plot the nueronal reference signal across all samples 
```{r}
par(cex=0.7)
plot(neuron_reference,type="l")
```
# Model fit
Fit the signal measured by probeset 1367660_at in the mixed samples with an expression model including all 3 populations 
```{r}
model1<-lm(expression["1367660_at",]~neuron_reference+astro_reference+
+ oligo_reference,subset=mixedsamples)
```
# Assess model fit
We can use component-plus-residual plots to visualize the dependence of expression on the 4 reference signals.

Component-plus-residual (CR) plot for quantitative variables and group-specific effects.
crplot(lm,quantv,g=NULL,newplot=TRUE,col=c(1,2,4),xlab=NULL,ylab='CR',...)
lm = fitted model
quantv	= character. Name of the quantitative independent variable ("component" to be plotted).
g = 	character vector. Interaction regressors coding for group-specific effects in the model.
```{r}
par(mfrow=c(1,3),cex=0.7)
crplot(model1,"neuron_reference",newplot=FALSE)
crplot(model1,"astro_reference",newplot=FALSE,ylim=c(-250,250))
crplot(model1,"oligo_reference",newplot=FALSE,ylim=c(-250,250))
```
# Insepect the fitted expression model
```{r}
summary(model1)
```
# Deconvolution
We now deconvolute the entire expression profile (i.e. all probesets) obtained for the mixed samples. We start by defining the full model matrix
```{r}
model_matrix<-cbind(intercept=1,neuron_reference,astro_reference,oligo_reference)

# specify the subset of models under consideration
model_subset<-em_quantvg(c(2,3,4), tnv=3, ng=1)

#fit each probeset with all models in the subset and select the best model.
models<-lmfitst(t(expression), model_matrix, model_subset, subset=mixedsamples)
```
# Extract coefficients, pvalues, and adjusted R^2 for the selected models 
```{r}
regressor_names<-as.character(1:4)
coefficients<-coefmat(models[[2]], regressor_names)
pvalues<-pvalmat(models[[2]], regressor_names)
models_summary<-lapply(models[[2]], summary)
adjusted_R2<-slt(models_summary, 'adj.r.squared')
```
# Filter satisfactory expression models
```{r}
negativecoefficient<-apply(coefficients[,-1]<0 & pvalues[,-1]<0.05,1,function(x){any(x, na.rm = T)})
average_expression<-apply(expression[,mixedsamples], 1, mean)
filter<-!negativecoefficient & (coefficients[,1] / average_expression) < 0.5 & adjusted_R2 > 0.6
```
# Results
```{r}
# number of probesets with non-negative coefficients:
sum(!negativecoefficient)
# Number of probesets with relative intercept < 0.5:
sum(coefficients[,1] / average_expression < 0.5)
# Number of probesets with adjusted R2 > 0.6:
sum(adjusted_R2 > 0.6)
# Number of probesets passing all 3 criteria:
sum(filter)
```
# Spot check
We can for instance inspect the expression model for probeset 1370431_at (replicates Sup- plementary Figure 4d, middle panel).
```{r}
selectedpsname<-"1370431_at"
selectedps<-which(rownames(expression)==selectedpsname)

```
# Coeffients
The parameter β (the regression coefficient) signifies the amount by which change in x must be multiplied to give the corresponding average change in y, or the amount y changes for a unit increase in x. In this way it represents the degree to which the line slopes upwards or downwards.
```{r}
# The neuron-specific expression for 1370431_at is
coefficients[selectedps,2]

# Corresponding p-value
pvalues[selectedps,2]

# The dependence on the neuronal reference signal is visualized as follows.
crplot(models[[2]][[selectedps]],"2",ylim=c(0,950))
```

