---
title: "Variance in expression data include effect of sex "
author: "Kimberly Olney, Ph.D."
date: "01/06/2023"
output:
  html_document:
    df_print: paged
  pdf_document: default
params:
  args: myarg
---
# Setup
```{r setup}
knitr::opts_knit$set(root.dir = ".")
```

```{r libraries, message=FALSE, warning=FALSE}
library('variancePartition')
library('edgeR')
library('BiocParallel')
library('Hmisc')
```

# User defined variables
```{r set_variables}
source(here::here("scripts/R", "file_paths_and_colours.R"))
source(here::here("scripts/R", "gtf_path.R"))
condition <- c("All_samples")
min_expression <- "min_exp1"

myContrasts <- c("LBDvsControl = LBD - CONTROL",
  "LBDvsAD = LBD - AD",
  "LBDvsPA = LBD - PA",
  "ADvsControl = AD - CONTROL", 
  "PAvsControl = PA - CONTROL",
  "ADvsPA = AD - PA")
tool = c("star")
typeOfCount <-  c(".bamReadsPerGene.out.tab")
```
# Read in DGE object & metadata
```{r}
dge <- readRDS(paste0("../../rObjects/", condition, ".dge.raw.rds"))
metadata <- dge$samples
dge.filtered.norm <- readRDS(paste0("../../rObjects/", condition, "_", min_expression, ".dge.filtered.norm.rds"))
# some samples are missing RIN values. 
# Replace NA with median RIN. 
# This is necessary to be able include RIN as a covariate in voom
# fill missing values of marks2 with median
dge.filtered.norm$samples$RIN <- impute(dge.filtered.norm$samples$RIN, median)
# one sample is missing VaD information
dge.filtered.norm$samples$VaD <- impute(dge.filtered.norm$samples$VaD, median)
dge.filtered.norm$samples$flowcell_and_lane <- factor(dge.filtered.norm$samples$flowcell_and_lane)
dge.filtered.norm$samples$APOE <- factor(dge.filtered.norm$samples$APOE)

info <- as.data.frame(dge.filtered.norm$samples)
genes <- dge.filtered.norm$genes
log2cpm.norm <- edgeR::cpm(dge.filtered.norm, log = TRUE)

# Add new condition information for saving results 
condition <- c("All_samples_sex_effect")
```

# BIC with forward stepwise regression
First, we will scale some of the continuous variables. 
In regression, it is often recommended to scale the features so that the predictors have a mean of 0. This makes it easier to interpret the intercept term as the expected value of Y when the predictor values are set to their means.
scale is generic function whose default method centers and/or scales the columns of a numeric matrix.

Secondly, obtian voom$E counts to use in the BIC. See BIC 


Finally, perform forwards stepwise regression Bayesian information criterion (BIC) to determine the best  model. 
see: https://rdrr.io/github/GabrielHoffman/mvIC/man/mvIC_fit.html 
```{r}
scaled.info <-
  info[c(
    "Race_numeric",
    "RIN",
    "Age",
    "PCT_CODING_BASES",
    "PCT_INTERGENIC_BASES",
    "PCT_INTRONIC_BASES",
    "APOE_E4_allele_count"
  )] %>% scale()
scaled.info.df <- as.data.frame(scaled.info)

info_with_scale <- cbind(info, scaled.info.df)
```

```{r eval=FALSE}
formula <- ( ~ 0 + TYPE)
voom_with_weights <-
  variancePartition::voomWithDreamWeights(
    counts = dge.filtered.norm$counts,
    formula = formula,
    data = dge.filtered.norm$samples,
    BPPARAM = BiocParallel::SnowParam(cores),
    plot = TRUE
  )
path <- paste0("../../results/", tool, "/voom/", condition, "_", min_expression, "_gene_mean_voom_with_weights_no_covariates")
saveToPDF(paste0(path, ".pdf"), width = 6, height = 4)
voomCounts <- voom_with_weights$E
```

```{r eval=FALSE}
baseFormula <- ~ (1 | TYPE)
# Combine responses on *rows*
Y = with(
  info,
  rbind(
    sex_inferred,
    Race,
    flowcell_and_lane,
    APOE,
    scaled.info.df$RIN,
    scaled.info.df$Age,
    scaled.info.df$APOE_E4_allele_count,
    scaled.info.df$PCT_CODING_BASES,
    scaled.info.df$PCT_INTERGENIC_BASES,
    scaled.info.df$PCT_INTRONIC_BASES
  )
)

rownames(Y) <-
  c(
    "sex_inferred",
    "Race",
    "flowcell_and_lane",
    "APOE",
    "RIN",
    "Age",
    "APOE_E4_allele_count",
    "PCT_CODING_BASES",
    "PCT_INTERGENIC_BASES",
    "PCT_INTRONIC_BASES"
  )
# variables to consider in the model
variables = c(
  "(1|sex_inferred)",
  "(1|Race)",
  "(1|flowcell_and_lane)",
  "(1|APOE)",
  "RIN",
  "Age",
  "APOE_E4_allele_count",
  "PCT_CODING_BASES",
  "PCT_INTERGENIC_BASES",
  "PCT_INTRONIC_BASES"
)

# fit forward stepwise regression starting
bestModel_voomcounts = mvForwardStepwise(voomCounts,
                                         baseFormula,
                                         data = info,
                                         variables = variables)
bestModel_voomcounts
```
Best model: ~ (1 | TYPE) + PCT_CODING_BASES + (1 | flowcell_and_lane) + (1 | TYPE:sex_inferred) + PCT_INTERGENIC_BASES + PCT_INTRONIC_BASES + RIN + Age 

# Design matrix
```{r}
design <-
  model.matrix(~ 0 + TYPE +
      TYPE*sex_inferred + 
      scale(RIN) + 
      scale(Age) + 
      flowcell_and_lane + 
      scale(PCT_CODING_BASES) + 
      scale(PCT_CODING_BASES) + 
      scale(PCT_INTERGENIC_BASES) + 
      scale(PCT_INTRONIC_BASES),
    dge.filtered.norm$samples
  )

colnames(design) <-
  c(
    CONTROL,
    AD,
    PA,
    LBD,
    "sex",
    "RIN",
    "Age",
    "Batch1",
    "Batch2",
    "Batch3",
    "Batch4",
    "Batch5",
    "Batch6",
    "Batch7",
    "PCT_CODING_BASES",
    "PCT_INTERGENIC_BASES",
    "PCT_INTRONIC_BASES", 
    "AD_SexEffect",
    "PA_SexEffect",
    "LBD_SexEffect"
  )
```
# Voom
When the library sizes are quite variable between samples, then the voom approach is theoretically more powerful than limma-trend. 
If the data are very noisy, one can apply the same between-array normalization methods as would be used for microarrays, for example:
```{r voom}
form <- (
  ~ 0 +
    TYPE +
    TYPE*sex_inferred +
    scale(RIN) +
    scale(Age) +
    flowcell_and_lane +
    scale(PCT_CODING_BASES) +
    scale(PCT_CODING_BASES) +
    scale(PCT_INTERGENIC_BASES) +
    scale(PCT_INTRONIC_BASES)
)
voom_with_weights_and_covariates <-
  variancePartition::voomWithDreamWeights(
    counts = dge.filtered.norm$counts,
    formula = form,
    data = dge.filtered.norm$samples,
    BPPARAM = BiocParallel::SnowParam(cores),
    plot = TRUE
  )
path <-
  paste0("../../results/",
         tool,
         "/voom/",
         condition,
         "_",
         min_expression,
         "_gene_mean_voom_with_weights_and_covariates")
saveToPDF(paste0(path, ".pdf"), width = 6, height = 4)

voomCounts_covariates <- voom_with_weights_and_covariates$E
voomCountsMatrix <- voom_with_weights_and_covariates$E
```
# Contrast plot
```{r}
# form <- ~ 0 + TYPE
myDreamContrasts <- c(LBDvsControl = "TYPELBD - TYPECONTROL",
  LBDvsAD = "TYPELBD - TYPEAD",
  LBDvsPA = "TYPELBD - TYPEPA",
  ADvsControl = "TYPEAD - TYPECONTROL", 
  PAvsControl = "TYPEPA - TYPECONTROL",
  ADvsPA = "TYPEAD - TYPEPA")
myDreamContrasts<- rev(myDreamContrasts)

L = makeContrastsDream(form, dge.filtered.norm$samples, contrasts = myDreamContrasts)
colnames(L) <- gsub("TYPE","",colnames(L))
rownames(L) <- gsub("TYPE","", rownames(L))

# Visualize contrast matrix
plotContrasts(L) 
```

```{r}
# fits linear model for each gene given a series of arrays
fit <- lmFit(voom_with_weights_and_covariates, design)
coef.fit <- fit$coefficients
head(coef.fit)
# contrast design for differential expression
contrasts <- makeContrasts(
  LBDvsControl = LBD - CONTROL,
  LBDvsAD = LBD - AD,
  LBDvsPA = LBD - PA,
  ADvsControl = AD - CONTROL, 
  PAvsControl = PA - CONTROL,
  ADvsPA = AD - PA, 
  levels = colnames(design))
head(contrasts)

# save contrast names
allComparisons <- colnames(contrasts)
allComparisons # check

# run contrast analysis
vfit <- contrasts.fit(fit, contrasts = contrasts)

# Compute differential expression based on the empirical Bayes moderation of the
# standard errors towards a common value.
# The logCPM values can then be used in any standard limma pipeline, using the trend=TRUE
# argument when running eBayes or treat. For example:
veBayesFit <- eBayes(vfit, trend = TRUE, robust=TRUE)
plotSA(veBayesFit, main = "Final Model: Mean-variance Trend")

# Rather than worry about the normalization too much, better to explore the data. E.g. try a BCV plot to look for dispersion outliers, or try robust=TRUE with eBayes() to downweight dispersion outliers. 

#disp <- estimateDisp(dge.filtered.norm, design, robust=TRUE)
#plotBCV(disp)
```

# DEGs summary
```{r}
pval <- 0.05
lfc.cutoff <- 0.25

sumTable <- 
  summary(decideTests(
    veBayesFit,  # object
    adjust.method = "BH", # by default the method = "separate"
    p.value = pval,
    lfc = lfc.cutoff  # numeric, minimum absolute log2-fold change required
  ))

print(paste0(" FDRq < ", pval,
             " & absolute log2-fold change > ", lfc.cutoff))
sumTable
```
# Add gene information to DEGs
reformat genes table to only include relevant information
```{r}
genes_relevant <- select(genes, 1:4,10:12)
```
Check AD vs Control
```{r}
ADvsControl <- topTable(
  veBayesFit, 
  coef = 4,  
  n = Inf, 
  p.value = 1,
  lfc = 0, 
  sort.by = "P", 
  genelist = genes_relevant, 
  confint = TRUE # column of confidence interval 
    )
head(ADvsControl, 20)
```
# Save objects
```{r}
saveRDS(veBayesFit, file = paste0("../../rObjects/", condition, "_", min_expression,".veBayesFit.rds"))
saveRDS(voomCounts_covariates, file = paste0("../../rObjects/", condition, "_", min_expression,".voomCountsMatrix.rds"))
```
# Output DEG tables
```{r}
coef <- 1

for (i in allComparisons) {
  vTopTableAll <- topTable(
    veBayesFit, 
    coef = coef,  
    n = Inf, 
    p.value = 1,
    lfc = 0, 
    sort.by = "P", 
    genelist = genes_relevant, 
    confint = TRUE # column of confidence interval 
    )
    saveRDS(vTopTableAll, file = 
            paste0("../../rObjects/", condition, "_", 
                   i,"_gene_table.rds"))
  path <- paste0("../../results/", tool, "/DEGs/", condition, "_", 
  i, "_gene_DEGs_FDRq1.00.txt", sep = "") 
  write.table(
    vTopTableAll,
    path,
    sep = "\t",
    row.names = FALSE,
    quote = FALSE
  )
  # p < 0.05, log2fc > 0
  vTopTable1 <-
    topTable( 
      veBayesFit,  
      coef = coef,  
      n = Inf, 
      p.value = pval,
      lfc = lfc.cutoff
    )
  path <- paste0("../../results/", tool, "/DEGs/", condition, "_", 
  i, "_gene_DEGs_FDRq0.05_logFC_0.25.txt", sep = "") 
  write.table(
    vTopTable1,
    path,
    sep = "\t",
    row.names = FALSE,
    quote = FALSE
  )
  # increment 
  coef <- coef + 1
}
remove(coef)
```

# PCA
```{r}
# Setting the N of genes to use
ntop = length(dge.filtered.norm$genes$gene_id)
# Sorting by the coefficient of variance
means <- rowMeans(voomCounts_covariates)
Pvars <- rowVars(voomCounts_covariates)
cv2 <- Pvars / means ^ 2
select <-
  order(cv2, decreasing = TRUE)[seq_len(min(ntop, length(cv2)))]
head(select)

highly_variable_exp <- ((voomCounts_covariates)[select,])
dim(highly_variable_exp)
# Running PCA
pca_exp <- prcomp(t(highly_variable_exp), scale = F, center = T)
# scale a logical value indicating whether the variables should be scaled
# to have unit variance before the analysis takes place.
# a logical value indicating whether the variables should be shifted to be zero centered.
head(pca_exp$x)[, 1:3]
summary(pca_exp)
# Dataframe with the first 10 PCs
dim1_10 <- data.frame(pca_exp$x[, 1:10])
# Adding metadata
pcaWithMetadata <- merge(dim1_10, dge.filtered.norm$samples, by = 0, all = TRUE)
pcaWithMetadata$group <- factor(pcaWithMetadata$TYPE,
                                 levels = c(CONTROL, AD, PA, LBD, NA))
# Plotting
ggplot(data = pcaWithMetadata, aes(
  x = PC2,
  y = PC3,
  shape = group,
  color = group
)) +
  geom_point(size = 2.5) +
  theme_bw() +
#  xlim(-150, 150) +
#  ylim(-150, 150) +
  scale_color_manual(values = c(control_color, AD_color,
                                PA_color, LBD_color,
                                "azure3")) +
  scale_shape_manual(values = c(control_shape, AD_shape,
                                PA_shape, LBD_shape)) +
  theme(
    plot.title = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(size = 10),
    axis.title.y = element_text(size = 10)
  ) +
  theme(legend.title = element_blank()) +
  xlab("PC1 ") +
  ylab("PC2 ")
```
# CCA PC1-10 & variables in model 
```{r, eval = FALSE}
form_PCA <- ~ TYPE + 
  flowcell_and_lane +
  sex_inferred + 
  Age + 
  RIN +
  PCT_CODING_BASES +
  PCT_INTERGENIC_BASES +
  PCT_INTRONIC_BASES +
  PC1 +
  PC2 +
  PC3 +
  PC4 +
  PC5 +
  PC6 +
  PC7 +
  PC8 +
  PC9 +
  PC10
  
C = canCorPairs( form_PCA, pcaWithMetadata)
# Plot correlation matrix
cor.mtest <- function(C, ...) {
    C <- as.matrix(C)
    n <- ncol(C)
    p.mat<- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
        for (j in (i + 1):n) {
            tmp <- cor.test(C[, i], C[, j], ...)
            p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
        }
    }
  colnames(p.mat) <- rownames(p.mat) <- colnames(C)
  p.mat
}
# matrix of the p-value of the correlation
p.mat <- cor.mtest(C)
col <- colorRampPalette(c("#4477AA", "#77AADD", "#FFFFFF", "#EE9988", "#BB4444"))
  corrplot(C, method="color", col=col(200),  
         type="upper", order="hclust", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=45, #Text label color and rotation
         diag=FALSE, col.lim = c(0, 1)
         )
path <- paste0("../../results/", tool ,"/varpart/", condition, "_CCA_covariartes_and_PC1_10")
saveToPDF(paste0(path, ".pdf"), width = 16, height = 16)
```
# Blood & white matter marker expression among groups 
```{r}
# get the gene id that corresponds to the gene name 
genes <- dge.filtered.norm$genes

bloodmarker_gene_info <- subset(genes, 
                                gene_name == "HBB" |  
                                gene_name == "HBA1" |
                                gene_name == "HBA2" |
                                gene_name == "HBD" |
                                gene_name == "HBQ1" |
                                gene_name == "MOG" |
                                gene_name == "MBP" |
                                gene_name == "RBFOX3")
bloodmarker_gene_id <- rownames(bloodmarker_gene_info)
# subset counts 
HBB_counts <- subset(voomCountsMatrix, rownames(voomCountsMatrix) %in% bloodmarker_gene_id)
HBB_melt <- reshape2::melt(HBB_counts)
names(HBB_melt)[names(HBB_melt) == "Var2"] <- "NPID"
names(HBB_melt)[names(HBB_melt) == "Var1"] <- "gene_id"
gene_id_gene_name <- data.frame(rownames(bloodmarker_gene_info), bloodmarker_gene_info$gene_name)
names(gene_id_gene_name)[names(gene_id_gene_name) == "rownames.bloodmarker_gene_info."] <- "gene_id"
names(gene_id_gene_name)[names(gene_id_gene_name) == "bloodmarker_gene_info.gene_name"] <- "gene_name"
df <- merge(gene_id_gene_name, HBB_melt, by = "gene_id")
df <- merge(df, info, by = "NPID")
```

```{r}
column_variables <- c("HBB", "HBA1", "HBA2", "HBD", "HBQ1", "MOG", "MBP", "RBFOX3")
my_comparisons <-
  list(
    c("CONTROL", "AD"),
    c("CONTROL", "PA"),
    c("CONTROL", "LBD"),
    c("AD", "PA"),
    c("AD", "LBD"),
    c("PA", "LBD")
  )
violin_plot_fun <- function(j) {
  df <- subset(df, gene_name == j)
  ggplot(df, aes(TYPE, value, fill = TYPE)) +
    geom_violin() +
    geom_boxplot(width = 0.1) + 
    geom_jitter(shape=16, position=position_jitter(0.2)) +
    theme_bw() + 
    ggtitle(j) +
    xlab("Disease group") +
    ylab(j) +
    stat_compare_means(comparisons = my_comparisons) +
    scale_fill_manual(values=TypeColors) +
    theme(legend.position = "none")
}
violin_plots <-
  Map(violin_plot_fun, j = column_variables)
violin_plots

plotnames = imap(violin_plots, ~paste0("../../results/", tool, "/blood_markers/Type.", .y, ".pdf")) %>%
  flatten()
plotnames
walk2(plotnames, violin_plots, ~ggsave(filename = .x, plot = .y, 
                                    height = 4, width = 5.5))


# blood marker expression between VaD classification 
column_variables <- c("HBB", "HBA1", "HBA2", "HBD", "HBQ1")
my_comparisons <- list(c("0", "1"))
df$VaD <- as.factor(df$VaD)
violin_plot_fun <- function(j) {
  df <- subset(df, gene_name == j)
  ggplot(df, aes(VaD, value, fill = VaD)) +
    geom_violin() +
    geom_boxplot(width = 0.1) + 
    geom_jitter(shape=16, position=position_jitter(0.2)) +
    theme_bw() + 
    ggtitle(j) +
    xlab("VaD classification") +
    ylab(j) +
    stat_compare_means(comparisons = my_comparisons) +
    theme(legend.position = "none")
}
violin_plots <-
  Map(violin_plot_fun, j = column_variables)
violin_plots

plotnames = imap(violin_plots, ~paste0("../../results/", tool, "/blood_markers/VaD.", .y, ".pdf")) %>%
  flatten()
plotnames
walk2(plotnames, violin_plots, ~ggsave(filename = .x, plot = .y, 
                                    height = 4, width = 5.5))
```

```{r}
sessionInfo()
```



