---
title: "Upset and correlations among pairwise DEGs"
author: "Kimberly Olney"
date: "01/06/2023"
output: html_document
---

# Setup
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = ".")
```
# Library packages
```{r packages}
library(UpSetR)
library(ComplexUpset)
library(ggplot2movies)
library(ggvenn)
library(forcats)
```
# Function to create tables
```{r}
fromList <- function (input) {
  # Same as original fromList()...
  elements <- unique(unlist(input))
  data <- unlist(lapply(input, function(x) {
      x <- as.vector(match(elements, x))
      }))
  data[is.na(data)] <- as.integer(0)
  data[data != 0] <- as.integer(1)
  data <- data.frame(matrix(data, ncol = length(input), byrow = F))
  data <- data[which(rowSums(data) != 0), ]
  names(data) <- names(input)
  # ... Except now it conserves your original value names!
  row.names(data) <- elements
  return(data)
}

makePaddedDataFrame <- function(l, ...) {
  maxlen <- max(sapply(l, length))
  data.frame(lapply(l, na.pad, len = maxlen), ...)
}
```
# User defined variables
```{r set_variables}
source(here::here("scripts/R", "file_paths_and_colours.R"))
condition <- c("protein_coding")
tool <- "star"
min_expression <- "min_exp1"
# save contrast names
allComparisons <- c("LBDvsControl", 
                    "LBDvsAD", 
                    "LBDvsPA", 
                    "ADvsControl", 
                    "PAvsControl", 
                    "ADvsPA")
allComparisons # check
```
# Read in DEGs
Read table with all genes (FDRq = 1).
```{r read_DEG_table}
coef <- 1
for (i in allComparisons) {
  filepath <- paste0(
    "../../results/",
    tool,
    "/DEGs/", condition, "_BIC_", 
    i,
    "_gene_DEGs_FDRq1.00.txt"
  )
  assign(paste0(i),
         read.delim(
           filepath,
           header = TRUE,
           sep = "\t",
           stringsAsFactors = FALSE
         ))
  # increment
  coef <- coef + 1
}
```

# DEGs by FDRq <= 0.05
```{r}
# sex as a covariate
LBDvsControl_up <- subset(LBDvsControl, logFC > 0.25 & adj.P.Val <= 0.05)
LBDvsControl_down <- subset(LBDvsControl, logFC < -0.25 & adj.P.Val <= 0.05)
ADvsControl_up <- subset(ADvsControl, logFC > 0.25 & adj.P.Val <= 0.05)
ADvsControl_down <- subset(ADvsControl, logFC < -0.25 & adj.P.Val <= 0.05)
PAvsControl_up <- subset(PAvsControl, logFC > 0.25 & adj.P.Val <= 0.05)
PAvsControl_down <- subset(PAvsControl, logFC < -0.25 & adj.P.Val <= 0.05)
```

# Upset of DEGs
### all pathologies compared to control 
```{r}
list_input <- list("LBD down-regulated" = LBDvsControl_down$gene_name,
                   "LBD up-regulated" = LBDvsControl_up$gene_name,
                   "AD down-regulated" = ADvsControl_down$gene_name,
                   "AD up-regulated" = ADvsControl_up$gene_name,
                   "PA down-regulated" = PAvsControl_down$gene_name,
                   "PA up-regulated" = PAvsControl_up$gene_name)
data <- fromList(list_input)

upset_gene <- upset(data, set_sizes = FALSE,
      c('LBD down-regulated',
        'AD down-regulated',
        'PA down-regulated',
        'LBD up-regulated',
        'AD up-regulated',
        'PA up-regulated'),
      themes=upset_modify_themes(
        list('intersections_matrix'=theme(text=element_text(size=10),
              plot.margin = margin(0, 0, 0, 0, "cm")),
          'overall_sizes' = theme(axis.text.x = element_text(size =10), 
              plot.margin = margin(0, 0, 0, 0, "cm")))),
  queries=list(
    upset_query(set='LBD down-regulated', fill='blue'),
    upset_query(set='AD down-regulated', fill='blue'),
    upset_query(set='PA down-regulated', fill='blue'),
    upset_query(set='LBD up-regulated', fill='red'),
    upset_query(set='AD up-regulated', fill='red'),
    upset_query(set='PA up-regulated', fill='red')
  ),
  intersections = list(c('LBD down-regulated','AD down-regulated', 'PA down-regulated'), 
                       c('LBD down-regulated', 'AD down-regulated'), 
                       c('LBD down-regulated', 'PA down-regulated'), 
                       c('AD down-regulated', 'PA down-regulated'), 
                       'LBD down-regulated',
                       'AD down-regulated', 
                       'PA down-regulated',
                       c('LBD up-regulated','AD up-regulated', 'PA up-regulated'), 
                       c('LBD up-regulated', 'AD up-regulated'), 
                       c('LBD up-regulated', 'PA up-regulated'), 
                       c('AD up-regulated', 'PA up-regulated'), 
                       'LBD up-regulated',
                       'AD up-regulated', 
                       'PA up-regulated',
                       c('AD up-regulated', 'PA down-regulated'), 
                       c('AD down-regulated', 'PA up-regulated')
  ),
  base_annotations=list(
    'Intersection size'=(
        intersection_size(
        size=2,
        text = list(size = 3),
        text_mapping = aes(),
        bar_number_threshold=3,  # show all numbers on top of bars
        width=0.3,   # reduce width of the bars
        mapping=aes(fill='bars_color')
      )
      + scale_fill_manual(values=c('grey'), guide='none')
      + scale_y_continuous(expand=expansion(mult=c(0, 0.5)))
      + theme(axis.text.y = element_text(size = 10),
              axis.title.y = element_text(size = 10),
          panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),
          plot.margin = margin(0, 0, 0, 0, "cm"), 
          axis.line=element_line(colour='black')))),
  matrix=intersection_matrix(
      geom=geom_point(
        shape='circle filled',
        size=2,
        stroke=0.45)) +
        theme(axis.text.y = element_text(size = 10),
              axis.title.y = element_text(size = 10), 
              plot.margin = margin(0, 0, 0, 0, "cm")),
    sort_sets=FALSE,
  sort_intersections=FALSE)
upset_gene

# save
path <- paste0("../../results/", tool, 
               "/upset/", condition, "_all_types_FDRq0.05_logFC_0.25")
saveToPDF(paste0(path, ".pdf"), width = 6, height = 4)

upset(data, intersect = c('LBD down-regulated',
        'AD down-regulated',
        'PA down-regulated',
        'LBD up-regulated',
        'AD up-regulated',
        'PA up-regulated'))

path <- paste0("../../results/", tool,
               "/upset/", condition, "_all_types_FDRq0.05_logFC_0.25_simple")
saveToPDF(paste0(path, ".pdf"), width = 9, height = 6)

# Binary table with colnames:
write.table(
  data,
  paste0("../../results/",
  tool,
  "/upset/", condition, 
  "_all_types_FDRq0.05_logFC_0.25_list_binary_results.txt"),
  sep = "\t",
  quote = FALSE
)

```
# Upset of DEGs LBDvsControl & ADvsControl 
```{r}
list_input <- list("LBD down-regulated" = LBDvsControl_down$gene_name,
                   "LBD up-regulated" = LBDvsControl_up$gene_name,
                   "AD down-regulated" = ADvsControl_down$gene_name,
                   "AD up-regulated" = ADvsControl_up$gene_name)
data <- fromList(list_input)

upset_gene <- upset(data, set_sizes = FALSE,
      c('LBD down-regulated',
        'AD down-regulated',
        'LBD up-regulated',
        'AD up-regulated'),
      themes=upset_modify_themes(
        list('intersections_matrix'=theme(text=element_text(size=10),
              plot.margin = margin(0, 0, 0, 0, "cm")),
          'overall_sizes' = theme(axis.text.x = element_text(size =10), 
              plot.margin = margin(0, 0, 0, 0, "cm")))),
  queries=list(
    upset_query(set='LBD down-regulated', fill='blue'),
    upset_query(set='AD down-regulated', fill='blue'),
    upset_query(set='LBD up-regulated', fill='red'),
    upset_query(set='AD up-regulated', fill='red')
  ),
  intersections = list(c('LBD down-regulated', 'AD down-regulated'), 
                       'LBD down-regulated',
                       'AD down-regulated', 
                       c('LBD up-regulated', 'AD up-regulated'), 
                       'LBD up-regulated',
                       'AD up-regulated', 
                       c('LBD up-regulated', 'AD down-regulated'),
                       c('AD up-regulated', 'LBD down-regulated')

  ),
  base_annotations=list(
    'Intersection size'=(
        intersection_size(
        size=2,
        text = list(size = 3),
        text_mapping = aes(),
        bar_number_threshold=3,  # show all numbers on top of bars
        width=0.3,   # reduce width of the bars
        mapping=aes(fill='bars_color')
      )
      + scale_fill_manual(values=c('grey'), guide='none')
      + scale_y_continuous(expand=expansion(mult=c(0, 0.5)))
      + theme(axis.text.y = element_text(size = 10),
              axis.title.y = element_text(size = 10),
          panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),
          plot.margin = margin(0, 0, 0, 0, "cm"), 
          axis.line=element_line(colour='black')))),
  matrix=intersection_matrix(
      geom=geom_point(
        shape='circle filled',
        size=2,
        stroke=0.45)) +
        theme(axis.text.y = element_text(size = 10),
              axis.title.y = element_text(size = 10), 
              plot.margin = margin(0, 0, 0, 0, "cm")),
    sort_sets=FALSE,
  sort_intersections=FALSE)
upset_gene

# save
path <- paste0("../../results/", tool, 
               "/upset/", condition, "_LBD_AD_FDRq0.05_logFC_0.25")
saveToPDF(paste0(path, ".pdf"), width = 6, height = 4)

upset(data, intersect = c('LBD down-regulated',
        'AD down-regulated',
        'LBD up-regulated',
        'AD up-regulated'))

path <- paste0("../../results/", tool,
               "/upset/", condition, "_LBD_AD_FDRq0.05_logFC_0.25_simple")
saveToPDF(paste0(path, ".pdf"), width = 9, height = 6)

# Binary table with colnames:
write.table(
  data,
  paste0("../../results/",
  tool,
  "/upset/", condition, 
  "_LBD_AD_FDRq0.05_logFC_0.25_list_binary_results.txt"),
  sep = "\t",
  quote = FALSE
)

```

# Correlation plot
### LBD vs AD 
```{r}
ADandLBD <-
  merge(ADvsControl,
        LBDvsControl,
        by = "gene_name",
        all = T)
ADandLBD <- ADandLBD[c(1,2,6,7,8:15,22:29)]
colnames(ADandLBD) = c("gene_name", "chr", "gene_id","gene_type", "AD_logFC", "AD_CI.L", "AD_CI.R", "AD_AveExpr", "AD_t","AD_P.val", "AD_adj.P.Val", "AD_B", "LBD_logFC", "LBD_CI.L", "LBD_CI.R", "LBD_AveExpr", "LBD_t","LBD_P.val", "LBD_adj.P.Val", "LBD_B")

ADandLBD_goi <- subset(ADandLBD, 
                        (AD_logFC < -0.5 & LBD_logFC < -0.5) | # down in both
                          (AD_logFC > .5 & LBD_logFC > 0.5) | # up in both
                          (AD_logFC > .5 & LBD_logFC < 0) | # up in AD, down in LBD
                          (AD_logFC < -.5 & LBD_logFC > 0)) # up in LBD, down in AD

ADandLBD_corr_plot <- ggplot(data = ADandLBD, 
  aes(x = AD_logFC, y = LBD_logFC, text = paste(gene_name))) +
  annotate(
    "rect",
    xmin = 0,
    xmax = 2.75,
    ymin = 2.75,
    ymax = 0,
    fill = "lightpink3",
    alpha = .5
  ) +  annotate(
    "rect",
    xmin = 0,
    xmax = -2.25,
    ymin = 0,
    ymax = -2.25,
    fill = "cadetblue3",
    alpha = .5) +
  geom_abline(color = "gray40") +
  geom_text_repel(
    data = ADandLBD_goi, 
    aes(
      x = AD_logFC, 
      y = LBD_logFC,
      label = gene_name
    ),
    color = "black",
    fontface = "italic",
    size = 2.5, 
    max.overlaps = getOption("ggrepel.max.overlaps", default = 10)
  ) +
  geom_point(size = 1) + 
  theme_bw() +
  theme(plot.title = element_text(size = 10)) +
  theme(legend.position = "none") +
  theme(axis.title.x = element_text(size = 10),
        axis.text.x = element_text(size = 10)) +
  theme(axis.title.y = element_text(size = 10),
        axis.text.y = element_text(size = 10)) +
  labs(
    title = "LBD and AD",
    x = expression(paste("  ", log[2](AD/control))),
    y = expression(paste("  ", log[2](LBD/control))))+
  scale_y_continuous(breaks = seq(-1.75, 2.725, by = .5), 
                     limits = c(-2.25, 2.75), expand = c(0,0)) +
  scale_x_continuous(breaks = seq(-2, 2.75, by = .5), 
                     limits = c(-2.25, 2.75), expand = c(0,0))
ADandLBD_corr_plot

path <-
  paste0(
    "../../results/",
    tool,
    "/correlation/",
    condition, 
    "_ADandLBD"
  )
saveToPDF(paste0(path, ".pdf"), width = 6.25, height = 4.5)

# Scatter plot with correlation coefficient
sp <- ggscatter(ADandLBD, x = "AD_logFC", y = "LBD_logFC",
                add = "reg.line",  # Add regression line
                add.params = list(color = "blue", fill = "lightgray"), 
                conf.int = TRUE)
# Add correlation coefficient
sp + stat_cor(method = "pearson", label.x = -1.25, label.y = 1)
path <-
  paste0(
    "../../results/",
    tool,
    "/correlation/",
    condition, 
    "_ADandLBD_corr_pval"
  )
saveToPDF(paste0(path, ".pdf"), width = 6.25, height = 4.5)
```
## LBD vs AD significant DEGs only 
```{r}
ADandLBD <-
  merge(ADvsControl,
        LBDvsControl,
        by = "gene_name",
        all = T)
ADandLBD <- ADandLBD[c(1,2,6,7,8:15,22:29)]
colnames(ADandLBD) = c("gene_name", "chr", "gene_id","gene_type", "AD_logFC", "AD_CI.L", "AD_CI.R", "AD_AveExpr", "AD_t","AD_P.val", "AD_adj.P.Val", "AD_B", "LBD_logFC", "LBD_CI.L", "LBD_CI.R", "LBD_AveExpr", "LBD_t","LBD_P.val", "LBD_adj.P.Val", "LBD_B")

write.table(
  ADandLBD,
  paste0(
    "../../results/",
    tool,
    "/upset/",
    "ADvsControl_and_LBDvsControl_DEGs_FDRq0.05_logFC_0.25.txt"
  ),
  sep = "\t",
  quote = FALSE,
  row.names = FALSE
)

# significant in either 
ADandLBD <- subset(ADandLBD, (AD_logFC < -.25 & AD_adj.P.Val < 0.05) | (AD_logFC > 0.25 & AD_adj.P.Val < 0.05) |
                     (LBD_logFC < -.25 & LBD_adj.P.Val < 0.05) | (LBD_logFC > 0.25 & LBD_adj.P.Val < 0.05) )
ADandLBD_goi <- subset(ADandLBD, 
                        (AD_logFC < -0.5 & LBD_logFC < -0.5) | # down in both
                          (AD_logFC > .5 & LBD_logFC > 0.5) | # up in both
                          (AD_logFC > .5 & LBD_logFC < 0) | # up in AD, down in LBD
                          (AD_logFC < -.5 & LBD_logFC > 0)) # up in LBD, down in AD

ADandLBD_corr_plot <- ggplot(data = ADandLBD, 
  aes(x = AD_logFC, y = LBD_logFC, text = paste(gene_name))) +
  annotate(
    "rect",
    xmin = 0,
    xmax = 2.75,
    ymin = 2.75,
    ymax = 0,
    fill = "lightpink3",
    alpha = .5
  ) +  annotate(
    "rect",
    xmin = 0,
    xmax = -2.25,
    ymin = 0,
    ymax = -2.25,
    fill = "cadetblue3",
    alpha = .5) +
  geom_abline(color = "gray40") +
  geom_text_repel(
    data = ADandLBD_goi, 
    aes(
      x = AD_logFC, 
      y = LBD_logFC,
      label = gene_name
    ),
    color = "black",
    fontface = "italic",
    size = 2.5, 
    max.overlaps = getOption("ggrepel.max.overlaps", default = 10)
  ) +
  geom_point(size = 1) + 
  theme_bw() +
  theme(plot.title = element_text(size = 10)) +
  theme(legend.position = "none") +
  theme(axis.title.x = element_text(size = 10),
        axis.text.x = element_text(size = 10)) +
  theme(axis.title.y = element_text(size = 10),
        axis.text.y = element_text(size = 10)) +
  labs(
    title = "DEGs shared and unique between LBD and AD",
    x = expression(paste("  ", log[2](AD/control))),
    y = expression(paste("  ", log[2](LBD/control))))+
  scale_y_continuous(breaks = seq(-1.75, 2.725, by = .5), 
                     limits = c(-2.25, 2.75), expand = c(0,0)) +
  scale_x_continuous(breaks = seq(-2, 2.75, by = .5), 
                     limits = c(-2.25, 2.75), expand = c(0,0))
ADandLBD_corr_plot

path <-
  paste0(
    "../../results/",
    tool,
    "/correlation/",
    condition, 
    "_ADandLBD_sig_genes_only"
  )
saveToPDF(paste0(path, ".pdf"), width = 6.25, height = 4.5)

# Scatter plot with correlation coefficient
sp <- ggscatter(ADandLBD, x = "AD_logFC", y = "LBD_logFC",
                add = "reg.line",  # Add regression line
                add.params = list(color = "blue", fill = "lightgray"), 
                conf.int = TRUE)
# Add correlation coefficient
sp + stat_cor(method = "pearson", label.x = -1.25, label.y = 1)
path <-
  paste0(
    "../../results/",
    tool,
    "/correlation/",
    condition, 
    "_ADandLBD_corr_pval_sig_genes_only"
  )
saveToPDF(paste0(path, ".pdf"), width = 6.25, height = 4.5)
```
```{r}
```
