---
title: "Glimma"
author: "Kimberly Olney"
date: "2023-05-10"
output:
  pdf_document: default
  html_document: default
---

```{r setup, echo = FALSE}
knitr::opts_knit$set(root.dir = ".")
library(rmarkdown)
source(here::here("/research/labs/neurology/fryer/m239830/LBD_CWOW/scripts/R", "file_paths_and_colours.R"))
```

```{r data, echo = FALSE}
condition <- c("TYPE")
tool <- c("star")
dge.filtered.norm <- readRDS(paste0("../../rObjects/dge.filtered.norm.rds"))
# some samples are missing RIN values. 
# Replace NA with median RIN. 
# This is necessary to be able include RIN as a covariate in voom
# fill missing values with median
dge.filtered.norm$samples$RIN <- impute(dge.filtered.norm$samples$RIN, median)
# one sample is missing VaD information
dge.filtered.norm$samples$VaD <- impute(dge.filtered.norm$samples$VaD, median)
dge.filtered.norm$samples$flowcell_and_lane <- factor(dge.filtered.norm$samples$flowcell_and_lane)
dge.filtered.norm$samples$APOE <- factor(dge.filtered.norm$samples$APOE)
info <- as.data.frame(dge.filtered.norm$samples)

genes <- dge.filtered.norm$genes # gene information 
lcpm <- edgeR::cpm(dge.filtered.norm$counts, log = TRUE) # obtain log CPM counts

biomarkers <- c("ENO2", "GFAP", "OLIG2", "CD34", "P2RY12")

for (i in biomarkers) {
  biomarker <- subset(genes, gene_name == i) # gene of interest 
  biomarker_counts <- subset(lcpm, rownames(lcpm) %in% biomarker)
  biomarker_melt <- reshape2::melt(biomarker_counts) # reshape data 
  # rename columns to merge with metadata 
  names(biomarker_melt)[names(biomarker_melt) == "value"] <- i 
  # rename columns to merge with metadata 
  names(biomarker_melt)[names(biomarker_melt) == "Var2"] <- "NPID" 
  names(biomarker_melt)[names(biomarker_melt) == "Var1"] <- "gene_id"
  biomarker_melt$gene_id <- NULL
  assign(paste0(i),biomarker_melt)
}
# add gene expression values into one dataframe 
# put all data frames into list
df_list <- list(ENO2, OLIG2, CD34, P2RY12, GFAP)
# merge all data frames in list
cell_biomarker_lcpm <- df_list %>% reduce(full_join, by='NPID')

df <- merge(cell_biomarker_lcpm, info, by = "NPID")
scaled.info <-
  df[c(
    "Race_numeric",
    "RIN",
    "Age",
    "PCT_CODING_BASES",
    "PCT_INTERGENIC_BASES",
    "PCT_INTRONIC_BASES",
    "APOE_E4_allele_count", 
    "ENO2", 
    "GFAP", 
    "OLIG2", 
    "CD34", 
    "P2RY12",
    "Brain.wt"
  )] %>% scale()
scaled.info.df <- as.data.frame(scaled.info)
# remove columns with unscaled data 
df <- (subset(df, select = -c(Race_numeric, RIN, Age, PCT_CODING_BASES, PCT_INTRONIC_BASES, APOE_E4_allele_count, ENO2, GFAP, OLIG2, CD34, P2RY12, Brain.wt)))
# Add scaled information to the metadata called "info"
info_with_scale <- cbind(df, scaled.info.df)

# replace sample information with the updated info that includes biomakrer expression
dge.filtered.norm$samples <- info_with_scale

veBayesFit <- readRDS(paste0("../../rObjects/", condition, ".veBayesFit.rds"))
```
## MDS 
The multidimensional scaling (MDS) plot to explore differences in samples. The first two dimensions explain the greatest variance between samples, and the amount of variance decreases monotonically with increasing dimension.

Interactions with the plot:
Scaling the points by library size (lib_size) using the scale_by field.
Changing the colour of points by group using the colour_by field.
Altering the shape of points by sample batch using the shape_by field.
```{r}
Sys.setenv(RSTUDIO_PANDOC="/usr/local/biotools/pandoc/3.1.2/bin/pandoc")
glimmaMDS(dge.filtered.norm)
glimmaVolcano(veBayesFit, dge = dge.filtered.norm)
```
