---
title: "gprofiler2"
author: "Kimberly Olney"
date: "06/09/2022"
output: html_document
---
# gprofiler for gene enrichment analysis
```{r setup, include=FALSE, echo=FALSE}
knitr::opts_knit$set(root.dir = ".")
```
```{r, echo=FALSE}
saveToPDF <- function(...) {
    d = dev.copy(pdf,...)
    dev.off(d)
}

saveToPNG <- function(...) {
    d = dev.copy(png,...)
    dev.off(d)
}
```

```{r packages, echo=FALSE, message=FALSE}
library(gprofiler2)
library(DESeq2)
library(airway)
library(enrichplot)
library(DOSE) 
library(plyr)
library(ggplot2)
require(gridExtra)
library(ggpubr)
library(scales)
library(forcats)
```

```{r set_variables, echo=FALSE}
tissue <- c("Brain") # Kidney or Brain
control <- "Control"
treatment <- "DLB"
control_color <- "gray29"
treatment_color <- "red"
tool = c("star")
```

# Read in DEGs
The DEGs are already ordered by smallest to largest adjusted p-value.
```{r}
star_all  <- read.delim("../../results/star/DEGs/LBD_gene_DEGs_FDRq0.05.txt", 
                    header = TRUE, sep = "\t")
star_XX <- read.delim("../../results/star/DEGs/LBD_gene_DEGs_FDRq0.05_XX.txt", 
                    header = TRUE, sep = "\t")
star_XY <- read.delim("../../results/star/DEGs/LBD_gene_DEGs_FDRq0.05_XY.txt", 
                     header = TRUE, sep = "\t")
```

# Subset lists
sort = FALSE so that the order of the DEGs remains. The DEGs are ordered by adjusted p-value
May also order by log2FC
```{r}
# sort by fold change. Biggest fold change is the first gene
star_all_sort_positiveFC <- star_XY[order(-star_all$adj.P.Val),]
# now repeat for negative fold changes
star_all_sort_negativeFC <- star_XY[order(star_all$adj.P.Val),]

star_XX_sort_positiveFC <- star_XX[order(-star_XX$adj.P.Val),]
star_XX_sort_negativeFC <- star_XX[order(star_XX$adj.P.Val),]

star_XY_sort_positiveFC <- star_XY[order(-star_XY$adj.P.Val),]
star_XY_sort_negativeFC <- star_XY[order(star_XY$adj.P.Val),]

# when subsetting, keep the order. I.e sort = FALSE. Do not resort 
up_star_all <- subset(star_all_sort_positiveFC$gene_name, 
                      star_all_sort_positiveFC$logFC > 0, sort = FALSE)
down_star_all <- subset(star_all_sort_negativeFC$gene_name, 
                        star_all_sort_negativeFC$logFC < 0, sort = FALSE)

up_star_XX <- subset(star_XX_sort_positiveFC$gene_name, 
                     star_XX_sort_positiveFC$logFC > 0, sort = FALSE)
down_star_XX <- subset(star_XX_sort_negativeFC$gene_name, 
                       star_XX_sort_negativeFC$logFC < 0, sort = FALSE)

up_star_XY <- subset(star_XY_sort_positiveFC$gene_name, 
                     star_XY_sort_positiveFC$logFC > 0, sort = FALSE)
down_star_XY <- subset(star_XY_sort_negativeFC$gene_name, 
                       star_XY_sort_negativeFC$logFC < 0, sort = FALSE)
```
# All samples (XX and XY)
### Enrichment analysis
ordered_query = TRUE specify that the genes are in a meaningful order. 
Ordered by adjusted p-value. 
```{r}
gp_up <- gost(up_star_all, ordered_query = TRUE, organism = "hsapiens")
gp_down <- gost(down_star_all, ordered_query = TRUE, organism = "hsapiens")

# inspect
head(gp_up$result)
head(gp_down$result)
```

```{r, echo=FALSE}
# up-regulated 
gostplot(gp_up, interactive = TRUE) # , capped = FALSE
p_up = gostplot(gp_up, interactive = FALSE)
publish_gostplot(p_up) #, highlight_terms = c("GO:0002376", "GO:0006952","GO:0034097", "KEGG:05169", "KEGG04668"))

path <- paste0("../../results/", tool, "/gprofiler/", treatment, "_",tolower(tissue),"_up_gprofiler")
saveToPDF(paste0(path, ".pdf"), width = 8, height = 6)
saveToPNG(paste0(path, ".png"), width = 8, height = 6, unit = "in", res = 300)

# down-regulated 
gostplot(gp_down, interactive = TRUE) # , capped = FALSE
p_down = gostplot(gp_down, interactive = FALSE)
publish_gostplot(p_down) #, highlight_terms = c("GO:0035633"))
```

```{r eval =FALSE, echo=FALSE}
path <- paste0("../../results/", tool, "/gprofiler/", treatment, "_",tolower(tissue),"_down_gprofiler")
saveToPDF(paste0(path, ".pdf"), width = 8, height = 6)
saveToPNG(paste0(path, ".png"), width = 8, height = 6, unit = "in", res = 300)
```
### Plot up and down enrichment together 
```{r}
multi_gp = gost(list("down-regulated" = down_star_all, 
                     "up-regulated" = up_star_all), ordered_query = TRUE, 
                     organism = "hsapiens")

# rearrange so that up-regualted is on top 
neworder <- c("up-regulated","down-regulated")
multi_gp$result <- arrange(transform(multi_gp$result,
             query=factor(query,levels=neworder)),query)

# interactive plot 
gostplot(multi_gp, interactive = TRUE)
```

```{r, echo=FALSE}
# needed to convert to enrichResult object
up_names = gconvert(up_star_all, organism = "hsapiens") 
down_names = gconvert(down_star_all,organism = "hsapiens")

# enrichment analysis using gene names
multi_gp = gost(list("up-regulated" = up_names$name, "down-regulated" = down_names$name), multi_query = FALSE, evcodes = TRUE, ordered_query = TRUE, 
                organism = "hsapiens")

#subset(multi_gp$result, term_id == "GO:0035633")
# reorder so that up-regulated is on top
neworder <- c("up-regulated","down-regulated")
multi_gp$result <- arrange(transform(multi_gp$result,
             query=factor(query,levels=neworder)),query)

# modify the g:Profiler data frame
gp_mod = multi_gp$result[,c("query", "source", "term_id", "term_name", "p_value", 
                            "query_size", "intersection_size","term_size", 
                            "effective_domain_size","intersection")] 
# get the gene name size for each GO 
gp_mod$GeneRatio = paste0(gp_mod$intersection_size, "/", gp_mod$query_size)
gp_mod$BgRatio = paste0(gp_mod$term_size, "/", gp_mod$effective_domain_size) 
names(gp_mod) = c("Cluster", "Category", "ID", "Description", "p.adjust",
"query_size", "Count", "term_size", "effective_domain_size",
"geneID", "GeneRatio", "BgRatio") 
gp_mod$geneID = gsub(",", "/", gp_mod$geneID)
row.names(gp_mod) = make.names(gp_mod$ID, unique = TRUE)

# define as compareClusterResult object
gp_mod_cluster = new("compareClusterResult", compareClusterResult = gp_mod)
# define as enrichResult object
gp_mod_enrich = new("enrichResult", result = gp_mod)

# output table 
write.table(gp_mod_enrich@result, paste0("../../results/", tool, "/gprofiler/", treatment, "_",tolower(tissue),"_up_and_down_gp_mod_enrich.txt"), sep = "\t", quote = FALSE, row.names = FALSE)
```
# XX only
### Enrichment analysis
```{r}
#gp_up <- gost(up_star_XX, ordered_query = TRUE, organism = "hsapiens")
gp_down <- gost(down_star_XX, ordered_query = TRUE, organism = "hsapiens")

# inspect
#head(gp_up$result)
head(gp_down$result)
```

### Plot enrichment analysis 
interactive = TRUE means the plot is interactive. You can select a point of interest to see what it is. \
interactive = FALSE is needed when saving the plot
capped = FALSE indicates if the Y-asix p-value should be capped at 16 or not. 
```{r, echo=FALSE}
# up-regulated 
#gostplot(gp_up, interactive = TRUE) # , capped = FALSE
#p_up = gostplot(gp_up, interactive = FALSE)
#publish_gostplot(p_up) #, highlight_terms = c("GO:0002376", "GO:0006952","GO:0034097", "KEGG:05169", "KEGG04668"))

# down-regulated 
gostplot(gp_down, interactive = TRUE) # , capped = FALSE
p_down = gostplot(gp_down, interactive = FALSE)
publish_gostplot(p_down) #, highlight_terms = c("GO:0035633"))
```

### Plot up and down enrichment together 
```{r, eval=FALSE, echo=FALSE}
multi_gp = gost(list("down-regulated" = down_star_XX, 
                     "up-regulated" = up_star_XX), ordered_query = TRUE, 
                     organism = "hsapiens")

# rearrange so that up-regualted is on top 
neworder <- c("up-regulated","down-regulated")
multi_gp$result <- arrange(transform(multi_gp$result,
             query=factor(query,levels=neworder)),query)

# interactive plot 
gostplot(multi_gp, interactive = TRUE)
```

```{r echo=FALSE}
# needed to convert to enrichResult object
up_names = gconvert(up_star_XX, organism = "hsapiens") 
down_names = gconvert(down_star_XX,organism = "hsapiens")

# enrichment analysis using gene names
multi_gp = gost(list("up-regulated" = up_names$name, "down-regulated" = down_names$name), multi_query = FALSE, evcodes = TRUE, ordered_query = TRUE, 
                organism = "hsapiens")

# reorder so that up-regulated is on top
neworder <- c("up-regulated","down-regulated")
multi_gp$result <- arrange(transform(multi_gp$result,
             query=factor(query,levels=neworder)),query)

# modify the g:Profiler data frame
gp_mod = multi_gp$result[,c("query", "source", "term_id", "term_name", "p_value", 
                            "query_size", "intersection_size","term_size", 
                            "effective_domain_size","intersection")] 
# get the gene name size for each GO 
gp_mod$GeneRatio = paste0(gp_mod$intersection_size, "/", gp_mod$query_size)
gp_mod$BgRatio = paste0(gp_mod$term_size, "/", gp_mod$effective_domain_size) 
names(gp_mod) = c("Cluster", "Category", "ID", "Description", "p.adjust",
"query_size", "Count", "term_size", "effective_domain_size",
"geneID", "GeneRatio", "BgRatio") 
gp_mod$geneID = gsub(",", "/", gp_mod$geneID)
row.names(gp_mod) = make.names(gp_mod$ID, unique = TRUE)

# define as compareClusterResult object
gp_mod_cluster = new("compareClusterResult", compareClusterResult = gp_mod)
# define as enrichResult object
gp_mod_enrich = new("enrichResult", result = gp_mod)

# output table 
write.table(gp_mod_enrich@result, paste0("../../results/", tool, "/gprofiler/", treatment, "_",tolower(tissue),"_up_and_down_gp_mod_enrich_XX.txt"), sep = "\t", quote = FALSE, row.names = FALSE)
```
# XY only
### Enrichment analysis
```{r}
gp_up <- gost(up_star_XY, ordered_query = TRUE, organism = "hsapiens")
gp_down <- gost(down_star_XY, ordered_query = TRUE, organism = "hsapiens")

# inspect
head(gp_up$result)
head(gp_down$result)
```

### Plot enrichment analysis 
```{r echo=FALSE}
# up-regulated 
gostplot(gp_up, interactive = TRUE) # , capped = FALSE
p_up = gostplot(gp_up, interactive = FALSE)
publish_gostplot(p_up)

path <- paste0("../../results/", tool, "/gprofiler/", treatment, "_",tolower(tissue),"_up_gprofiler_XY")
saveToPDF(paste0(path, ".pdf"), width = 8, height = 6)
saveToPNG(paste0(path, ".png"), width = 8, height = 6, unit = "in", res = 300)

# down-regulated 
gostplot(gp_down, interactive = TRUE) # , capped = FALSE
p_down = gostplot(gp_down, interactive = FALSE)
publish_gostplot(p_down) #, highlight_terms = c("GO:0035633"))
```

```{r eval=FALSE, echo=FALSE}
path <- paste0("../../results/", tool, "/gprofiler/", treatment, "_",tolower(tissue),"_down_gprofiler_XY")
saveToPDF(paste0(path, ".pdf"), width = 8, height = 6)
saveToPNG(paste0(path, ".png"), width = 8, height = 6, unit = "in", res = 300)
```
### Plot up and down enrichment together 
```{r}
multi_gp = gost(list("down-regulated" = down_star_XY, 
                     "up-regulated" = up_star_XY), ordered_query = TRUE, 
                     organism = "hsapiens")

# rearrange so that up-regualted is on top 
neworder <- c("up-regulated","down-regulated")
multi_gp$result <- arrange(transform(multi_gp$result,
             query=factor(query,levels=neworder)),query)

# interactive plot 
gostplot(multi_gp, interactive = TRUE)
```

```{r echo=FALSE}
# needed to convert to enrichResult object
up_names = gconvert(up_star_XY, organism = "hsapiens") 
down_names = gconvert(down_star_XY,organism = "hsapiens")

# enrichment analysis using gene names
multi_gp = gost(list("up-regulated" = up_names$name, "down-regulated" = down_names$name), multi_query = FALSE, evcodes = TRUE, ordered_query = TRUE, 
                organism = "hsapiens")

# reorder so that up-regulated is on top
neworder <- c("up-regulated","down-regulated")
multi_gp$result <- arrange(transform(multi_gp$result,
             query=factor(query,levels=neworder)),query)

# modify the g:Profiler data frame
gp_mod = multi_gp$result[,c("query", "source", "term_id", "term_name", "p_value", 
                            "query_size", "intersection_size","term_size", 
                            "effective_domain_size","intersection")] 
# get the gene name size for each GO 
gp_mod$GeneRatio = paste0(gp_mod$intersection_size, "/", gp_mod$query_size)
gp_mod$BgRatio = paste0(gp_mod$term_size, "/", gp_mod$effective_domain_size) 
names(gp_mod) = c("Cluster", "Category", "ID", "Description", "p.adjust",
"query_size", "Count", "term_size", "effective_domain_size",
"geneID", "GeneRatio", "BgRatio") 
gp_mod$geneID = gsub(",", "/", gp_mod$geneID)
row.names(gp_mod) = make.names(gp_mod$ID, unique = TRUE)

# define as compareClusterResult object
gp_mod_cluster = new("compareClusterResult", compareClusterResult = gp_mod)
# define as enrichResult object
gp_mod_enrich = new("enrichResult", result = gp_mod)

# output table 
write.table(gp_mod_enrich@result, paste0("../../results/", tool, "/gprofiler/", treatment, "_",tolower(tissue),"_up_and_down_gp_mod_enrich_XY.txt"), sep = "\t", quote = FALSE, row.names = FALSE)
```

