---
title: "Examine clinical stats for representative samples"
author: "Kimberly Olney, Ph.D"
date: "05/11/2023"
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: inline
---
This script will examine the variables in the metadata to determine if there are differences by disease type and/or sex differences for clinical data such as brain weight. 
Repersentative samples were selected following the outlier_and_representative_sample.Rmd script 
# Setup
```{r setup, message=FALSE, warning=FALSE, tidy=TRUE}
knitr::opts_knit$set(root.dir = ".")
```
# User defined variables
sourcing file_paths_and_colours will load the libraries and user defined variables for colors of the graphs 
```{r set_variables, message=FALSE, warning=FALSE, tidy=TRUE}
source(here::here("scripts/R", "file_paths_and_colours.R"))
remove(metadata, metadata_ATS, df)
```
# Read data
```{r read_data, message=FALSE, warning=FALSE, tidy=TRUE}
# read in metadata
metadata <-
  read.delim(paste0("repersentative_samples.tsv"),
             header = TRUE,
             sep = "\t")
metadata$TYPE <-
  factor(metadata$TYPE, levels = c("CONTROL", "PA", "AD", "LBD"))
```

# Library size 
```{r library_size, message=FALSE, warning=FALSE, tidy=TRUE}
hist(metadata$TOTAL_READS, breaks = 100)
summary(metadata$TOTAL_READS)
```

## RNA sample count
```{r bar_plot, message=FALSE, warning=FALSE, tidy=TRUE}
# create bar plot showing number of samples for each group
bar <- ggplot(metadata, aes(TYPE, after_stat(count), fill = TYPE)) +
  geom_bar() +
  theme_bw() +
  xlab("Disease group") +
  ggtitle("Bulk RNAseq sample count by disease group") +
  geom_text(
    stat = 'count',
    aes(label = after_stat(count)),
    vjust = 1.6,
    color = "white",
    size = 3.5
  ) +
    theme_bw() + 
  scale_fill_manual(values = TypeColors)
bar

# pie chart
data <- metadata %>%
  group_by(TYPE) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(perc = `n` / sum(`n`)) %>%
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))

pie <- ggplot(data, aes(x = "", y = perc, fill = TYPE)) +
  geom_col(color = "black") +
  geom_label(
    aes(label = labels),
    color = c("white"),
    position = position_stack(vjust = 0.5),
    show.legend = FALSE
  ) +
  guides(fill = guide_legend(title = "Disease group")) +
  scale_fill_manual(values = TypeColors) +
  coord_polar(theta = "y") +
  theme_void() +
  theme(legend.position = "none")
pie
```
save plot
```{r message=FALSE, warning=FALSE, tidy=TRUE}
ggarrange(bar,
          pie,
          ncol = 2,
          widths = c(2.5, 1))

path <- paste0("../../results/clinical_representative_samples/TYPE.RNA.count")
saveToPDF(paste0(path, ".pdf"), width = 7.25, height = 3.5)

# clean up
remove(bar, pie, data)
```

## Split by sex 
```{r message=FALSE, warning=FALSE, tidy=TRUE}
#SexColors <- c("orange", "olivedrab")
bar_sex <- ggplot(metadata, aes(x = TYPE, fill = sex_inferred)) + 
  geom_bar(aes(y = after_stat(count)), position = "dodge") + 
  theme_bw() +
  xlab("Disease group") + 
  ggtitle("Bulk RNAseq sample count by inferred sex") +
  geom_text(stat='count', aes(x = TYPE, label=after_stat(count)), 
            position = position_dodge(width = 1), 
              vjust=-.25, 
            color="black", size=3.5) + 
  scale_fill_manual(values = SexColors) +
  guides(fill = guide_legend(title = "Sex inferred"))
bar_sex

data <- metadata %>% 
  group_by(sex_inferred,TYPE) %>%
  dplyr::count() %>% 
  ungroup(sex_inferred) %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))

pie_sex <- ggplot(data, aes(x = "", y = perc, fill = sex_inferred)) +
  geom_col(color = "black") +
  geom_label(aes(label = labels), color = c("white"),
            position = position_stack(vjust = 0.5),
            show.legend = FALSE) +
    facet_wrap(~TYPE, ncol = 4) +
  guides(fill = guide_legend(title = "Sex inferred")) +
  scale_fill_manual(values=SexColors) +
  coord_polar(theta = "y") + 
  theme_void() +
  theme(legend.position = "none") +
    theme(strip.text = element_blank())
pie_sex
```
```{r message=FALSE, warning=FALSE, tidy=TRUE}
ggarrange(bar_sex,
          pie_sex,
          nrow = 2,
          widths = c(2, 1), heights = c(2.8,1))

path <- paste0("../../results/clinical_representative_samples/TYPE.sex.RNA.count")
saveToPDF(paste0(path, ".pdf"), width = 6.88, height = 4.33)

# clean up
remove(bar_sex, pie_sex, data)
```

# RNA sample count by A-T-S scores
Group 1: A-T-S amyloid + synuclein + tau (A >= 1, T >= 1, S >= 1 )
Group 2: Ahigh-S high amyloid + synuclein (A >= 2, T = 0, S >= 1 )
Group 3: Alow-S low amyloid + synuclein  (A < 2 , T = 0, S >= 1 )
Group 4: S "pure" synuclein (A = 0, T = 0, S >= 1)
Group 5: A-T amyloid + tau (A >= 1, T >= 1, S = 0)
Group 6: A amyloid (A >= 1, T = 0, S = 0)
Group 7: Min minimal (i.e normal controls) (A = 0, T = 0, S = 0) 

plot by A-T-S score
```{r message=FALSE, warning=FALSE, tidy=TRUE}
# Samples with missing A-T-S scores
subset(metadata, ATS_names == "NA")
subset(metadata, is.na(metadata$ATS_names))
# Drop samples with NA ATS score
df <- metadata %>% drop_na(ATS_names)
df <- subset(df, ATS_names != "NA")
df$ATS_names <-
  factor(
    df$ATS_names,
    levels = c(
      "no pathology",
      "amyloid",
      "amyloid + tau",
      "pure synuclein",
      "low amyloid + synuclein",
      "high amyloid + synuclein",
      "amyloid + synuclein + tau"
    )
  )
metadata$ATS_names <-
  factor(
    metadata$ATS_names,
    levels = c(
      "no pathology",
      "amyloid",
      "amyloid + tau",
      "pure synuclein",
      "low amyloid + synuclein",
      "high amyloid + synuclein",
      "amyloid + synuclein + tau"
    )
  )
```
```{r message=FALSE, warning=FALSE, tidy=TRUE}
# create bar plot showing number of samples for each group
bar <- ggplot(df, aes(TYPE, after_stat(count), fill = ATS_names)) +
  geom_bar() +
  theme_bw() +
  xlab("Disease group") +
  ggtitle("Bulk RNAseq sample count by A-T-S scores") +
  theme_bw() + 
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust=1),
        axis.title.y = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        plot.title = element_text(size = 10),
        legend.text=element_text(size=10),
        legend.title=element_text(size=10)) +
  scale_fill_manual(values = colorbindColors)
bar
path <- paste0("../../results/clinical_representative_samples/ATS.RNA.count.ByTYPE")
saveToPDF(paste0(path, ".pdf"), width = 7, height = 5)

# create bar plot showing number of samples for each end point criteria
bar <- ggplot(df, aes(ATS_names, after_stat(count), fill = ATS_names)) +
  geom_bar() +
  theme_bw() +
  xlab("Disease group") +
  ggtitle("Bulk RNAseq sample count by A-T-S scores") +
  theme_bw() + 
  geom_text(
    stat = 'count',
    aes(label = after_stat(count)),
    vjust = 1.6,
    color = "black",
    size = 3.5
  ) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(size = 10, angle = 75, vjust = 1, hjust=1),
        axis.title.y = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        plot.title = element_text(size = 10),
        legend.text=element_text(size=10),
        legend.title=element_text(size=10)) +
  scale_fill_manual(values = colorbindColors)
bar
path <- paste0("../../results/clinical_representative_samples/ATS.RNA.count")
saveToPDF(paste0(path, ".pdf"), width = 7, height = 5)
```

```{r message=FALSE, warning=FALSE, tidy=TRUE}
# pie chart
data <- df %>%
  group_by(ATS_names) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(perc = `n` / sum(`n`)) %>%
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))

pie <- ggplot(data, aes(x = "", y = perc, fill = ATS_names)) +
  geom_col(color = "black") +
  geom_label(
    aes(label = labels),
    color = c("black"),
    position = position_stack(vjust = 0.5),
    show.legend = FALSE
  ) +
  guides(fill = guide_legend(title = "Disease group")) +
  scale_fill_manual(values = colorbindColors) +
  coord_polar(theta = "y") +
  theme_void() +
  theme(legend.position = "none")
pie
```
save plot
```{r message=FALSE, warning=FALSE, tidy=TRUE}
ggarrange(bar,
          pie,
          ncol = 2,
          widths = c(2.5, 1))

path <- paste0("../../results/clinical_representative_samples/ATS.RNA.count.pie")
saveToPDF(paste0(path, ".pdf"), width = 9, height = 6)

# clean up
remove(bar, data, pie) # do not remove df, used for split by sex 
```
## A-T-S split by sex 
```{r message=FALSE, warning=FALSE, tidy=TRUE}
bar_sex <- ggplot(df, aes(x = ATS_names, fill = sex_inferred)) + 
  geom_bar(aes(y = after_stat(count)), position = "dodge") + 
  theme_bw() + 
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust=1),
        axis.title.y = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        plot.title = element_text(size = 10),
        legend.text=element_text(size=10),
        legend.title=element_text(size=10)) +
  xlab("Disease group") + 
  ggtitle("Bulk RNAseq sample count by inferred sex") +
  geom_text(stat='count', aes(x = ATS_names, label=after_stat(count)), 
            position = position_dodge(width = 1), 
              vjust=-.25, 
            color="black", size=3.5) + 
  scale_fill_manual(values = SexColors) +
  guides(fill = guide_legend(title = "Sex inferred"))
bar_sex

data <- df %>% 
  group_by(sex_inferred,ATS_names) %>%
  dplyr::count() %>% 
  ungroup(sex_inferred) %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))

pie_sex <- ggplot(data, aes(x = "", y = perc, fill = sex_inferred)) +
  geom_col(color = "black") +
  geom_label(aes(label = labels), color = c("white"),
            position = position_stack(vjust = 0.5),
            show.legend = FALSE) +
    facet_wrap(~ATS_names, ncol = 6) +
  guides(fill = guide_legend(title = "Sex inferred")) +
  scale_fill_manual(values=SexColors) +
  coord_polar(theta = "y") + 
  theme_void() +
  theme(legend.position = "none") +
    theme(strip.text = element_blank())
pie_sex
```
```{r message=FALSE, warning=FALSE, tidy=TRUE}
ggarrange(bar_sex,
          pie_sex,
          nrow = 2, 
          heights = c(2,1))

path <- paste0("../../results/clinical_representative_samples/ATS.sex.RNA.count")
saveToPDF(paste0(path, ".pdf"), width = 8, height = 7)

remove(bar_sex, pie_sex, data, df)
```

# Categorial variables
### TYPE
```{r message=FALSE, warning=FALSE, tidy=TRUE}
metadata_categorical <-
  data.frame(
    metadata$LBD.type,
    metadata$AD.subtype,
    metadata$CDLB,
    metadata$TDP.type,
    metadata$APOE,
    metadata$MAPT,
    metadata$GRN,
    metadata$TMEM106b,
    metadata$Braak.NFT,
    metadata$Thal.amyloid,
    metadata$MF.Tau,
    metadata$MF.Amyloid,
    metadata$VaD,
    metadata$TDP.43,
    metadata$FHx,
    metadata$MF.SP,
    metadata$MF.NFT,
    metadata$MF.LB,
    metadata$Cing.LB,
    metadata$Cing.Synuclein,
    metadata$Race
  )
# To do it for all names
column_variables <-
  c(
    "LBD.type",
    "AD.subtype",
    "CDLB",
    "TDP.type",
    "APOE",
    "MAPT",
    "GRN",
    "TMEM106b",
    "Braak.NFT",
    "Thal.amyloid",
    "MF.Tau",
    "MF.Amyloid",
    "VaD",
    "TDP.43",
    "FHx",
    "MF.SP",
    "MF.NFT",
    "MF.LB",
    "Cing.LB",
    "Cing.Synuclein",
    "Race"
  )

bar_plot_fun <- function(i, j) {
  ggplot(metadata, aes(TYPE, after_stat(count), fill = factor(i))) +
    geom_bar(colour = "black") +
    theme_bw() +
    ggtitle(j) + 
    guides(fill = guide_legend(title = j)) +
    xlab("Disease group") +
    ylab("Count") +
    scale_fill_manual(values=colorbindColors) +
    geom_text(
      stat = 'count',
      aes(label = after_stat(count)),
      position = position_stack(vjust = .5),
      color = "black",
      size = 3.5
    )
}
bar_plots <-
  Map(bar_plot_fun, i = metadata_categorical, j = column_variables)
bar_plots

pie_plot_fun <- function(j) {
  data <- metadata %>%
    group_by(metadata[j], TYPE) %>%
    dplyr::count() %>%
    ungroup(all_of(j)) %>%
    mutate(perc = `n` / sum(`n`)) %>%
    arrange(perc) %>%
    mutate(labels = scales::percent(perc))
  k <- unlist(data[, c(1)])
  
  pie <- ggplot(data, aes(x = "", y = perc, fill = factor(k))) +
    geom_col(color = "black") +
    geom_label(
      aes(label = labels),
      color = c("black"),
      position = position_stack(vjust = 0.5),
      show.legend = FALSE
    ) +
    facet_wrap( ~ TYPE, ncol = 4) +
    guides(fill = guide_legend(title = j)) +
    coord_polar(theta = "y") +
    theme_void() +
    theme(legend.position = "none")+ 
    scale_fill_manual(values=colorbindColors)
}
pie_plots <- Map(pie_plot_fun, j = column_variables)

p <- list()
for (i in 1:length(column_variables)) {
  p[[i]] <- grid.arrange(grobs = c(bar_plots[i], pie_plots[i]), heights = c(2,1))
}

plotnames = imap(pie_plots, ~ paste0("../../results/clinical_representative_samples/TYPE.", .y, ".pdf")) %>%
  flatten()
plotnames
walk2(plotnames,
      p,
      ~ ggsave(
        filename = .x,
        plot = .y,
        height = 7,
        width = 6.5
      ))
```

```{r message=FALSE, warning=FALSE, tidy=TRUE}
# clean up
remove(bar_plots, pie_plots, plotnames, p) # do not remove metadata_categorical
```

# Continuous variables 
### TYPE
```{r message=FALSE, warning=FALSE, tidy=TRUE}
metadata_continuous <-
  data.frame(
    metadata$Brain.wt,
    metadata$Duration,
    metadata$Age,
    metadata$PMI,
    metadata$Concentration.ng.ul,
    metadata$Volume.ul,
    metadata$Total.RNA.ng,
    metadata$RIN
  )
column_variables <-
  c(
    "Brain.wt",
    "Duration",
    "Age",
    "PMI",
    "Concentration.ng.ul",
    "Volume.ul",
    "Total.RNA.ng",
    "RIN"
  )
TYPE_comparison <-
  list(
    c("CONTROL", "PA"),
    c("CONTROL", "AD"),
    c("CONTROL", "LBD"),
    c("PA", "AD"),
    c("PA", "LBD"),
    c("AD", "LBD")
  )

violin_plot_fun <- function(i, j) {
  ggplot(metadata, aes(TYPE, i, fill = TYPE)) +
    geom_violin() +
    geom_boxplot(width = 0.1, outlier.shape = NA) + 
    geom_jitter(shape=16, position=position_jitter(0.2)) +
    theme_bw() + 
    ggtitle(j) +
    xlab("Disease group") +
    ylab(j) +
    stat_compare_means(comparisons = TYPE_comparison) +
    scale_fill_manual(values=TypeColors) +
    theme(legend.position = "none")
}
violin_plots <-
  Map(violin_plot_fun, i = metadata_continuous, j = column_variables)
violin_plots

plotnames = imap(violin_plots, ~paste0("../../results/clinical_representative_samples/TYPE.", .y, ".pdf")) %>%
  flatten()
plotnames <- gsub("metadata.", "", plotnames)
walk2(plotnames, violin_plots, ~ggsave(filename = .x, plot = .y, 
                                    height = 4, width = 5.5))

remove(violin_plots) 
```

## TYPE.sex
```{r message=FALSE, warning=FALSE, tidy=TRUE}
violin_plot_fun <- function(i, j) {
  ggplot(metadata, aes(TYPE, i, fill = TYPE)) +
    geom_violin() +
    geom_boxplot(width = 0.1, outlier.shape = NA) + 
    geom_jitter(shape=16, position=position_jitter(0.2)) +
    theme_bw() + 
    ggtitle(j) +
    xlab("Disease group") +
    ylab(j) +
    stat_compare_means(comparisons = TYPE_comparison) +
    scale_fill_manual(values=TypeColors) +
    theme(legend.position = "none") +
    facet_grid(. ~ sex_inferred)
}
violin_plots <-
  Map(violin_plot_fun, i = metadata_continuous, j = column_variables)
violin_plots

plotnames = imap(violin_plots, ~paste0("../../results/clinical_representative_samples/TYPE.Sex.", .y, ".pdf")) %>%
  flatten()
plotnames <- gsub("metadata.", "", plotnames)
walk2(plotnames, violin_plots, ~ggsave(filename = .x, plot = .y, 
                                    height = 4, width = 7.5))
remove(violin_plots) 
```

# Alignment metrics
### TYPE
```{r message=FALSE, warning=FALSE, tidy=TRUE}
metrics_metadata <-
  data.frame(
    metadata$PCT_PF_READS,
    metadata$PCT_CODING_BASES,
    metadata$PCT_INTERGENIC_BASES,
    metadata$PCT_INTRONIC_BASES,
    metadata$TOTAL_READS
  )
column_variables <-
  c(
    "PCT_PF_READS",
    "PCT_CODING_BASES",
    "PCT_INTERGENIC_BASES",
    "PCT_INTRONIC_BASES",
    "TOTAL_READS"
  )

violin_plot_fun <- function(i, j) {
  ggplot(metadata, aes(TYPE, i, fill = TYPE)) +
    geom_violin() +
    geom_boxplot(width = 0.1, outlier.shape = NA) + 
    geom_jitter(shape=16, position=position_jitter(0.2)) +
    theme_bw() + 
    ggtitle(j) +
    xlab("Disease group") +
    ylab(j) +
    stat_compare_means(comparisons = TYPE_comparison) +
    scale_fill_manual(values=TypeColors) +
    theme(legend.position = "none")
}
violin_plots <-
  Map(violin_plot_fun, i = metrics_metadata, j = column_variables)
violin_plots

plotnames = imap(violin_plots, ~paste0("../../results/clinical_representative_samples/TYPE.", .y, ".pdf")) %>%
  flatten()
plotnames <- gsub("metadata.", "", plotnames)
walk2(plotnames, violin_plots, ~ggsave(filename = .x, plot = .y, 
                                    height = 4, width = 5.5))

remove(violin_plots) 
```

# Correlation heatmap
scale() function is used to ensure that variables measured in different scales (e.g. age of death vs RIN) are comparable.
same results with and without the use of scale
```{r correlation, message=FALSE, warning=FALSE, tidy=TRUE}
# scale continous data 
metadata <-
  metadata[, c(
    "MF.NFT",
    "MF.Amyloid",
    "MF.SP",
    "MF.Tau",
    "Brain.wt",
    "Duration",
    "Age",
    "PMI",
    "RIN",
    "Braak.NFT",
    "Thal.amyloid",
    "Cing.LB",
    "Cing.Synuclein"
  )] %>% scale()
info <- as.data.frame(metadata)

# correlation
cor_mat <- rcorr(as.matrix(info))
corrplot(
  cor_mat$r,
  method = "color",
  col = correlationColors(200),
  type = "upper",
  order = "hclust",
  p.mat = cor_mat$P,
  addCoef.col = "black",
  # Add coefficient of correlation
  tl.col = "black",
  tl.srt = 45,
  #Text label color and rotation
  diag = FALSE,
  col.lim = c(-1, 1)
)
path <- paste0("../../results/clinical_representative_samples/correlation_continuous_variables")
saveToPDF(paste0(path, ".pdf"), width = 10, height = 10)
```


