---
title: "RNAseq representative sample selection"
author: "Kimberly Olney"
date: "2023-05-11"
output:
  pdf_document: default
  html_document: default
---
# Outlier detection and representative sample selection via PCA 
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = ".")
```
# Libraries packages
```{r}
# remotes::install_github("wanchanglin/mt")
library(mt) #Metabolomics Data Analysis Toolbox
library('rlist') # for appending lists
library(dplyr)
```
# Read in counts & sample information object 
```{r}
dge.filtered.norm <- readRDS("../../rObjects/dge.filtered.norm.rds")
```
# Reformat data for PCA outlier detection 
```{r}
# create counts matrix
lcpm <- edgeR::cpm(dge.filtered.norm$counts, log = TRUE)
tcounts <- t(lcpm) # transpose so rows are sample IDs
df <- as.data.frame(tcounts) # save as dataframe
df$TYPE <- dge.filtered.norm$samples$TYPE # add group information as a column

sample_info <- dge.filtered.norm$samples
# clean up
remove(lcpm, tcounts)
```
# PCA outlier detection
### All samples 
```{r}
# plot 
pca.outlier(df[,1:14183], adj=-0.5, groups = dge.filtered.norm$samples$TYPE)
# save to obtain sample IDs that are outliers 
all_samples <- pca.outlier(df[,1:14183], adj=-0.5, groups = dge.filtered.norm$samples$TYPE)
exclude_samples <- as.data.frame(all_samples$outlier)
all_outlier_samples <- rownames(exclude_samples) # list of sample outliers

# clean up
remove(all_samples, exclude_samples)
```
## Each disease type, RIN >= 6 
Find representative samples within each disease type. 
Let us also look at samples with a RIN >= 6
```{r}
df$RIN <- dge.filtered.norm$samples$RIN # add RIN information as a column
summary(df$RIN) # median RIN is 6 and mean is 5.8
df_RIN6 <- subset(df, RIN >= 6) # subset to keep samples with RIN equal or greater than 6

# PCA with subset by RIN 
pca.outlier(df_RIN6[,1:14183], adj=-0.5, groups = df_RIN6$TYPE)
```
```{r}
disease_types <- c("CONTROL","PA", "AD", "LBD")
for (i in disease_types){
  df_RIN6 <- subset(df_RIN6, TYPE == "CONTROL") 
  pca.outlier(df_RIN6[,1:14183], adj=-0.5) # plot
  samples <- pca.outlier(df_RIN6[,1:14183], adj=-0.5)
  exclude_samples <- as.data.frame(samples$outlier)
  outlier_samples <- rownames(exclude_samples) # list of sample outliers
  assign(paste0(i, "_outlier_samples"), outlier_samples)
  # clean up
  remove(samples, exclude_samples, outlier_samples)
}
```
# Remove outliers, 
# Randomly select representatives 
```{r}
outliers <-
  list.append(
    CONTROL_outlier_samples,
    PA_outlier_samples,
    AD_outlier_samples,
    LBD_outlier_samples
  )

# subset sample info to keep sample with RIN >= 6 
sample_info_RIN6 <- subset(sample_info, RIN >= 6)
# remove outliers
sample_info_RIN6_no_outliers <- sample_info_RIN6[!sample_info_RIN6$NPID %in% outliers,]

PA_samples <- subset(sample_info_RIN6_no_outliers, TYPE == "PA")
set.seed(28) # to get the same random next time
disease_types <- c("CONTROL", "AD", "LBD")
for (i in disease_types){
  female <- subset(sample_info_RIN6_no_outliers, TYPE == i & sex_inferred == "female")
  female_8 <- sample_n(female, 8)
  assign(paste0(i, "_female"), female_8)
  male <- subset(sample_info_RIN6_no_outliers, TYPE == i & sex_inferred == "male")
  male_8 <- sample_n(male, 8)
  assign(paste0(i, "_male"), male_8)
  remove(female, female_8, male, male_8)
}

rep_samples <-
  rbind(CONTROL_female,
        CONTROL_male,
        PA_samples,
        AD_female,
        AD_male,
        LBD_female,
        LBD_male)

write.table(
  rep_samples,
  "repersentative_samples.tsv",
  quote = FALSE,
  row.names = FALSE,
  sep = "\t"
)
```
