---
title: "Figure 1. LBD vs Control"
author: "Kimberly Olney"
date: "07/10/2023"
output: html_document
---
# Setup
```{r setup}
knitr::opts_knit$set(root.dir = ".")
```
# User defined variables
```{r set_variables}
source(here::here("scripts/R", "file_paths_and_colours.R"))
library("readxl")
#source(here::here("scripts/R", "gtf_path.R"))
condition <- c("TYPE")
tool <- "star"
dge.filtered.norm <- readRDS(paste0("../../../rObjects/dge.filtered.norm.rds"))
```
# Binary list shared gene function
```{r}
fromList <- function (input) {
  # Same as original fromList()...
  elements <- unique(unlist(input))
  data <- unlist(lapply(input, function(x) {
      x <- as.vector(match(elements, x))
      }))
  data[is.na(data)] <- as.integer(0)
  data[data != 0] <- as.integer(1)
  data <- data.frame(matrix(data, ncol = length(input), byrow = F))
  data <- data[which(rowSums(data) != 0), ]
  names(data) <- names(input)
  # ... Except now it conserves your original value names!
  row.names(data) <- elements
  return(data)
  }
```
# LBD vs Control Volcano 
```{r volcano, warning=FALSE}
i <- "LBDvsControl"
group1_vs_group2 <- read.delim(paste0("../../../results/", tool, "/DEGs/", condition, "_", i, "_gene_DEGs_FDRq1.00.txt"))
  color_values <- vector()
  max <- nrow(group1_vs_group2)
  for (row in 1:max) {
    if (group1_vs_group2$adj.P.Val[row] < 0.05) {
      if (group1_vs_group2$logFC [row] > 0.25) {
        color_values <- c(color_values, 1)
      }
      else if (group1_vs_group2$logFC[row] < -.25) {
        color_values <- c(color_values, 2)
      }
      else {
        color_values <- c(color_values, 3)
      }
    }
    else{
      color_values <- c(color_values, 3)
    }
  }
  group1_vs_group2$color_adjpval_0.05 <- factor(color_values)
  data <- group1_vs_group2
  # plot only if there are DEGs with p_val_adj < 0.05
  num <- subset(data, (adj.P.Val < 0.05 & logFC < -.25)  | (adj.P.Val < 0.05 & logFC > .25 ))
  num <- nrow(num)
    up <- data[data$color_adjpval_0.05 == 1,]
    up10 <- up[1:10,]
    upFold <- subset(up, logFC > .75)
    upFold <- upFold[!(upFold$gene_name %in% up10$gene_name),]
    down <- data[data$color_adjpval_0.05 == 2,]
    down10 <- down[1:10,]
    downFold <- subset(down, logFC < -.75)
    downFold <- downFold[!(downFold$gene_name %in% down10$gene_name),]
    if (!1 %in% unique(data$color_adjpval_0.05)) {
      my_colors <- c("blue", "gray")
    } else if (!2 %in% unique(data$color_adjpval_0.05)) {
      my_colors <- c("red", "gray")
    } else if (!1 %in% unique(data$color_adjpval_0.05) &&
               !2 %in% unique(data$color_adjpval_0.05)) {
      my_colors <- c("gray")
    } else {
      my_colors <- c("red", "blue", "gray")
    }
    hadjpval <- (-log10(max(data$P.Value[data$adj.P.Val < 0.05],
                            na.rm = TRUE)))
    negFC <- c(-.25)
    posFC <- c(.25) 
    i <- gsub("vs", " vs ", i)
volcano <- ggplot(data = data,
             aes(
               x = logFC,
               y = -log10(P.Value),
               color = color_adjpval_0.05
             )) +
      geom_point(alpha = 0.8, size = 1) +
      theme_bw() +
      theme(legend.position = "none", 
          plot.margin = margin(0.1, 0.2, 0.5, 0.3, "cm"), #t,r,b,l
             axis.title.x = element_text(size = 8),
             axis.text.x = element_text(size = 8),
             axis.title.y = element_text(size = 8),
             axis.text.y = element_text(size = 8),
             plot.title = element_text(size = 8, margin = margin(0,0,0.2,0))) +
      scale_color_manual(values = my_colors) +
      labs(
        title = "",
        x = expression(log[2](FC)),
        y = expression(-log[10] ~ "(" ~ italic("p") ~ "-value)")
      )  +
      #theme(plot.title = element_text(size = 7, margin = margin(0,0,0,0))) +
      geom_hline(yintercept = hadjpval,
                 #  horizontal line
                 colour = "#000000",
                 linetype = "dashed") +
      geom_vline(xintercept = negFC,
                 #  horizontal line
                 colour = "#000000",
                 linetype = "dashed") +
      geom_vline(xintercept = posFC,
                 #  horizontal line
                 colour = "#000000",
                 linetype = "dashed") +
      ggtitle(paste0(i)) +
      geom_text_repel(
        data = up10,
        aes(
          x = logFC,
          y = -log10(P.Value),
          label = gene_name
        ),
        nudge_x = 0.1,
        color = "maroon",
        fontface = "italic",
        size = 2,
        max.overlaps = getOption("ggrepel.max.overlaps", default = 8)
      ) +
      geom_text_repel(
        data = upFold,
        aes(
          x = logFC,
          y = -log10(P.Value),
          label = gene_name
        ),
        nudge_x = 0.1,
        color = "maroon",
        fontface = "italic",
        size = 2,
        max.overlaps = getOption("ggrepel.max.overlaps", default = 5)
      ) +
      geom_text_repel(
        data = down10,
        aes(
          x = logFC,
          y = -log10(P.Value),
          label = gene_name
        ),
        x_nudge = -.1,
        y_nudge = 1,
        color = "navyblue",
        fontface = "italic",
        size = 2,
        max.overlaps = getOption("ggrepel.max.overlaps", default = 8)
      ) +
      geom_text_repel(
        data = downFold,
        aes(
          x = logFC,
          y = -log10(P.Value),
          label = gene_name
        ),
        x_nudge = -.3,
        color = "navyblue",
        fontface = "italic",
        size = 2,
        max.overlaps = getOption("ggrepel.max.overlaps", default = 5)
      ) +
  scale_x_continuous(limits = c(-1.75, 1.75), breaks = seq(-2, 2, .5))
volcano
i <- gsub(" vs ", "vs", i)
```

```{r cleanup}
# clean up
remove(up, up10, upFold, downFold, down10, data)
```

# Metascape Enrichment 
### Format the dataframes
```{r}
# read in enrichment analysis results
up_enrich_results <-
  read_excel(
    paste0(
      "../../../results/",
      tool,
      "/metascape/",
      i,
      "_up_regulated/metascape_result.xlsx"
    ), sheet = 2
  )

down_enrich_results <-
  read_excel(
    paste0(
      "../../../results/",
      tool,
      "/metascape/",
      i,
      "_down_regulated/metascape_result.xlsx"
    ), sheet = 2
  )
```

```{r}
# select the GO term IDs we want to show in the plot
GO_ID_up <-
  c(
    "1_Summary",
    "2_Summary",
    "3_Summary",
    "4_Summary",
    "5_Summary",
    "6_Summary",
    "7_Summary",
    "8_Summary",
    "9_Summary", 
    "10_Summary",
    "11_Summary",
    "12_Summary",
    "13_Summary",
    "14_Summary",
    "15_Summary",
    "16_Summary",
    "17_Summary",
    "18_Summary",
    "19_Summary", 
    "20_Summary"
  )
GO_ID_down <- c(
    "1_Summary",
    "2_Summary",
    "3_Summary",
    "4_Summary",
    "5_Summary",
    "6_Summary",
    "7_Summary",
    "8_Summary",
    "9_Summary", 
    "10_Summary",
    "11_Summary",
    "12_Summary",
    "13_Summary",
   # "14_Summary",
   # "15_Summary",
    "16_Summary"
  #  "17_Summary",
  #  "18_Summary",
   # "19_Summary", 
   # "20_Summary"
)

up_enrich_results_subset <-
  up_enrich_results[up_enrich_results$GroupID %in% GO_ID_up, ]
up_enrich_results_subset$Cluster <- c("up-regulated")
up_enrich_results_subset$Description <-
  factor(up_enrich_results_subset$Description,
         levels = up_enrich_results_subset$Description)
up_enrich_results_subset$Description <-
  fct_rev(up_enrich_results_subset$Description)

down_enrich_results_subset <-
  down_enrich_results[down_enrich_results$GroupID %in% GO_ID_down, ]
down_enrich_results_subset$Cluster <- c("down-regulated")
down_enrich_results_subset$Description <-
  factor(down_enrich_results_subset$Description,
         levels = down_enrich_results_subset$Description)
down_enrich_results_subset$Description <-
  fct_rev(down_enrich_results_subset$Description)

# get the number of genes in each summary
up_gene_count <-
  strsplit(as.character(up_enrich_results_subset$InTerm_InList),
           "/",
           fixed = T)
up_gene_count_df <-
  data.frame(matrix(
    unlist(up_gene_count),
    nrow = length(up_gene_count),
    byrow = TRUE
  ), stringsAsFactors = FALSE)
up_enrich_results_subset$InTerm <- as.numeric(up_gene_count_df$X1)

down_gene_count <-
  strsplit(as.character(down_enrich_results_subset$InTerm_InList),
           "/",
           fixed = T)
down_gene_count_df <-
  data.frame(matrix(
    unlist(down_gene_count),
    nrow = length(down_gene_count),
    byrow = TRUE
  ), stringsAsFactors = FALSE)
down_enrich_results_subset$InTerm <- as.numeric(down_gene_count_df$X1)

# combine together
up_and_down_enrich_results_subset <- rbind(up_enrich_results_subset, down_enrich_results_subset)

up_and_down_enrich_results_subset$Description <-
      gsub(
       "regulation of tumor necrosis factor superfamily cytokine production",
        "reg tumor necrosis factor cytokine production",
        up_and_down_enrich_results_subset$Description
      )
up_and_down_enrich_results_subset$Description <-
      gsub(
       "Trans-sulfuration, one-carbon metabolism and related pathways",
        "Trans-sulfuration, one-carbon metabolism & pathways",
        up_and_down_enrich_results_subset$Description
      )
up_and_down_enrich_results_subset$Description <-
  factor(up_and_down_enrich_results_subset$Description, levels = unique(up_and_down_enrich_results_subset$Description)) 

up_and_down_enrich_results_subset$Description <- fct_rev(up_and_down_enrich_results_subset$Description)
up_and_down_enrich_results_subset$Cluster <- factor(up_and_down_enrich_results_subset$Cluster, levels = c("up-regulated", "down-regulated"))
```

remove files 
```{r}
remove(
  down_enrich_results,
  down_gene_count,
  down_gene_count_df,
  up_enrich_results,
  up_gene_count,
  up_gene_count_df
)
```

### Enrichment plot showing log10 P-value
#### down and up seperately
```{r}
down_enrich_plot <-
  ggplot(data = down_enrich_results_subset, aes(x = InTerm, y = Description)) +
  ggplot2::facet_grid(~ Cluster) +
  geom_bar(stat = "identity", aes(fill = LogP), width = .65) +
  theme(axis.title.x = element_text(size = 8),
        axis.text.x = element_text(size = 8)) +
  theme(axis.title.y = element_text(size = 8),
        axis.text.y = element_text(size = 8)) +
  theme_bw() +
  scale_y_discrete(expand = c(0, 0)) +
  labs(x = "Gene count", y = "") +
  #  xlim(0,60) +
  guides(fill = guide_legend(
    size = 1,
    title = expression(log[10] ~ "(" ~ italic("p") ~ "-value)")
  )) +
  scale_fill_gradientn(
    colours = c("#800026FF", "#FD8D3CFF", "#FFFFCCFF"), 
    guide = "legend",
    limits = c(-45,-1),
  ) +
  theme(legend.text = element_text(size = 8),
        legend.title = element_text(size = 8))
down_enrich_plot

up_enrich_plot <-
  ggplot(data = up_enrich_results_subset, aes(x = InTerm, y = Description)) +
  ggplot2::facet_grid(~ Cluster) +
  geom_bar(stat = "identity", aes(fill = LogP), width = .65) +
  theme(axis.title.x = element_text(size = 8),
        axis.text.x = element_text(size = 8)) +
  theme(axis.title.y = element_text(size = 8),
        axis.text.y = element_text(size = 8)) +
  theme(legend.text = element_text(size = 8)) +
  theme_bw() +
  scale_y_discrete(expand = c(0, 0)) +
  labs(x = "Gene count", y = "") +
  guides(fill = guide_legend(title = expression(log[10] ~ "(" ~ italic("p") ~ "-value)"))) +
  scale_fill_gradientn(
    colours = c("#800026FF", "#FD8D3CFF", "#FFFFCCFF"),
    guide = "legend",
    limits = c(-45,-1)
  ) +
  theme(legend.text = element_text(size = 8),
        legend.title = element_text(size = 8))
up_enrich_plot
```

#### down and up together
```{r}
up_and_down_enrich_plot <-
  ggplot(data = up_and_down_enrich_results_subset, aes(x = InTerm, y = Description)) +
  #ggplot2::facet_wrap(~ Cluster, scales = "free_y", ncol = 1) +
 # facet_grid(Cluster ~ ., space="free", scales="free") +
  ggforce::facet_col(facets = vars(Cluster), 
                     scales = "free_y", 
                     space = "free") +
  geom_bar(stat = "identity", aes(fill = LogP), width = .7, position = position_dodge(width = .2)) +
  theme_bw() +
  ggtitle(paste0("Enrichment summaries")) +
  labs(x = "Gene count", y = NULL) +
  guides(fill = guide_legend(title = expression(log[10] ~ "(" ~ italic("p") ~ "-value)"))) +
  scale_fill_gradientn(
    colours = c("#800026FF", "#FD8D3CFF", "gray"),
    guide = "legend",
    limits = c(-22,-3)
  ) +
    theme(strip.text = element_text(size = 8), 
          axis.text.y = element_text(size = 8),
          axis.text.x = element_text(size = 8), 
          axis.title.x = element_text(size = 8), 
          legend.text = element_text(size = 8),
          legend.title = element_text(size = 8),
          legend.margin=margin(0,0.5,0,0),
          legend.box.margin=margin(-10,-2,-10,-7.5), 
          plot.margin = margin(0.1, 0.2, 0.5, 0.3, "cm"), #t,r,b,l
          plot.title = element_text(size = 8, hjust = -2.65, vjust=0, margin = margin(0,0,0,0)))
 #  theme(aspect.ratio = 1.5) # to reduce the size of the boxes 
addSmallLegend <- function(myPlot, pointSize = 4, textSize = 7, spaceLegend = .5) {
    myPlot +
        guides(shape = guide_legend(override.aes = list(size = pointSize)),
               color = guide_legend(override.aes = list(size = pointSize))) +
        theme(legend.title = element_text(size = textSize), 
              legend.text  = element_text(size = textSize),
              legend.key.size = unit(spaceLegend, "lines"))
}

# Apply on original plot
up_and_down_enrich_plot <- addSmallLegend(up_and_down_enrich_plot)
up_and_down_enrich_plot
```

```{r}
remove(up_enrich_plot,
       down_enrich_plot)
```
## Combine volcano and enrichment plot 
```{r}
row1 <- ggarrange(
  volcano,
  up_and_down_enrich_plot,
  ncol = 2,
  labels = c("a)", "b)"),
  widths = c(2.5, 4.25), 
  font.label = list(size = 8)
  )
row1
#path <- paste0("../../../results/manuscript_figures/Figure_1_row1_test")
#saveToPDF(paste0(path, ".pdf"), width = 8.5, height = 5)
```
# DLBD vs TLBD volcano
```{r DLBDvsTLBD}
i <- "DLBDvsTLBD"
tool <- "star_LBD_sub_type"
condition <- c("LBD.type.")
group1_vs_group2 <- read.delim(paste0("../../../results/", tool, "/DEGs/", condition, "_", i, "_gene_DEGs_FDRq1.00.txt"))
  color_values <- vector()
  max <- nrow(group1_vs_group2)
  for (row in 1:max) {
    if (group1_vs_group2$adj.P.Val[row] < 0.05) {
      if (group1_vs_group2$logFC [row] > 0.25) {
        color_values <- c(color_values, 1)
      }
      else if (group1_vs_group2$logFC[row] < -.25) {
        color_values <- c(color_values, 2)
      }
      else {
        color_values <- c(color_values, 3)
      }
    }
    else{
      color_values <- c(color_values, 3)
    }
  }
  group1_vs_group2$color_adjpval_0.05 <- factor(color_values)
  data <- group1_vs_group2
  # plot only if there are DEGs with p_val_adj < 0.05
  num <- subset(data, (adj.P.Val < 0.05 & logFC < -.25)  | (adj.P.Val < 0.05 & logFC > .25 ))
  num <- nrow(num)
    up <- data[data$color_adjpval_0.05 == 1,]
    up10 <- up[1:10,]
    upFold <- subset(up, logFC > .45)
    upFold <- upFold[!(upFold$gene_name %in% up10$gene_name),]
    down <- data[data$color_adjpval_0.05 == 2,]
    down10 <- down[1:10,]
    downFold <- subset(down, logFC < -.75)
    downFold <- downFold[!(downFold$gene_name %in% down10$gene_name),]
    if (!1 %in% unique(data$color_adjpval_0.05)) {
      my_colors <- c("blue", "gray")
    } else if (!2 %in% unique(data$color_adjpval_0.05)) {
      my_colors <- c("red", "gray")
    } else if (!1 %in% unique(data$color_adjpval_0.05) &&
               !2 %in% unique(data$color_adjpval_0.05)) {
      my_colors <- c("gray")
    } else {
      my_colors <- c("red", "blue", "gray")
    }
    hadjpval <- (-log10(max(data$P.Value[data$adj.P.Val < 0.05],
                            na.rm = TRUE)))
    negFC <- c(-.25)
    posFC <- c(.25) 
    i <- gsub("vs", " vs ", i)
DLBDvsTLBD_volcano <- ggplot(data = data,
             aes(
               x = logFC,
               y = -log10(P.Value),
               color = color_adjpval_0.05
             )) +
      geom_point(alpha = 0.8, size = 1) +
      theme_bw() +
      theme(legend.position = "none", 
          plot.margin = margin(0.1, 0.2, 0.5, 0.3, "cm"), #t,r,b,l
             axis.title.x = element_text(size = 8),
             axis.text.x = element_text(size = 8),
             axis.title.y = element_text(size = 8),
             axis.text.y = element_text(size = 8),
             plot.title = element_text(size = 8, margin = margin(0,0,0.3,0))) +
      scale_color_manual(values = my_colors) +
      labs(
        title = "",
        x = expression(log[2](FC)),
        y = expression(-log[10] ~ "(" ~ italic("p") ~ "-value)")
      )  +
      #theme(plot.title = element_text(size = 7, margin = margin(0,0,0,0))) +
      geom_hline(yintercept = hadjpval,
                 #  horizontal line
                 colour = "#000000",
                 linetype = "dashed") +
      geom_vline(xintercept = negFC,
                 #  horizontal line
                 colour = "#000000",
                 linetype = "dashed") +
      geom_vline(xintercept = posFC,
                 #  horizontal line
                 colour = "#000000",
                 linetype = "dashed") +
      ggtitle(paste0(i)) +
      geom_text_repel(
        data = up10,
        aes(
          x = logFC,
          y = -log10(P.Value),
          label = gene_name
        ),
        color = "maroon",
        fontface = "italic",
        size = 2,
        max.overlaps = getOption("ggrepel.max.overlaps", default = 8)
      ) +
      geom_text_repel(
        data = upFold,
        aes(
          x = logFC,
          y = -log10(P.Value),
          label = gene_name
        ),
        color = "maroon",
        fontface = "italic",
        size = 2,
        max.overlaps = getOption("ggrepel.max.overlaps", default = 5)
      ) +
      geom_text_repel(
        data = down10,
        aes(
          x = logFC,
          y = -log10(P.Value),
          label = gene_name
        ),
        x_nudge = -.1,
        y_nudge = 1,
        color = "navyblue",
        fontface = "italic",
        size = 2,
        max.overlaps = getOption("ggrepel.max.overlaps", default = 8)
      ) +
      geom_text_repel(
        data = downFold,
        aes(
          x = logFC,
          y = -log10(P.Value),
          label = gene_name
        ),
        x_nudge = -.3,
        color = "navyblue",
        fontface = "italic",
        size = 2,
        max.overlaps = getOption("ggrepel.max.overlaps", default = 5)
      ) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, .5)) +
  scale_y_continuous(limits = c(0, 8), breaks = seq(0, 8, 2))

DLBDvsTLBD_volcano
i <- gsub(" vs ", "vs", i)
```

# DLBD vs TLBD metascape 
```{r}
# read in enrichment analysis results
i <- c("DLBD_vs_TLBD")
up_enrich_results <-
  read_excel(
    paste0(
      "../../../results/",
      tool,
      "/metascape/",
      i,
      "_up/metascape_result.xlsx"
    ), sheet = 2
  )

down_enrich_results <-
  read_excel(
    paste0(
      "../../../results/",
      tool,
      "/metascape/",
      i,
      "_down/metascape_result.xlsx"
    ), sheet = 2
  )

# select the GO term IDs we want to show in the plot
GO_ID_up <-
  c(
    "1_Summary",
    "2_Summary",
    "3_Summary",
    "4_Summary",
    "5_Summary",
    "6_Summary",
    "7_Summary",
    "8_Summary",
    "9_Summary", 
    "10_Summary",
    "11_Summary",
    "12_Summary",
    "13_Summary",
    "14_Summary",
    "15_Summary",
    "16_Summary",
    "17_Summary",
    "18_Summary",
    "19_Summary", 
    "20_Summary"
  )
GO_ID_down <- c(
    "1_Summary",
    "2_Summary",
    "3_Summary",
    "4_Summary",
    "5_Summary",
    "6_Summary",
    "7_Summary",
    "8_Summary",
    "9_Summary", 
    "10_Summary",
    "11_Summary",
    "12_Summary",
    "13_Summary",
    "14_Summary",
    "15_Summary",
    "16_Summary",
    "17_Summary",
    "18_Summary",
    "19_Summary", 
    "20_Summary"
)

up_enrich_results_subset <-
  up_enrich_results[up_enrich_results$GroupID %in% GO_ID_up, ]
up_enrich_results_subset$Cluster <- c("up-regulated")
up_enrich_results_subset$Description <-
  factor(up_enrich_results_subset$Description,
         levels = up_enrich_results_subset$Description)
up_enrich_results_subset$Description <-
  fct_rev(up_enrich_results_subset$Description)

down_enrich_results_subset <-
  down_enrich_results[down_enrich_results$GroupID %in% GO_ID_down, ]
down_enrich_results_subset$Cluster <- c("down-regulated")
down_enrich_results_subset$Description <-
  factor(down_enrich_results_subset$Description,
         levels = down_enrich_results_subset$Description)
down_enrich_results_subset$Description <-
  fct_rev(down_enrich_results_subset$Description)

# get the number of genes in each summary
up_gene_count <-
  strsplit(as.character(up_enrich_results_subset$InTerm_InList),
           "/",
           fixed = T)
up_gene_count_df <-
  data.frame(matrix(
    unlist(up_gene_count),
    nrow = length(up_gene_count),
    byrow = TRUE
  ), stringsAsFactors = FALSE)
up_enrich_results_subset$InTerm <- as.numeric(up_gene_count_df$X1)

down_gene_count <-
  strsplit(as.character(down_enrich_results_subset$InTerm_InList),
           "/",
           fixed = T)
down_gene_count_df <-
  data.frame(matrix(
    unlist(down_gene_count),
    nrow = length(down_gene_count),
    byrow = TRUE
  ), stringsAsFactors = FALSE)
down_enrich_results_subset$InTerm <- as.numeric(down_gene_count_df$X1)

# combine together
up_and_down_enrich_results_subset <- rbind(up_enrich_results_subset, down_enrich_results_subset)

up_and_down_enrich_results_subset$Description <-
      gsub(
       "negative regulation of cellular component organization",
        "negative reg of cellular component organization",
        up_and_down_enrich_results_subset$Description
      )
up_and_down_enrich_results_subset$Description <-
      gsub(
       "positive regulation of cytosolic calcium ion concentration",
        "positive reg of cytosolic calcium ion concentration",
        up_and_down_enrich_results_subset$Description
      )
up_and_down_enrich_results_subset$Description <-
  factor(up_and_down_enrich_results_subset$Description, levels = unique(up_and_down_enrich_results_subset$Description)) 

up_and_down_enrich_results_subset$Description <- fct_rev(up_and_down_enrich_results_subset$Description)
up_and_down_enrich_results_subset$Cluster <- factor(up_and_down_enrich_results_subset$Cluster, levels = c("up-regulated", "down-regulated"))

remove(
  down_enrich_results,
  down_gene_count,
  down_gene_count_df,
  up_enrich_results,
  up_gene_count,
  up_gene_count_df
)

DLBDvsTLBD_up_and_down_enrich_plot <-
  ggplot(data = up_and_down_enrich_results_subset, aes(x = InTerm, y = Description)) +
  #ggplot2::facet_wrap(~ Cluster, scales = "free_y", ncol = 1) +
 # facet_grid(Cluster ~ ., space="free", scales="free") +
  ggforce::facet_col(facets = vars(Cluster), 
                     scales = "free_y", 
                     space = "free") +
  geom_bar(stat = "identity", aes(fill = LogP), width = .7, position = position_dodge(width = .2)) +
  theme_bw() +
  ggtitle(paste0("Enrichment summaries")) +
  labs(x = "Gene count", y = NULL) +
  guides(fill = guide_legend(title = expression(log[10] ~ "(" ~ italic("p") ~ "-value)"))) +
  scale_fill_gradientn(
    colours = c("#800026FF", "#FD8D3CFF", "gray"),
    guide = "legend",
    limits = c(-13,-1)
  ) +
    theme(strip.text = element_text(size = 8), 
          axis.text.y = element_text(size = 8),
          axis.text.x = element_text(size = 8), 
          axis.title.x = element_text(size = 8), 
          legend.text = element_text(size = 8),
          legend.title = element_text(size = 8),
          legend.margin=margin(0,0.5,0,0),
          legend.box.margin=margin(-10,-2,-10,-7.5), 
          plot.margin = margin(0.1, 0.2, 0.5, 0.3, "cm"), #t,r,b,l
          plot.title = element_text(size = 8, hjust = -3, vjust=0, margin = margin(0,0,0,0)))
 #  theme(aspect.ratio = 1.5) # to reduce the size of the boxes 
addSmallLegend <- function(myPlot, pointSize = 4, textSize = 7, spaceLegend = .5) {
    myPlot +
        guides(shape = guide_legend(override.aes = list(size = pointSize)),
               color = guide_legend(override.aes = list(size = pointSize))) +
        theme(legend.title = element_text(size = textSize), 
              legend.text  = element_text(size = textSize),
              legend.key.size = unit(spaceLegend, "lines"))
}

# Apply on original plot
DLBDvsTLBD_up_and_down_enrich_plot <- addSmallLegend(DLBDvsTLBD_up_and_down_enrich_plot)
DLBDvsTLBD_up_and_down_enrich_plot
```
# DLBD Tau vs DLBD no Tau volcano
```{r DLBDvsTLBD}
group1_vs_group2 <- read.delim(paste0("../../../results/star_LBD_sub_type/DEGs/LBD.type.tau_DLBD_Tau_vs_DLBD_NoTau_gene_DEGs_FDRq1.00.txt"))
  color_values <- vector()
  max <- nrow(group1_vs_group2)
  for (row in 1:max) {
    if (group1_vs_group2$adj.P.Val[row] < 0.05) {
      if (group1_vs_group2$logFC [row] > 0.25) {
        color_values <- c(color_values, 1)
      }
      else if (group1_vs_group2$logFC[row] < -.25) {
        color_values <- c(color_values, 2)
      }
      else {
        color_values <- c(color_values, 3)
      }
    }
    else{
      color_values <- c(color_values, 3)
    }
  }

group1_vs_group2$color_adjpval_0.05 <- factor(color_values)
data <- group1_vs_group2

upFold <- subset(data, logFC > .30)
downFold <- subset(data, logFC < -.35)
almost_sig <- subset(data, P.Value < .001)
SERPINA5 <- subset(data, gene_name == "SERPINA5")


DLBD_Tau_vs_DLBD_NoTau_volcano <- ggplot(data = data,
             aes(
               x = logFC,
               y = -log10(P.Value),
               color = color_adjpval_0.05
             )) +
      geom_point(alpha = 0.8, size = 1) +
      theme_bw() +
      theme(legend.position = "none", 
          plot.margin = margin(0.1, 0.2, 0.5, 0.3, "cm"), #t,r,b,l
             axis.title.x = element_text(size = 8),
             axis.text.x = element_text(size = 8),
             axis.title.y = element_text(size = 8),
             axis.text.y = element_text(size = 8),
             plot.title = element_text(size = 8, margin = margin(0,0,0.3,0))) +
      scale_color_manual(values = "gray") + #
      labs(
        title = "",
        x = expression(log[2](FC)),
        y = expression(-log[10] ~ "(" ~ italic("p") ~ "-value)")
      )  +
      geom_vline(xintercept = negFC,
                 #  horizontal line
                 colour = "#000000",
                 linetype = "dashed") +
      geom_vline(xintercept = posFC,
                 #  horizontal line
                 colour = "#000000",
                 linetype = "dashed") +
      ggtitle("DLBD Tau+ vs DLBD Tau-") +
        geom_text_repel(
        data = upFold,
        aes(
          x = logFC,
          y = -log10(P.Value),
          label = gene_name
        ),
        color = "grey25",
        fontface = "italic",
        size = 2,
        max.overlaps = getOption("ggrepel.max.overlaps", default = 5)
      ) +
      geom_text_repel(
        data = downFold,
        aes(
          x = logFC,
          y = -log10(P.Value),
          label = gene_name
        ),
        color = "grey25",
        fontface = "italic",
        size = 2,
        max.overlaps = getOption("ggrepel.max.overlaps", default = 5)
      ) +
        geom_text_repel(
        data = almost_sig,
        aes(
          x = logFC,
          y = -log10(P.Value),
          label = gene_name
        ),
        color = "grey25",
        fontface = "italic",
        size = 2,
        max.overlaps = getOption("ggrepel.max.overlaps", default = 5)
      ) +
        geom_text_repel(
        data = SERPINA5,
        aes(
          x = logFC,
          y = -log10(P.Value),
          label = gene_name
        ),
        color = "grey25",
        fontface = "italic",
        size = 2,
        max.overlaps = getOption("ggrepel.max.overlaps", default = 20)
      ) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, .5)) +
  scale_y_continuous(limits = c(0, 8), breaks = seq(0, 8, 2))
DLBD_Tau_vs_DLBD_NoTau_volcano
```
# Combine all plots 
```{r}
row2 <- ggarrange(
  DLBDvsTLBD_volcano,
  DLBDvsTLBD_up_and_down_enrich_plot,
  labels = c("c)", "d)"), 
  widths = c(2.5, 4.25), 
  font.label = list(size = 8)
  )

row3 <- ggarrange(
  DLBD_Tau_vs_DLBD_NoTau_volcano,
  NULL,
  labels = c("e)", ""), 
  widths = c(2.5, 4.25), 
  font.label = list(size = 8)
)
combind <-
  ggarrange(
    row1,
    row2,
    row3,
    nrow = 3,
    heights = c(1.2, 1, 1)
    )
combind

path <- paste0("../../../results/manuscript_figures/Figure_1_LBD_vs_Control_DLBD_vs_TLBD")
saveToPDF(paste0(path, ".pdf"), width = 8.5, height = 11)
```

# Gene violins 
```{r eval = FALSE}
dge.filtered.norm <- readRDS(paste0("../../../rObjects/dge.filtered.norm.rds"))
voomCountsMatrix <- readRDS("/research/labs/neurology/fryer/m239830/LBD_CWOW/rObjects/TYPE.voomCountsMatrix.rds")
# get the gene id that corresponds to the gene name 
genes <- dge.filtered.norm$genes # gene information 
goi <- subset(genes, gene_name == "SERPINA5" | gene_name == "MAG" | gene_name == "LGR6" | gene_name == "HBB") # gene of interest 

# subset counts 
goi_counts <- subset(voomCountsMatrix, rownames(voomCountsMatrix) %in% rownames(goi))
goi_melt <- reshape2::melt(goi_counts) # reshape data 
names(goi_melt)[names(goi_melt) == "Var2"] <- "NPID" # rename columns to merge with metadata 
names(goi_melt)[names(goi_melt) == "Var1"] <- "gene_id"
df <- merge(goi_melt, metadata, by = "NPID")

colors <- c("#4682B4", "thistle3", "thistle4")
df$LBD.type <- factor(df$LBD.type, levels =c("no LBs", "TLBD", "DLBD", "BLBD"))
df <- subset(df, LBD.type != "BLBD")
df <- subset(df, TYPE != "AD")
df <- subset(df, TYPE != "PA")

df$LBD.type <-gsub("no LBs","Control",df$LBD.type)
df$LBD.type <- factor(df$LBD.type, levels =c("Control", "TLBD", "DLBD"))
```
### MAG
```{r MAG}
exp_data <- subset(df, gene_id == "ENSG00000105695.15")
MAG_exp_plot <- ggplot(exp_data, aes(LBD.type, value, fill = LBD.type)) +
    geom_violin() +
    geom_boxplot(width = 0.1, outlier.shape=NA, color = "black") + 
   # geom_jitter(shape=16, position=position_jitter(0.2), alpha = 0.1) +
    theme_bw() +
    theme(legend.position = "none", 
          plot.margin = margin(0.1, 0.2, 0.5, 0.3, "cm"), #t,r,b,l
             axis.title.x = element_text(size = 8),
             axis.text.x = element_text(size = 8),
             axis.title.y = element_text(size = 8),
             axis.text.y = element_text(size = 8),
             plot.title = element_text(size = 8, margin = margin(0,0,0.3,0))) +
    labs(
        title = "MAG expression",
        x = "",
        y = expression(log[2] ~ "(CPM)")
      )  +
    scale_fill_manual(values=alpha(c(colors),0.5)) #  scale_fill_manual(values = alpha(c("blue", "red"), 0.2))
MAG_exp_plot
```
### HBB
```{r HBB}
exp_data <- subset(df, gene_id == "ENSG00000244734.4")
HBB_exp_plot <- ggplot(exp_data, aes(LBD.type, value, fill = LBD.type)) +
    geom_violin() +
    geom_boxplot(width = 0.1, outlier.shape=NA, color = "black") + 
   # geom_jitter(shape=16, position=position_jitter(0.2), alpha = 0.1) +
    theme_bw() +
    theme(legend.position = "none", 
          plot.margin = margin(0.1, 0.2, 0.5, 0.3, "cm"), #t,r,b,l
             axis.title.x = element_text(size = 8),
             axis.text.x = element_text(size = 8),
             axis.title.y = element_text(size = 8),
             axis.text.y = element_text(size = 8),
             plot.title = element_text(size = 8, margin = margin(0,0,0.3,0))) +
    labs(
        title = "HBB expression",
        x = "",
        y = expression(log[2] ~ "(CPM)")
      )  +
    scale_fill_manual(values=alpha(c(colors),0.5)) #  scale_fill_manual(values = alpha(c("blue", "red"), 0.2))
HBB_exp_plot
```
### LGR6
```{r LGR6}
exp_data <- subset(df, gene_id == "ENSG00000133067.18")
LGR6_exp_plot <- ggplot(exp_data, aes(LBD.type, value, fill = LBD.type)) +
    geom_violin() +
    geom_boxplot(width = 0.1, outlier.shape=NA, color = "black") + 
   # geom_jitter(shape=16, position=position_jitter(0.2), alpha = 0.1) +
    theme_bw() +
    theme(legend.position = "none", 
          plot.margin = margin(0.1, 0.2, 0.5, 0.3, "cm"), #t,r,b,l
             axis.title.x = element_text(size = 8),
             axis.text.x = element_text(size = 8),
             axis.title.y = element_text(size = 8),
             axis.text.y = element_text(size = 8),
             plot.title = element_text(size = 8, margin = margin(0,0,0.3,0))) +
    labs(
        title = "LGR6 expression",
        x = "",
        y = expression(log[2] ~ "(CPM)")
      )  +
    scale_fill_manual(values=alpha(c(colors),0.5)) #  scale_fill_manual(values = alpha(c("blue", "red"), 0.2))
LGR6_exp_plot
```
### SERPINA5
```{r SERPINA5}
exp_data <- subset(df, gene_id == "ENSG00000188488.14")
SERPINA5_exp_plot <- ggplot(exp_data, aes(LBD.type, value, fill = LBD.type)) +
    geom_violin() +
    geom_boxplot(width = 0.1, outlier.shape=NA, color = "black") + 
   # geom_jitter(shape=16, position=position_jitter(0.2), alpha = 0.1) +
    theme_bw() +
    theme(legend.position = "none", 
          plot.margin = margin(0.1, 0.2, 0.5, 0.3, "cm"), #t,r,b,l
             axis.title.x = element_text(size = 8),
             axis.text.x = element_text(size = 8),
             axis.title.y = element_text(size = 8),
             axis.text.y = element_text(size = 8),
             plot.title = element_text(size = 8, margin = margin(0,0,0.3,0))) +
    labs(
        title = "SERPINA5 expression",
        x = "",
        y = expression(log[2] ~ "(CPM)")
      )  +
    scale_fill_manual(values=alpha(c(colors),0.5)) #  scale_fill_manual(values = alpha(c("blue", "red"), 0.2))
SERPINA5_exp_plot

```
``
```{r}
```
