---
title: "LBD CWOW Brain Single Nucleus Find Markers"
author: "Kimberly Olney"
date: "06/08/2023"
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

# Set working directory
```{r working_directory}
knitr::opts_knit$set(root.dir = ".")
```
# Libraris, paths, colors
```{r}
source(here::here("snRNA/scripts/R", "file_paths_and_colours.R"))
tissue <- "Brain"
treatment <- "pilot"
```

```{r}
dataObject.annotated <- readRDS(paste0("../../rObjects/",tolower(tissue),"_annotated_harmony.rds"))
```
# Find markers - Differential expression 
Differential expressed genes across conditions
Identify differential expressed genes across conditions
Now that we’ve aligned the beads and debris cells, we can start to do comparative analyses and look at the differences induced by the wet lab protocol. One way to look broadly at these changes is to plot the average expression of both the beads and debris cells and look for genes that are visual outliers on a scatter plot. 
```{r DE, warning=FALSE, message=FALSE}
celltype <- unique(dataObject.annotated$annotated_clusters)
df <- data.frame()

for (i in celltype) {
  print(i) # inspect 
  # subset object and set idents
  cluster <- subset(dataObject.annotated, annotated_clusters == i)
  Idents(cluster) <- "sample"
  # differential expression
  markers <- FindMarkers(object = cluster,
                         ident.1 = "beads",
                         ident.2 = "debris",
                         test.use = "MAST",
                         verbose = TRUE,
                         assay = "RNA")
  # if no markers found go to next iteration
  if(nrow(markers) == 0) {
    print(paste0("no markers for ", i))
    next
  }
  # add markers to master df for each comparison
  markers$cluster <- i
  markers$gene <- rownames(markers)
  # combine
  df <- rbind(df, markers)
  # cleanup
  remove(cluster)
}
# reformat table
colnames(df)[c(3,4)] <- c("percent_beads","percent_debris")
rownames(df) <- 1:nrow(df)
df$percent_difference <- abs(df[,3] - df[,4])
df <- df[,c(6,7,1,5,2,3,4,8)]
  
# write table
path <- "../../results/seurat/DEGs/beads_vs_debris_DEGs.tsv"
write.table(df, path, quote = FALSE, row.names = FALSE, sep = "\t")
```

## Volcano
```{r volcano_plot, message=FALSE, eval=FALSE}
all_clusters <- unique(dataObject.annotated$annotated_clusters)
comparison <- c("beads_vs_debris")
for (i in comparison) {
  group1_vs_group2 <-
    read.delim(paste0("../../results/seurat/DEGs/", i,
                      "_DEGs.tsv"))
  print(i)
  color_values <- vector()
  max <- nrow(group1_vs_group2)
  for (row in 1:max) {
    if (group1_vs_group2$p_val_adj[row] < 0.05) {
      if (group1_vs_group2$avg_log2FC [row] > 0.25) {
        color_values <- c(color_values, 1)
      }
      else if (group1_vs_group2$avg_log2FC[row] < -.25) {
        color_values <- c(color_values, 2)
      }
      else {
        color_values <- c(color_values, 3)
      }
    }
    else{
      color_values <- c(color_values, 3)
    }
  }
  group1_vs_group2$color_adjpval_0.05 <- factor(color_values)
  # loop through clusters
  for (j in all_clusters) {
    data <- subset(group1_vs_group2, cluster == j)
    num <- subset(data, (p_val_adj < 0.05 & avg_log2FC < -.25)  | (p_val_adj < 0.05 & avg_log2FC > .25 ))
    num <- nrow(num)
    if (num != 0) {
      up <- data[data$color_adjpval_0.05 == 1,]
      up10 <- up[1:10,]
      upFold <- subset(up, avg_log2FC > 1)
      upFold <- upFold[!(upFold$gene %in% up10$gene),]
      down <- data[data$color_adjpval_0.05 == 2,]
      down10 <- down[1:10,]
      downFold <- subset(down, avg_log2FC < -1)
      downFold <- downFold[!(downFold$gene %in% down10$gene),]
    if (!1 %in% unique(data$color_adjpval_0.05)) {
      my_colors <- c("blue", "gray")
    } else if (!2 %in% unique(data$color_adjpval_0.05)) {
      my_colors <- c("red", "gray")
    } else if (!1 %in% unique(data$color_adjpval_0.05) &&
               !2 %in% unique(data$color_adjpval_0.05)) {
      my_colors <- c("gray")
    } else {
      my_colors <- c("red", "blue", "gray")
    }
    hadjpval <- (-log10(max(data$p_val[data$p_val_adj < 0.05],
                            na.rm = TRUE)))
    negFC <- c(-.25)
    posFC <- c(.25) 
    print(j)
    }
  i <- gsub("_vs_", " vs ", i)
  p <-
      ggplot(data = data,
             aes(
               x = avg_log2FC,
               y = -log10(p_val),
               color = color_adjpval_0.05
             )) +
      geom_point(alpha = 0.8, size = 1) +
      theme_bw() +
      theme(legend.position = "none") +
      scale_color_manual(values = my_colors) +
      labs(
        title = "",
        x = expression(log[2](FC)),
        y = expression(-log[10] ~ "(" ~ italic("p") ~ "-value)")
      )  +
      geom_hline(yintercept = hadjpval,
                 #  horizontal line
                 colour = "#000000",
                 linetype = "dashed") +
      geom_vline(xintercept = negFC,
                 #  horizontal line
                 colour = "#000000",
                 linetype = "dashed") +
      geom_vline(xintercept = posFC,
                 #  horizontal line
                 colour = "#000000",
                 linetype = "dashed") +
      ggtitle(paste0(i, "
", j)) +
      geom_text_repel(
        data = up10,
        aes(
          x = avg_log2FC,
          y = -log10(p_val),
          label = gene
        ),
        color = "maroon",
        fontface = "italic",
        size = 3,
        max.overlaps = getOption("ggrepel.max.overlaps", default = 10)
      ) +
      geom_text_repel(
        data = upFold,
        aes(
          x = avg_log2FC,
          y = -log10(p_val),
          label = gene
        ),
        color = "maroon",
        fontface = "italic",
        size = 3,
        max.overlaps = getOption("ggrepel.max.overlaps", default = 10)
      ) +
      geom_text_repel(
        data = down10,
        aes(
          x = avg_log2FC,
          y = -log10(p_val),
          label = gene
        ),
        color = "navyblue",
        fontface = "italic",
        size = 3,
        max.overlaps = getOption("ggrepel.max.overlaps", default = 10)
      ) +
      geom_text_repel(
        data = downFold,
        aes(
          x = avg_log2FC,
          y = -log10(p_val),
          label = gene
        ),
        color = "navyblue",
        fontface = "italic",
        size = 3,
        max.overlaps = getOption("ggrepel.max.overlaps", default = 10)
      )
    p
    i <- gsub(" vs ", "_vs_", i)
    j <- gsub(" / ", "&", j)
    # save
    path <- paste0("../../results/seurat/volcano/", i, "_", j,
        "")
    pdf(paste0(path, ".pdf"),
        height = 5,
        width = 5)
    print(p)
    dev.off()
    print(paste("i =", i))
  } 
}
```
## Scatter plot
```{r}
theme_set(theme_cowplot())
endothelial.cells <- subset(dataObject.annotated, idents = "Endothelial")
Idents(endothelial.cells) <- "sample"
avg.endothelial.cells <- as.data.frame(log1p(AverageExpression(endothelial.cells, verbose = FALSE)$RNA))
avg.endothelial.cells$gene <- rownames(avg.endothelial.cells)

genes.to.label = c("HERC4", "EYS")
p1 <- ggplot(avg.endothelial.cells, aes(beads, debris)) + geom_point() + ggtitle("Endothelial") +
  geom_abline(color = "gray40")
p1 <- LabelPoints(plot = p1, points = genes.to.label, repel = TRUE)
p1

subset(avg.endothelial.cells, gene == "EYS")
```

```{r cleanup}
# clean up
remove(up, up10, upFold, group1_vs_group2, downFold, down10, data, p)
```

# Upset of DEGs
### Function to create tables
```{r}
fromList <- function (input) {
  # Same as original fromList()...
  elements <- unique(unlist(input))
  data <- unlist(lapply(input, function(x) {
      x <- as.vector(match(elements, x))
      }))
  data[is.na(data)] <- as.integer(0)
  data[data != 0] <- as.integer(1)
  data <- data.frame(matrix(data, ncol = length(input), byrow = F))
  data <- data[which(rowSums(data) != 0), ]
  names(data) <- names(input)
  # ... Except now it conserves your original value names!
  row.names(data) <- elements
  return(data)
}

makePaddedDataFrame <- function(l, ...) {
  maxlen <- max(sapply(l, length))
  data.frame(lapply(l, na.pad, len = maxlen), ...)
}
```
### Plot
```{r}
data <- read.delim(paste0("../../results/seurat/DEGs/beads_vs_debris_DEGs.tsv"))
up <- subset(data, avg_log2FC > 0.25 & p_val_adj <= 0.05 & percent_difference > 0.2)
down <- subset(data, avg_log2FC < -0.25 & p_val_adj <= 0.05 & percent_difference > 0.2)
path <- "../../results/seurat/goi/beads_vs_debris_DEGs_up_logFC0.25_FDRq0.05_perdiff_0.2.tsv"
write.table(up, path, quote = FALSE, row.names = FALSE, sep = "\t")
path <- "../../results/seurat/goi/beads_vs_debris_DEGs_down_logFC0.25_FDRq0.05_perdiff_0.2.tsv"
write.table(down, path, quote = FALSE, row.names = FALSE, sep = "\t")


list_input <- list("up" = LBDvsControl_down$gene_name,
                   "down" = LBDvsControl_up$gene_name,
                   "AD down-regulated" = ADvsControl_down$gene_name,
                   "AD up-regulated" = ADvsControl_up$gene_name,
                   "PA down-regulated" = PAvsControl_down$gene_name,
                   "PA up-regulated" = PAvsControl_up$gene_name)
data <- fromList(list_input)

upset_gene <- upset(data, set_sizes = FALSE,
      c('LBD down-regulated',
        'AD down-regulated',
        'PA down-regulated',
        'LBD up-regulated',
        'AD up-regulated',
        'PA up-regulated'),
      themes=upset_modify_themes(
        list('intersections_matrix'=theme(text=element_text(size=10),
              plot.margin = margin(0, 0, 0, 0, "cm")),
          'overall_sizes' = theme(axis.text.x = element_text(size =10), 
              plot.margin = margin(0, 0, 0, 0, "cm")))),
  queries=list(
    upset_query(set='LBD down-regulated', fill='blue'),
    upset_query(set='AD down-regulated', fill='blue'),
    upset_query(set='PA down-regulated', fill='blue'),
    upset_query(set='LBD up-regulated', fill='red'),
    upset_query(set='AD up-regulated', fill='red'),
    upset_query(set='PA up-regulated', fill='red')
  ),
  intersections = list(c('LBD down-regulated','AD down-regulated', 'PA down-regulated'), 
                       c('LBD down-regulated', 'AD down-regulated'), 
                       c('LBD down-regulated', 'PA down-regulated'), 
                       c('AD down-regulated', 'PA down-regulated'), 
                       'LBD down-regulated',
                       'AD down-regulated', 
                       'PA down-regulated',
                       c('LBD up-regulated','AD up-regulated', 'PA up-regulated'), 
                       c('LBD up-regulated', 'AD up-regulated'), 
                       c('LBD up-regulated', 'PA up-regulated'), 
                       c('AD up-regulated', 'PA up-regulated'), 
                       'LBD up-regulated',
                       'AD up-regulated', 
                       'PA up-regulated',
                       c('AD up-regulated', 'PA down-regulated'), 
                       c('AD down-regulated', 'PA up-regulated')
  ),
  base_annotations=list(
    'Intersection size'=(
        intersection_size(
        size=2,
        text = list(size = 3),
        text_mapping = aes(),
        bar_number_threshold=3,  # show all numbers on top of bars
        width=0.3,   # reduce width of the bars
        mapping=aes(fill='bars_color')
      )
      + scale_fill_manual(values=c('grey'), guide='none')
      + scale_y_continuous(expand=expansion(mult=c(0, 0.5)))
      + theme(axis.text.y = element_text(size = 10),
              axis.title.y = element_text(size = 10),
          panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),
          plot.margin = margin(0, 0, 0, 0, "cm"), 
          axis.line=element_line(colour='black')))),
  matrix=intersection_matrix(
      geom=geom_point(
        shape='circle filled',
        size=2,
        stroke=0.45)) +
        theme(axis.text.y = element_text(size = 10),
              axis.title.y = element_text(size = 10), 
              plot.margin = margin(0, 0, 0, 0, "cm")),
    sort_sets=FALSE,
  sort_intersections=FALSE)
upset_gene

```

